### 2009-11-30 papers/mt_marathon10_SuffixArray/paper.tex
Working on suffix array paper
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1167 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

%% Sample file $Id: PBML_article.tex 92 2007-12-18 22:59:46Z zw $
% 
% This is a sample file for the PBML article.
% 
% The Prague Bulletin of Mathematical Linguistics is typeset by XeLaTeX. This
% means that it is not guaranteed that the line and page breaks will come up
% exactly the same in the final printed version as in your computer. It is
% not required that you use XeLaTeX and the same fonts as will be used in the
% journal but it is better to do so if you could. You can even used standard
% LaTeX if XeLaTeX is not available on your computer.
% 
% 
% Document class
% ==============
% 
% The Prague Bulletin of Mathematical Linguistics uses its own document class.
% You will thus start your document with:

\documentclass{pbml}

% If you process the document with XeLaTeX, fonts Minion Pro (distributed
% with Adobe Reader) and DejaVu
% (http://dejavu.sourceforge.net/wiki/index.php/Main_Page). If you do not
% have these fonts and are not able to install them, you can instruct XeLaTeX
% to use its default fonts by issuing \document[nofonts]{pbml}.
% 
% 
% Packages
% ========
% 
% I this place you will load required packages by the \usepackage commands.
% Remember that other articles may require other packages and it is known
% that some of them are in conflict. Please use only those packages that you
% really need.
% 
% The following packages are loaded automatically by the class:
% euler (for math fonts)
% graphicx (for inclusion of images)
% multicol (for multicolumn typesetting)
% fullname (for bibliography citations)
% 
% If XeLaTeX is used, fontspec and xltxtra are loaded as well.
% 
% If standard LaTeX is used, the font encoding is switched to T1 and Latin
% Modern fonts are used if they are found in your TeX distribution.
% 
% 
% Graphics
% ========
% 
% The most portable graphics package is tikz which is a front end to pgf.
% This package is fully supported both in standard LaTeX and XeLaTeX.
% PSTricks is not currently available for XeLaTeX. It will most probably be
% supported in the future.
% 
% 
% Fonts
% =====
% 
% If you want to typset examples in east Asian scripts, you have to use
% OpenType Unicode fonts that are freely redistributable and you have to
% include them with your article. If you must use nonfree or non-Unicode
% fonts, you must supply the examples as EPS or PDF with fonts embedded in
% the files (as the required subset).
% 
% 
% Definitions
% ===========
% 
% You can use your own macros defined by \newcommand, \providecommand,
% \DeclareRobustCommand (or even by \def) as well as environments declared by
% \newenvironment. These definitions will automatically be local to your
% article and will not clash with other parts of the journal. You can also
% use \newcounter in order to declare your own counters.
% 
% 
% Beginning of the document
% =========================
% 
% Use the standard command:

%\usepackage{newcite}

%\newcommand{\url}{\begin{verbatim}URL\end{verbatim}}
\usepackage{url}
\usepackage{moreverb}

\newcommand{\newcite}{\namecite}

\begin{document}

% Document title
% ==============
% 
% Due to journal organization the article title and authors must be specified
% AFTER \begin{document}. Give the article title:

\title{Hierarchical Phrase-Based Grammar Extraction in Joshua}

% You may optionally specify a subtitle:

\subtitle{Suffix Arrays and Prefix Trees}

% Now put the author names and adresses, one command for each authors. The
% order of authors printed below the title will be the same as the order of
% your commands:

\author{firstname=Lane, surname=Schwartz,
       address={University of Minnesota, Minneapolis}}

% The braces are mandatory only if the value contains a comma as in the first
% author's address but they never make any harm, you can safely put all
% values into braces.
% 
% The title and authors' names are used in the running head. If they are
% long, you should define short versions. These definitions are optional. You
% define them only if they are needed. The example follows:

\shorttitle{Short}
\shortauthor{L. Schwartz}

% Now print the title by:

\maketitle

% Abstract
% ========
% 
% The abstract is placed within the "abstract" environment. It is a mandatory
% part of the article.

\begin{abstract}
The abstract of the article...
\end{abstract}

% The body of the article
% =======================

\section{Introduction}

Phrase-based model \cite{Koehn2003,Marcu2002,Och2004,Och1999}.

Suffix arrays  \cite{Manber1990}. Suffix arrays have been shown to be useful in efficient natural language processing; \newcite{Yamamoto2001} show how suffix arrays can be used to efficiently calculate the term frequency and document frequency of every substring in a corpus \footnote{The source code for performing these calculations is available for download at \url{http://www.mibel.cs.tsukuba.ac.jp/~myama/tfdf/index.html}, and is reproduced as appendix material in \cite{Yamamoto2001}}.

Extracting phrases \cite{Koehn2003,Kumar2006}.

Example-based machine translation (EBMT) \cite{Nagao1981,Sato1990,Somers2003}. EBMT using efficient data structures for pattern matching \cite{Brown2004}. Recent implementation that uses suffix arrays in Cunei \cite{Phillips2009}.

Suffix arrays for phrase-based model \newcite{Callison-Burch2005b} and \newcite{Zhang2005}. Extended by \cite{Lopez2008} to use deterministic sampling; this is necessary for the approach to integrate well with the minimum error rate training (MERT) procedure \cite{Och2003c} used to tune model parameters.

Inversion transduction grammar (ITG) \cite{Wu1996}.

Hierarchical phrase-based model introduced by \cite{Chiang2005,Chiang2007}.

Suffix arrays for pattern matching for phrases with gaps \cite{Lopez2007}, described in more detail in \cite{Lopez2008}.

Trie data structure \cite{Fredkin1960}.

Tight versus loose extraction heuristic \cite{Ayan2006a}.

\section{Related Work}

Moses has suffix array \footnote{\url{http://sourceforge.net/projects/mosesdecoder}}.

Joshua \cite{Joshua-WMT}\footnote{\url{http://sourceforge.net/projects/joshua}}.

Cunei \cite{Phillips2009}\footnote{\url{http://sourceforge.net/projects/cunei}}.

\section{Data Structures}

\subsection{Symbol Table}

\subsection{Suffix Array}

\subsection{Prefix Tree with Suffix Links}

\subsection{Memory Mapping}

\section{Using Joshua}

Grammar extraction using command line.

Grammar extraction using Apache Ant \cite{ant1.7.1}.

\begin{figure}
\begin{verbatimtab}[4]
<project name="Joshua Grammar Extraction in Ant" default="extract">

	<property name="classes.dir" value="bin"/>

	<taskdef name="extractRules" 
		classname="joshua.prefix_tree.ExtractRules" 
		classpath="${classes.dir}"/>

	<taskdef name="compileJosh" 
		classname="joshua.corpus.suffix_array.Compile" 
		classpath="${classes.dir}"/>

	<target name="compile" 
		description="Compile aligned parallel corpus
			 into memory-mapped files">

		<compileJosh
			sourceCorpus="/path/to/source.txt"
			targetCorpus="/path/to/target.txt"
			alignments="/path/to/align.txt"
			outputDir="/path/to/model.josh"
		/>
	</target>

	<target name="extract" 
		description="Extract a translation grammar 
			from a compiled parallel corpus">

		<extractRules 
			joshDir="/path/to/model.josh" 
			testFile="/path/to/target-test.txt" 
			outputFile="/path/to/test-set-specific.grammar"
		/>    
	</target>

</project>
\end{verbatimtab}
\caption{Sample Apache Ant build file defining how to extract a translation grammar from an aligned parallel corpus.}
\end{figure}

\begin{figure}
\begin{description}

	\item[joshDir] Path to directory containing the binary files representing memory-mappable aligned parallel corpus.

	\item[testFile] Path to plain text file containing a source language test file. The grammar extracted by {\tt extractRules} will be all of the rules required to translate the sentences in this test file.

	\item[outputFile] Path where extracted grammar will be placed. This grammar will consist of all of the rules required to translate the sentences in the test file defined in the {\tt testFile} option.
\end{description}
\caption{Mandatory options must be configured in the {\tt extractRules} target.}
\end{figure}

\begin{figure}
\begin{description}

	\item[maxPhraseSpan] Defines the maximum span (in the source corpus) of any extracted SCFG rule. Default value is "10".

	\item[maxPhraseLength] Defines the maximum number of tokens (terminals plus nonterminals) allowed in the source right-hand side of any extracted SCFG rule. Default value is "10".

	\item[maxNonterminals] Defines the maximum number of nonterminal symbols allowed in the source side of any synchronous context-free rule extracted. Note: the number and type of nonterminals is the same in the source and target right-hand sides of a SCFG rule. Default value is "2".

	\item[cacheSize]

	\item[encoding] Defines the file encoding scheme of the input test file and the output grammar file. Default value is "UTF-8".

	\item[ruleSampleSize] When extracting SCFG rules for a given source language phrase, this option defines the number of instances of that source phrase will be sampled from the source training corpus for use in rule extraction. Default value is "300".

	\item[startingSentence] Defines the (1-based) sentence index in the test file where grammar extraction will begin. Default value is 1, indicating that a grammar will be extracted capable of translating sentences starting with the first sentence in the test file.

	\item[maxTestSentences] Defines the number of sentences in the test file over which grammar extraction will be performed. Default value is {\tt Integer.MAX\_VALUE}. For example, given a test file of 100 sentences, the options {\tt startingSentence="51"  maxTestSentences="25"} would cause grammar extraction to be performed over test sentences 51--75.

\end{description}
\caption{Options can be configured in the {\tt extractRules} target.}
\end{figure}

\begin{figure}
\begin{description}

	\item[sentenceInitialX] Boolean option indicates whether rules with an initial source-side nonterminal should be extracted from phrases at the start of a sentence, even though such rules do not have supporting corporal evidence. This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true. Default value is "true".

	\item[sentenceFinalX] Boolean option indicates whether rules with a final source-side nonterminal should be extracted from phrases at the end of a sentence, even though such rules do not have supporting corporal evidence. This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true. Default value is "true".

	\item[edgeXViolates] Boolean option indicates whether rules with an initial or final source-side nonterminal should be extracted when the source corpus phrase span for the rule, discounting the initial or final nonterminal, is already equal to the maximum phrase span value. Since nonterminals conceptually correspond to elided elements in the training corpus, setting this value to true allows phrases which have a longer phrase span than the maximum allowed phrase span. This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true. Default value is "true".

	\item[requireTightSpans] Boolean option indicates whether

\end{description}
\caption{Options can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}} .}
\end{figure}

\begin{figure}
\begin{description}
	\item[keepTree] Boolean option indicates whether the prefix tree should persist from sentence to sentence during grammar extraction. If set to false, a new prefix tree will be created to process each sentence in the test file. Default value is "true".

	\item[printPrefixTree] Boolean option indicates whether a representation of the prefix tree should be printed to standard output. If set to true, the tree will be printed after processing each sentence in the test file. Default value is "false".

\end{description}
\caption{Additional options can be configured in the {\tt extractRules} target to affect behavior of the prefix tree.}
\end{figure}

\begin{figure}
\begin{description}
	\item[sourceFileName] Path to binary file containing integer-ized memory-mappable source language training corpus. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/source.corpus}.

	\item[targetFileName] Path to binary file containing integer-ized memory-mappable target language training corpus. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/target.corpus}.

	\item[alignmentsFileName] Path to binary file containing integer-ized memory-mappable source-to-target word alignments data. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/alignment.grids}. The format of this file is defined in {\tt joshua.corpus.alignment.AlignmentGrids\#writeExternal} and in the {\tt joshua.corpus.alignment.MemoryMappedAlignmentGrids} constructors.
\end{description}
\caption{Additional options can be configured in the {\tt extractRules} target. These options are normally set automatically as a result of setting the {\tt joshDir} option. They are listed here for completeness.}
\end{figure}

% 
% The PBML class is modelled after the standard article class. This means
% that you can use almost everything that is allowed in articles as described
% in the textbooks of LaTeX. We support sectioning commands \section,
% \subsection and \subsubsection. In addition to standard LaTeX we support
% \subsection and \subsubsection with an empty title (i.e. just with empty
% braces {} that must not even contain a space). Such subsections will be
% automatically numbered and properly formatted.
% 
% 
% Figures and tables
% ==================
% 
% Use figure and table environments. The \caption command is redefined in
% order to comply with the graphic design.
% 
% 
% Inclusion of images and other files
% ===================================
% 
% All files required for your articles must reside in the same directory. You
% are allowed to use subdirectories only exceptionally. When loading the file
% you must use a relative path, never the full path. Load the images by
% \includegraphics and other files by \input, never use \include.
% 
% 
% Cross references
% ================
% 
% You can safely used \label, \ref and \pageref with any identifiers. These
% macros are modified in order not to collide with identifiers in other
% articles.
% 
% 
% Bibliography
% ============
% 
% You may either enter the bibliography manually, possibly making use of
% \label and \ref, or use BibTeX. The bibliography style is set
% automatically. The \cite command is also modified so that it does not
% collide with other articles. You process the bibliography by BibTeX in the
% standard way and include it by:

\bibliography{../bibliography}

% End of the article
% ==================
% 
% The article must end with:

\end{document}

### 2009-11-30 papers/mt_marathon10_SuffixArray/paper.tex
More suffix array paper
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1168 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
+\author{firstname=Chris, surname=Callison-Burch,
+       address={Johns Hopkins University, Baltimore}}
+
 % The braces are mandatory only if the value contains a comma as in the first
@@
 
-\shorttitle{Short}
-\shortauthor{L. Schwartz}
+\shorttitle{Grammar Extraction}
+\shortauthor{L. Schwartz, C. Callison-Burch}
 
@@
 \subsection{Memory Mapping}
+
+\section{Algorithms}
+
+\subsection{Caching}
+Discuss caching procedure
+
+\subsection{Query Intersection}
+\label{query-intersection}
+Discuss query intersect
+
+\begin{figure}
+
+Put algorithm here.
+
+\caption{Query intersection algorithm implemented in Joshua.
+This algorithm is based on a corrected version (Lopez, p.c.) of query intersection {\protect \cite{Lopez2008}}, with minor modifications as noted in section \ref{query-intersection}}
+\end{figure}
 

### 2009-11-30 papers/mt_marathon10_SuffixArray/paper.tex
Added algorithm for Query_Intersection
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1172 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \usepackage{moreverb}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \begin{figure}
+%\begin{algorithm}
+%\caption{The basic {\sc QueryIntersect} algorithm}
+\noindent {\bf Algorithm} {\sc Query\_Intersect}
 
-Put algorithm here.
+\noindent {\bf Input:} Sorted list of corpus locations matching source language pattern $a\alpha$:  $M_{a\alpha}$
 
-\caption{Query intersection algorithm implemented in Joshua.
-This algorithm is based on a corrected version (Lopez, p.c.) of query intersection {\protect \cite{Lopez2008}}, with minor modifications as noted in section \ref{query-intersection}}
+\noindent {\bf Input:} Sorted list of corpus locations matching source language pattern $\alpha{}b$:  $M_{\alpha{}b}$
+
+\begin{algorithmic}[1]
+\Function{Query\_Intersect}{$M_{a\alpha}$,  $M_{\alpha{}b}$}
+	\State $M_{a\alpha{}b} \leftarrow \emptyset$ \Comment{Result list is initially empty}
+	\State $I \leftarrow |M_{a\alpha}|$ \Comment{Number of instances of pattern $a\alpha$ in the source corpus}
+	\State $J \leftarrow |M_{\alpha{}b}|$  \Comment{Number of instances of pattern $\alpha{}b$ in the source corpus}
+	\State $j \leftarrow 0$
+	\State $i \leftarrow 0$
+
+	\While {$i<I$ and $j<J$}
+		\State \Comment{Ignore elements in $M_{\alpha{}b}$ that are}
+		\State \Comment{too distant to intersect with $M_{a\alpha}[i]$}
+		\While {$j<J$ and $M_{a\alpha}[i] \ddot{>} M_{\alpha{}b}[j]$}
+			\State $j \leftarrow j + 1$ 
+		\EndWhile
+
+		\State  \Comment{Verify that the corpus index}
+		\State \Comment{for the first terminal symbol}
+		\State $k \leftarrow i$ \Comment{in pattern $\alpha{}b$ is the same }
+		%\State \Comment{is the same as the corpus index for the first terminal symbol in pattern $\alpha{}b$ at }
+		\While {$M_{\alpha{}b}[i]_{,1}$ = $M_{\alpha{}b}[k]_{,1}$} \Comment{for locations $M_{\alpha{}b}[i]$ and $M_{\alpha{}b}[k]$}
+			\State $\ell \leftarrow j$
+			\While {$\ell < J$ and not $M_{a\alpha}[i] \ddot{<} M_{\alpha{}b}[\ell]$}
+				\If {$M_{a\alpha}[i] \ddot{=} M_{\alpha{}b}[\ell]$}
+					\State {Intersect $M_{a\alpha}[i]$ with $M_{\alpha{}b}[\ell]$ and append result to  $M_{a\alpha{}b}$}
+				\EndIf
+				\State $\ell \leftarrow \ell + 1$ \Comment{Proceed to next element in $M_{\alpha{}b}$}
+			\EndWhile
+			\State {$i \leftarrow i + 1$} \Comment{Proceed to next element in $M_{a\alpha}$}
+		\EndWhile
+	\EndWhile
+	\State \Return {$M_{a\alpha{}b}$}
+\EndFunction
+\end{algorithmic}
+
+\noindent {\bf Result:} Sorted list of corpus locations matching source language pattern $a\alpha{}b$ : $M_{a\alpha{}b}$
+
+%\end{algorithm}
+\caption{Query intersection algorithm implemented in Joshua. This algorithm is adapted from a corrected version (Lopez, p.c.) of query intersection {\protect \cite{Lopez2008}}.}
 \end{figure}
@@
 	<property name="classes.dir" value="bin"/>
+	<property name="experiment.dir" value="/path/to/my/data"/>
 
@@
 		<compileJosh
-			sourceCorpus="/path/to/source.txt"
-			targetCorpus="/path/to/target.txt"
-			alignments="/path/to/align.txt"
-			outputDir="/path/to/model.josh"
+			sourceCorpus="${experiment.dir}/source.txt"
+			targetCorpus="${experiment.dir}/target.txt"
+			alignments="${experiment.dir}/align.txt"
+			outputDir="${experiment.dir}/model.josh"
 		/>
@@
 		<extractRules 
-			joshDir="/path/to/model.josh" 
-			testFile="/path/to/target-test.txt" 
-			outputFile="/path/to/test-set-specific.grammar"
+			joshDir="${experiment.dir}/model.josh" 
+			testFile="${experiment.dir}/target-test.txt" 
+			outputFile="${experiment.dir}/test-set-specific.grammar"
 		/>    
@@
 \end{verbatimtab}
-\caption{Sample Apache Ant build file defining how to extract a translation grammar from an aligned parallel corpus.}
+\caption{Sample Apache Ant build file defining how to extract a translation grammar from an aligned parallel corpus.
+This example assumes that the files {\tt source.txt}, {\tt target.txt}, {\tt align.txt} and {\tt target-test.txt} already exist in the directory specified in the {\tt experiment.dir} variable.
+Running ant with this file will create the directory {\tt model.josh} and extract {\tt test-set-specific.grammar}.}
 \end{figure}
@@
 
-	\item[maxPhraseSpan] Defines the maximum span (in the source corpus) of any extracted SCFG rule. Default value is "10".
+	\item[maxPhraseSpan] Defines the maximum span (in the source corpus) of any extracted SCFG rule. Default value is ``10".
 
-	\item[maxPhraseLength] Defines the maximum number of tokens (terminals plus nonterminals) allowed in the source right-hand side of any extracted SCFG rule. Default value is "10".
+	\item[maxPhraseLength] Defines the maximum number of tokens (terminals plus nonterminals) allowed in the source right-hand side of any extracted SCFG rule. Default value is ``10".
 
@@
 Note: the number and type of nonterminals is the same in the source and target right-hand sides of a SCFG rule.
-Default value is "2".
+Default value is ``2".
 
@@
 
-	\item[encoding] Defines the file encoding scheme of the input test file and the output grammar file. Default value is "UTF-8".
+	\item[encoding] Defines the file encoding scheme of the input test file and the output grammar file. Default value is ``UTF-8".
 
 	\item[ruleSampleSize] When extracting SCFG rules for a given source language phrase, this option defines the number of instances of that source phrase will be sampled from the source training corpus for use in rule extraction.
-Default value is "300".
+Default value is ``300".
 
@@
 This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true.
-Default value is "true".
+Default value is ``true".
 
@@
 If set to false, a new prefix tree will be created to process each sentence in the test file.
-Default value is "true".
+Default value is ``true".
 
@@
 If set to true, the tree will be printed after processing each sentence in the test file.
-Default value is "false".
+Default value is ``false".
 
@@
 
+\section*{Acknowledgements}
+
+Special thanks to Adam Lopez for his help and advice, and for making the \LaTeX code for his algorithms available.
+
 \bibliography{../bibliography}

### 2009-12-01 papers/mt_marathon10_SuffixArray/paper.tex
Added introduction and related work sections
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1202 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-Phrase-based model \cite{Koehn2003,Marcu2002,Och2004,Och1999}.
+Significant recent research in statistical machine translation has focused on modeling translation based on contiguous strings of words, called {\textit phrases}, in the source language and corresponding phrases in the target language.
+Such phrase-based models \cite{Koehn2003,Marcu2002,Och2004,Och1999} have been very successful, and state-of-the-art machine translation systems are commonly based on these approaches.
 
-Suffix arrays  \cite{Manber1990}.
-Suffix arrays have been shown to be useful in efficient natural language processing; \newcite{Yamamoto2001} show how suffix arrays can be used to efficiently calculate the term frequency and document frequency of every substring in a corpus \footnote{The source code for performing these calculations is available for download at \url{http://www.mibel.cs.tsukuba.ac.jp/~myama/tfdf/index.html}, and is reproduced as appendix material in \cite{Yamamoto2001}}.
+A critical component in phrase-based translation is the estimation of a translation model from an aligned parallel text.
+The phrase table required for these models is typically extracted prior to translation in a preprocessing stage \cite{Koehn2003,Kumar2006}.
+This approach is found in the training scripts provided as part of the open source phrase-based Moses toolkit \cite{Moses}; the scripts perform this task, calculating a translation model from a sentence-aligned parallel corpus.
+Hierarchical phrase-based translation \cite{Chiang2005} extends phrase-based translation by allowing phrases with gaps, modeled as a synchronous context-free grammar (SCFG).
+The original Hiero implementation \cite{Chiang2007} trains its SFCG translation model in a similar preprocessing stage.
 
-Extracting phrases \cite{Koehn2003,Kumar2006}.
+By contrast, example-based machine translation (EBMT) approaches \cite{Nagao1981,Sato1990,Somers2003} are notable for their use of aligned parallel corpora at run time.
+EBMT research has successfully explored how various efficient data structures for pattern matching can be leveraged to allow the decoder to access at decode-time portions of the training text that are most relevant to the text currently being translated.
+The Cunei machine translation toolkit \cite{Phillips2009} is an open-source, statistical EBMT system that follows this approach.
 
-Example-based machine translation (EBMT) \cite{Nagao1981,Sato1990,Somers2003}.
-EBMT using efficient data structures for pattern matching \cite{Brown2004}.
-Recent implementation that uses suffix arrays in Cunei \cite{Phillips2009}.
+Suffix arrays are compact data structures which allow efficient pattern matching to be performed over all text in a corpus \cite{Manber1990}.
+\newcite{Callison-Burch2005b} and \newcite{Zhang2005} showed that suffix arrays can be adapted to allow phrase-based translation to calculate translation options for the translation model at run-time.
+For a given source phrase, this approach selected a random sample of instances of that phrase in the source corpus.
+By accessing the target corpus and word alignment data, translation options and associated model parameters can be calculated at run-time.
+\newcite{Lopez2007} showed that hierarchical phrases can also by obtained at run time using a suffix array.
+This work incorporated a trie-based grammar mechanism \cite{Fredkin1960} and used deterministic, rather than random, sampling to allow predictable integration with the minimum error rate training (MERT) procedure \cite{Och2003c} used to tune model parameters.
 
-Suffix arrays for phrase-based model \newcite{Callison-Burch2005b} and \newcite{Zhang2005}.
-Extended by \cite{Lopez2008} to use deterministic sampling; this is necessary for the approach to integrate well with the minimum error rate training (MERT) procedure \cite{Och2003c} used to tune model parameters.
+This paper implements the basic techniques described in \newcite{Lopez2008} for accessing using an aligned parallel corpus to extraction translation rules for an open source hierarchical machine translation system, Joshua \cite{Joshua-WMT}.
+The implementation described here enables users of Joshua to begin translating sentences using an aligned parallel corpus in lieu of a pre-extracted SCFG.
+This implementation can also be used during a pre-processing stage to extract an SCFG for a specific test set.
 
-Inversion transduction grammar (ITG) \cite{Wu1996}.
-
-Hierarchical phrase-based model introduced by \cite{Chiang2005,Chiang2007}.
-
-Suffix arrays for pattern matching for phrases with gaps \cite{Lopez2007}, described in more detail in \cite{Lopez2008}.
-
-Trie data structure \cite{Fredkin1960}.
-
-Tight versus loose extraction heuristic \cite{Ayan2006a}.
+The remainder of this paper is structured as follows.
+Section \ref{related-work} briefly overviews the grammar extraction mechanisms employed by other prominent open source machine translation systems.
+Section \ref{data-structures} reviews the data structures that are used for efficient run-time calculation of SCFG rules.
+Section \ref{algorithms} presents two core algorithms of this implementation that differ from those presented in \newcite{Lopez2008}.
+Finally, section \ref{using-joshua} provides details for configuring and running the Joshua grammar extraction tools.
 
 \section{Related Work}
+\label{related-work}
 
-Moses has suffix array \footnote{\url{http://sourceforge.net/projects/mosesdecoder}}.
+While example-based machine translation has long used corpus information at run-time, statistical phrase-based approaches typically include a pre-processing stage where an aligned parallel corpus is split into phrases, and parameter values are calculated for each phrase using simple relative frequency estimates.
 
-Joshua \cite{Joshua-WMT}\footnote{\url{http://sourceforge.net/projects/joshua}}.
+The Pharaoh \cite{Koehn2004} and Moses \cite{Moses} phrase-based decoders take this approach, providing users with scripts to estimate a translation model from a sentence-aligned parallel corpus.
+Similarly, the Hiero \cite{Chiang2007} and syntax-augmented machine translation (SAMT) \cite{samt2006} approaches each require a pre-processing stage to extract an SCFG-based translation model.
+Recent work in Moses (Huang and Koehn, p.c.) provides similar functionality for extracting an SCFG-based translation model during a pre-processing stage.
 
-Cunei \cite{Phillips2009}\footnote{\url{http://sourceforge.net/projects/cunei}}.
+While \newcite{Callison-Burch2005b}, \newcite{Zhang2005}, and \newcite{Lopez2008} all describe implementations of traditional phrase-based models extended to take advantage of suffix array data structures to extract phrase translation options at run-time, generally available open-source implementations of these are not available.
+Preliminary work has investigated integrating these techniques into Moses, but this work is not complete\footnote{Much of this preliminary work was conducted by Chris Callison-Burch, Andreas Eisele, Juri Ganitkevitch, and Adam Lopez at the Second Machine Translation Marathon in 2008.}.
+
+\newcite{Lopez2008} provides a fast implementation of SCFG grammar extraction for Hiero which uses suffix arrays.
+This implementation allows Hiero to use an aligned parallel corpus at run-time in lieu of a pre-extracted SCFG.
+However, this implementation is not available as free open source software.
+Cunei \cite{Phillips2009}\footnote{\url{http://sourceforge.net/projects/cunei}} is a statistical open-source EBMT system that uses suffix arrays to extract relevant phrase pairs from an aligned parallel corpus at run-time.
 
 \section{Data Structures}
+\label{data-structures}
 
@@
 \section{Algorithms}
+\label{algorithms}
 
@@
 \section{Using Joshua}
+\label{using-joshua}
 

### 2009-12-01 papers/mt_marathon10_SuffixArray/paper.tex
Work on implementation section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1203 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 Section \ref{related-work} briefly overviews the grammar extraction mechanisms employed by other prominent open source machine translation systems.
-Section \ref{data-structures} reviews the data structures that are used for efficient run-time calculation of SCFG rules.
-Section \ref{algorithms} presents two core algorithms of this implementation that differ from those presented in \newcite{Lopez2008}.
+Section \ref{data-structures} reviews the data structures and algorithms that are used for efficient run-time calculation of SCFG rules.
 Finally, section \ref{using-joshua} provides details for configuring and running the Joshua grammar extraction tools.
@@
 
-\section{Data Structures}
+\section{Implementation: Data Structures and algorithms}
 \label{data-structures}
 
-\subsection{Symbol Table}
-
-\subsection{Suffix Array}
-
-\subsection{Prefix Tree with Suffix Links}
-
-\subsection{Memory Mapping}
-
-\section{Algorithms}
-\label{algorithms}
-
-\subsection{Caching}
-Discuss caching procedure
-
-\subsection{Query Intersection}
-\label{query-intersection}
-Discuss query intersect
+\begin{figure}
+\begin{algorithmic}[1]
+\Function{Compare\_Elements}{$index1$,  $index2$, $max$, $corpusEnd$}
+	\For {$i=0$; $i<max$; $i++$}
+		\If {$index1+i < corpusEnd$ and $index2+i > corpusEnd$}
+			\State \Return{$1$}
+		\ElsIf {$index2+i < corpusEnd$ and $index1+i>corpusEnd$}
+			\State \Return{$-1$}
+		\ElsIf {$corpus[index1+i]$ is lexicographically < $corpus[index2+i]$}
+			\State \Return {$-1$}
+		\ElsIf {$corpus[index1+i]$ is lexicographically > $corpus[index2+i]$}
+			\State \Return {$1$}
+		\EndIf
+	\EndFor
+	\State \Return {$0$}
+\EndFunction
+\end{algorithmic}
+\caption{}
+\label{sort}
+\end{figure}
 
@@
 \caption{Query intersection algorithm implemented in Joshua. This algorithm is adapted from a corrected version (Lopez, p.c.) of query intersection {\protect \cite{Lopez2008}}.}
+\label{query}
 \end{figure}
+
+To extract hierarchical translation rules at run-time, the decoder must have access to the aligned parallel corpus.
+Internally, Joshua treats all source and target words as 32-bit integers.
+Each unique string that is encountered is assigned a unique integer.
+A hash map maintains the mapping from string to integer, while a corresponding list of strings maintains the mapping from integer back to string; together, these data structures comprise the {\bf symbol table}.
+
+A corpus can be considered a simple list whose size is equal to the number of words in the corpus.
+Using this approach with the symbol table, Joshua stores the source corpus as an integer array.
+An auxiliary array, with size equal to the number of sentences in the source corpus, is maintained; elements in this auxiliary sentence array indicate the corpus index where each sentence begins.
+The target corpus is likewise represented by a {\bf corpus array} and corresponding sentence array.
+
+Once the source and target corpus arrays are available, the corresponding {\bf suffix array}s can be constructed.
+Given a corpus array $c$ and a symbol table, a second array is created of equal size to the corpus array.
+This array $s$ is initialized such that $s[x] = x]$.
+Where each integer in $c$ represents a word string, each integer in $s$ represents an index into $c$.
+The contents of $s$ are sorted, using the element comparison function defined in figure \ref{sort}.
+After sorting, the indices of all instances in the corpus of any given phrase are located in a contiguous segment in the suffix array $s$.
+
+While a phrase-based decoder can simply look up any required phrase in a suffix array, hierarchical decoders must deal with discontinuous phrases that include gaps.
+To deal with such phrases, \namecite{Lopez2008} defines an incremental algorithm for constructing a specialized trie \cite{Fredkin1960} to represent the SCFG translation grammar.
+Given a source sentence, this algorithm constructs a {\bf prefix tree with suffix links} by first examining all possible contiguous source phrases, and uses the source suffix array to look up translations for contiguous phrases.
+Hierarchical phrases that consist of a contiguous phrase preceded or followed by a single nonterminal $X$ can be constructed directly from the corresponding contiguous phrase.
+In this manner, the tree is gradually constructed into a grammar containing contiguous phrases and simple hierarchical phrases.
+
+More complex hierarchical phrases are constructed according to the {\sc Query\_Intersect} function in figure \ref{query}.
+This function takes two smaller phrases $a\alpha$ and $\alpha{}b$, along with the list of indices where these phrases are located.
+These two lists can be efficiently processed to determine the locations where the two phrases intersect to form the more complex phrase $a\alpha{}b$.
+In this way, all source hierarchical phrases can be located.
+
+Each node in the prefix tree corresponds to a unique source phrase; each node stores the complete list of all indices in the source corpus where that node's phrase occurs.
+These locations are used in conjunction with the target corpus array and the word alignment data to construct SCFG translation rules.
 
@@
 \label{using-joshua}
-
-Grammar extraction using command line.
-
-Grammar extraction using Apache Ant \cite{ant1.7.1}.
 
@@
 \end{figure}
+
+Grammar extraction using command line.
+
+Grammar extraction using Apache Ant \cite{ant1.7.1}.
 

### 2009-12-02 papers/mt_marathon10_SuffixArray/paper.tex
Version submitted to MT Marathon
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1209 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \begin{abstract}
-The abstract of the article...
+While example-based machine translation has long used corpus information at run-time, statistical phrase-based approaches typically include a pre-processing stage where an aligned parallel corpus is split into phrases, and parameter values are calculated for each phrase using simple relative frequency estimates.
+This paper describes an open source implementation of the crucial algorithms presented in \cite{Lopez2008} which allow direct run-time calculation of SCFG translation rules in Joshua.
 \end{abstract}
@@
 
-Significant recent research in statistical machine translation has focused on modeling translation based on contiguous strings of words, called {\textit phrases}, in the source language and corresponding phrases in the target language.
+Significant recent research in statistical machine translation has focused on modeling translation based on contiguous strings of words, called {\it phrases}, in the source language and corresponding phrases in the target language.
 Such phrase-based models \cite{Koehn2003,Marcu2002,Och2004,Och1999} have been very successful, and state-of-the-art machine translation systems are commonly based on these approaches.
@@
 This implementation allows Hiero to use an aligned parallel corpus at run-time in lieu of a pre-extracted SCFG.
-However, this implementation is not available as free open source software.
+However, this implementation is not available as generally available open source software.
 Cunei \cite{Phillips2009}\footnote{\url{http://sourceforge.net/projects/cunei}} is a statistical open-source EBMT system that uses suffix arrays to extract relevant phrase pairs from an aligned parallel corpus at run-time.
 
-\section{Implementation: Data Structures and algorithms}
+\section{Implementation: Data Structures and Algorithms}
 \label{data-structures}
@@
 \noindent {\bf Input:} Sorted list of corpus locations matching source language pattern $\alpha{}b$:  $M_{\alpha{}b}$
+
+\ \\
 
@@
 
+\ \\
+
 \noindent {\bf Result:} Sorted list of corpus locations matching source language pattern $a\alpha{}b$ : $M_{a\alpha{}b}$
@@
 Given a corpus array $c$ and a symbol table, a second array is created of equal size to the corpus array.
-This array $s$ is initialized such that $s[x] = x]$.
+This array $s$ is initialized such that $s[x] = x$.
 Where each integer in $c$ represents a word string, each integer in $s$ represents an index into $c$.
@@
 
-\section{Using Joshua}
-\label{using-joshua}
+% Caching
+Ideally, once translation rules have been extracted for a given source phrase, those rules would be stored and not calculated again.
+Unfortunately, memory constraints typically dictate that not all rules are stored.
+Rather than storing the translation rules for a given source phrase at the corresponding node in the prefix tree, a single least-recently-used (LRU) cache is maintained.
+This cache maps from source phrase to the corresponding set of translation rules.
 
-\begin{figure}
+% Memory mapping
+Another technique used to save memory is the option of using memory-mapped data structures.
+Memory-mapped version of the corpus array, suffix array, and alignment grids data structures are implemented and used by default.
+
+\begin{figure}[hptb]
 \begin{verbatimtab}[4]
@@
 Running ant with this file will create the directory {\tt model.josh} and extract {\tt test-set-specific.grammar}.}
+\label{ant-file}
 \end{figure}
 
-Grammar extraction using command line.
-
-Grammar extraction using Apache Ant \cite{ant1.7.1}.
-
-\begin{figure}
-\begin{description}
-
-	\item[joshDir] Path to directory containing the binary files representing memory-mappable aligned parallel corpus.
-
-	\item[testFile] Path to plain text file containing a source language test file.
-The grammar extracted by {\tt extractRules} will be all of the rules required to translate the sentences in this test file.
-
-	\item[outputFile] Path where extracted grammar will be placed. This grammar will consist of all of the rules required to translate the sentences in the test file defined in the {\tt testFile} option.
-\end{description}
-\caption{Mandatory options must be configured in the {\tt extractRules} target.}
-\end{figure}
-
-\begin{figure}
-\begin{description}
-
-	\item[maxPhraseSpan] Defines the maximum span (in the source corpus) of any extracted SCFG rule. Default value is ``10".
-
-	\item[maxPhraseLength] Defines the maximum number of tokens (terminals plus nonterminals) allowed in the source right-hand side of any extracted SCFG rule. Default value is ``10".
-
-	\item[maxNonterminals] Defines the maximum number of nonterminal symbols allowed in the source side of any synchronous context-free rule extracted.
-Note: the number and type of nonterminals is the same in the source and target right-hand sides of a SCFG rule.
-Default value is ``2".
-
-	\item[cacheSize]
-
-	\item[encoding] Defines the file encoding scheme of the input test file and the output grammar file. Default value is ``UTF-8".
-
-	\item[ruleSampleSize] When extracting SCFG rules for a given source language phrase, this option defines the number of instances of that source phrase will be sampled from the source training corpus for use in rule extraction.
-Default value is ``300".
-
-	\item[startingSentence] Defines the (1-based) sentence index in the test file where grammar extraction will begin.
-Default value is 1, indicating that a grammar will be extracted capable of translating sentences starting with the first sentence in the test file.
-
-	\item[maxTestSentences] Defines the number of sentences in the test file over which grammar extraction will be performed.
-Default value is {\tt Integer.MAX\_VALUE}.
-For example, given a test file of 100 sentences, the options {\tt startingSentence="51"  maxTestSentences="25"} would cause grammar extraction to be performed over test sentences 51--75.
-
-\end{description}
-\caption{Options can be configured in the {\tt extractRules} target.}
-\end{figure}
-
-\begin{figure}
-\begin{description}
-
-	\item[sentenceInitialX] Boolean option indicates whether rules with an initial source-side nonterminal should be extracted from phrases at the start of a sentence, even though such rules do not have supporting corporal evidence.
-This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true.
-Default value is ``true".
-
-	\item[sentenceFinalX] Boolean option indicates whether rules with a final source-side nonterminal should be extracted from phrases at the end of a sentence, even though such rules do not have supporting corporal evidence.
-This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true.
-Default value is ``true".
-
-	\item[edgeXViolates] Boolean option indicates whether rules with an initial or final source-side nonterminal should be extracted when the source corpus phrase span for the rule, discounting the initial or final nonterminal, is already equal to the maximum phrase span value.
-Since nonterminals conceptually correspond to elided elements in the training corpus, setting this value to true allows phrases which have a longer phrase span than the maximum allowed phrase span.
-This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true.
-Default value is ``true".
-
-	\item[requireTightSpans] Boolean option indicates whether
-
-\end{description}
-\caption{Options can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}} .}
-\end{figure}
-
-\begin{figure}
-\begin{description}
-	\item[keepTree] Boolean option indicates whether the prefix tree should persist from sentence to sentence during grammar extraction.
-If set to false, a new prefix tree will be created to process each sentence in the test file.
-Default value is ``true".
-
-	\item[printPrefixTree] Boolean option indicates whether a representation of the prefix tree should be printed to standard output.
-If set to true, the tree will be printed after processing each sentence in the test file.
-Default value is ``false".
-
-\end{description}
-\caption{Additional options can be configured in the {\tt extractRules} target to affect behavior of the prefix tree.}
-\end{figure}
-
-\begin{figure}
-\begin{description}
-	\item[sourceFileName] Path to binary file containing integer-ized memory-mappable source language training corpus.
+%\begin{figure}
+%\begin{description}
+%	\item[sourceFileName] Path to binary file containing integer-ized memory-mappable source language training corpus.
 When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/source.corpus}.
 
-	\item[targetFileName] Path to binary file containing integer-ized memory-mappable target language training corpus.
+%	\item[targetFileName] Path to binary file containing integer-ized memory-mappable target language training corpus.
 When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/target.corpus}.
 
-	\item[alignmentsFileName] Path to binary file containing integer-ized memory-mappable source-to-target word alignments data.
+%	\item[alignmentsFileName] Path to binary file containing integer-ized memory-mappable source-to-target word alignments data.
 When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/alignment.grids}.
 The format of this file is defined in {\tt joshua.corpus.alignment.AlignmentGrids\#writeExternal} and in the {\tt joshua.corpus.alignment.MemoryMappedAlignmentGrids} constructors.
-\end{description}
-\caption{Additional options can be configured in the {\tt extractRules} target.
+%\end{description}
+%\caption{Additional options can be configured in the {\tt extractRules} target.
 These options are normally set automatically as a result of setting the {\tt joshDir} option.
 They are listed here for completeness.}
-\end{figure}
+%\end{figure}
 
@@
 
+\section{Using Joshua}
+\label{using-joshua}
+
+Given a word-aligned parallel corpus, the first step in extracting a grammar, either at run-time or during a pre-processing stage, is to compile the memory-mappable data structures to binary files on disk.
+The {\tt joshua.corpus.suffix\_array.Compile} program takes four parameters: source corpus text file, target corpus text file, word alignments text file, and an output directory path.
+This program may be run directly, or may be accessed as an Apache Ant task (see figure \ref{ant-file}).
+The output directory, by convention, is named with the suffix .josh; this directory will contain binary representations of the symbol table, source and target corpus arrays, and source and target suffix arrays.
+
+Once these binary files have been constructed into a .josh directory, the decoder can be configured to extract translation rules directly at run-time.
+To run the decoder in this mode, specify the .josh directory in the {\tt tm\_file=$\ldots$} line of the Joshua configuration file.
+
+It is often useful (especially during MERT) to explicitly extract a test-set specific grammar in a pre-processing stage.
+The {\tt joshua.prefix\_tree.ExtractRules} program performs this task.
+When run directly, this program accepts either three arguments (a compiled .josh directory, file name for grammar to extract, and test file) or five arguments (source corpus text file, target corpus text file, word alignments text file, file name for grammar to extract, and test file).
+This program can also be accessed as an Apache Ant task (see figure \ref{ant-file}).
+The following subsections document the mandatory and optional parameters that can be passed to this program through the {\tt extractRules} ant task.
+
+\subsection{Mandatory parameters}
+%\begin{figure}
+\begin{description}
+
+	\item[testFile] Path to plain text file containing a source language test file.
+The grammar extracted by {\tt extractRules} will be all of the rules required to translate the sentences in this test file.
+
+	\item[outputFile] Path where extracted grammar will be placed. This grammar will consist of all of the rules required to translate the sentences in the test file defined in the {\tt testFile} option.
+
+	\item[joshDir] Path to directory containing the binary files representing memory-mappable aligned parallel corpus.
+
+\end{description}
+
+The following parameters may be specified instead of the {\tt joshDir} parameter.
+
+\begin{description}
+	\item[sourceFileName] Path to text file containing source corpus.
+%this value is automatically set.
+Otherwise, this  to  Path to binary file containing integer-ized memory-mappable source language training corpus.
+When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/source.corpus}.
+
+	\item[targetFileName] Path to text file containing target corpus.
+
+	%\item[targetFileName] Path to binary file containing integer-ized memory-mappable target language training corpus.
+When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/target.corpus}.
+
+	\item[alignmentsFileName]  Path to text file containing word alignment data.
+
+	%\item[alignmentsFileName] Path to binary file containing integer-ized memory-mappable source-to-target word alignments data.
+When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/alignment.grids}.
+The format of this file is defined in {\tt joshua.corpus.alignment.AlignmentGrids\#writeExternal} and in the {\tt joshua.corpus.alignment.MemoryMappedAlignmentGrids} constructors.
+
+\end{description}
+%\caption{Mandatory options must be configured in the {\tt extractRules} target.}
+%\end{figure}
+
+\subsection{Optional parameters}
+%\begin{figure}
+\begin{description}
+
+	\item[maxPhraseSpan] Defines the maximum span (in the source corpus) of any extracted SCFG rule. Default value is ``10".
+
+	\item[maxPhraseLength] Defines the maximum number of tokens (terminals plus nonterminals) allowed in the source right-hand side of any extracted SCFG rule. Default value is ``10".
+
+	\item[maxNonterminals] Defines the maximum number of nonterminal symbols allowed in the source side of any synchronous context-free rule extracted.
+Note: the number and type of nonterminals is the same in the source and target right-hand sides of a SCFG rule.
+Default value is ``2".
+
+	\item[cacheSize] Maximum number of source phrases for which translation rules will be maintained in the least-recently-used (LRU) cache.
+
+	\item[encoding] Defines the file encoding scheme of the input test file and the output grammar file. Default value is ``UTF-8".
+
+	\item[ruleSampleSize] When extracting SCFG rules for a given source language phrase, this option defines the number of instances of that source phrase will be sampled from the source training corpus for use in rule extraction.
+Default value is ``300".
+
+	\item[startingSentence] Defines the (1-based) sentence index in the test file where grammar extraction will begin.
+Default value is 1, indicating that a grammar will be extracted capable of translating sentences starting with the first sentence in the test file.
+
+	\item[maxTestSentences] Defines the number of sentences in the test file over which grammar extraction will be performed.
+Default value is {\tt Integer.MAX\_VALUE}.
+For example, given a test file of 100 sentences, the options {\tt startingSentence="51"  maxTestSentences="25"} would cause grammar extraction to be performed over test sentences 51--75.
+
+\end{description}
+%\caption{Options can be configured in the {\tt extractRules} target.}
+%\end{figure}
+
+The following parameters can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}} .
+
+%\begin{figure}
+\begin{description}
+
+	\item[sentenceInitialX] Boolean option indicates whether rules with an initial source-side nonterminal should be extracted from phrases at the start of a sentence, even though such rules do not have supporting corporal evidence.
+This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true.
+Default value is ``true".
+
+	\item[sentenceFinalX] Boolean option indicates whether rules with a final source-side nonterminal should be extracted from phrases at the end of a sentence, even though such rules do not have supporting corporal evidence.
+This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true.
+Default value is ``true".
+
+	\item[edgeXViolates] Boolean option indicates whether rules with an initial or final source-side nonterminal should be extracted when the source corpus phrase span for the rule, discounting the initial or final nonterminal, is already equal to the maximum phrase span value.
+Since nonterminals conceptually correspond to elided elements in the training corpus, setting this value to true allows phrases which have a longer phrase span than the maximum allowed phrase span.
+This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true.
+Default value is ``true".
+
+	\item[requireTightSpans] Boolean option; if true, follow the heuristic from \cite{Chiang2005}: where multiple initial phrase pairs contain the same set of alignment points, consider only the smallest when performing rule extraction.
+For compatibility with \newcite{Lopez2008}, set this parameter to false.
+Default value is ``true".
+
+\end{description}
+%\caption{Options can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}} .}
+%\end{figure}
+
+Additional options can be configured in the {\tt extractRules} target to affect behavior of the prefix tree.
+
+%\begin{figure}
+\begin{description}
+	\item[keepTree] Boolean option indicates whether the prefix tree should persist from sentence to sentence during grammar extraction.
+If set to false, a new prefix tree will be created to process each sentence in the test file.
+Default value is ``true".
+
+	\item[printPrefixTree] Boolean option indicates whether a representation of the prefix tree should be printed to standard output.
+If set to true, the tree will be printed after processing each sentence in the test file.
+Default value is ``false".
+
+\end{description}
+%\caption{Additional options can be configured in the {\tt extractRules} target to affect behavior of the prefix tree.}
+%\end{figure}
+
 \section*{Acknowledgements}

### 2009-12-14 papers/mt_marathon10_SuffixArray/paper.tex
Revised Lane's text
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1241 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \begin{abstract}
-While example-based machine translation has long used corpus information at run-time, statistical phrase-based approaches typically include a pre-processing stage where an aligned parallel corpus is split into phrases, and parameter values are calculated for each phrase using simple relative frequency estimates.
+While example-based machine translation has long used corpus information at run-time, statistical phrase-based approaches typically include a preprocessing stage where an aligned parallel corpus is split into phrases, and parameter values are calculated for each phrase using simple relative frequency estimates.
 This paper describes an open source implementation of the crucial algorithms presented in \cite{Lopez2008} which allow direct run-time calculation of SCFG translation rules in Joshua.
@@
 
-Significant recent research in statistical machine translation has focused on modeling translation based on contiguous strings of words, called {\it phrases}, in the source language and corresponding phrases in the target language.
-Such phrase-based models \cite{Koehn2003,Marcu2002,Och2004,Och1999} have been very successful, and state-of-the-art machine translation systems are commonly based on these approaches.
+A significant amount of the recent research in statistical machine translation has focused on modeling translation based on contiguous strings of words, called {\it phrases}, in the source language and corresponding phrases in the target language.
+Phrase-based translation \cite{Och1999,Koehn2003,Marcu2002,Och2004} have proved to be very successful, and many state-of-the-art machine translation systems are based on these approaches.
 
-A critical component in phrase-based translation is the estimation of a translation model from an aligned parallel text.
-The phrase table required for these models is typically extracted prior to translation in a preprocessing stage \cite{Koehn2003,Kumar2006}.
-This approach is found in the training scripts provided as part of the open source phrase-based Moses toolkit \cite{Moses}; the scripts perform this task, calculating a translation model from a sentence-aligned parallel corpus.
+A critical component in phrase-based translation is the estimation of a translation model from a word-aligned parallel text.
+A phrase table containing the source phrases, their target translations and their associated probabilities that  is typically extracted in a preprocessing stage before decoding a test set \cite{Koehn2003,Kumar2006}.
+An example of this preprocessing approach is found in the training scripts provided as part of the open source phrase-based Moses toolkit \cite{Moses}.
 Hierarchical phrase-based translation \cite{Chiang2005} extends phrase-based translation by allowing phrases with gaps, modeled as a synchronous context-free grammar (SCFG).
@@
 By contrast, example-based machine translation (EBMT) approaches \cite{Nagao1981,Sato1990,Somers2003} are notable for their use of aligned parallel corpora at run time.
-EBMT research has successfully explored how various efficient data structures for pattern matching can be leveraged to allow the decoder to access at decode-time portions of the training text that are most relevant to the text currently being translated.
-The Cunei machine translation toolkit \cite{Phillips2009} is an open-source, statistical EBMT system that follows this approach.
+EBMT research has successfully explored how various efficient data structures for pattern matching \cite{Brown2004} can be leveraged to allow the decoder to access at decode-time portions of the training text that are most relevant to the text currently being translated.
+The Cunei machine translation toolkit \cite{Phillips2009} is an open source, statistical EBMT system that follows this approach.
 
@@
 \newcite{Callison-Burch2005b} and \newcite{Zhang2005} showed that suffix arrays can be adapted to allow phrase-based translation to calculate translation options for the translation model at run-time.
-For a given source phrase, this approach selected a random sample of instances of that phrase in the source corpus.
-By accessing the target corpus and word alignment data, translation options and associated model parameters can be calculated at run-time.
+A subsample of occurrences of given source phrase are used to calculate translation probabilities.
+By accessing the target corpus and word alignment data, the phrasal translations and their associated model parameters can be calculated at run-time.
 \newcite{Lopez2007} showed that hierarchical phrases can also by obtained at run time using a suffix array.
-This work incorporated a trie-based grammar mechanism \cite{Fredkin1960} and used deterministic, rather than random, sampling to allow predictable integration with the minimum error rate training (MERT) procedure \cite{Och2003c} used to tune model parameters.
+%This work incorporated a trie-based grammar mechanism \cite{Fredkin1960}.
 
-This paper implements the basic techniques described in \newcite{Lopez2008} for accessing using an aligned parallel corpus to extraction translation rules for an open source hierarchical machine translation system, Joshua \cite{Joshua-WMT}.
-The implementation described here enables users of Joshua to begin translating sentences using an aligned parallel corpus in lieu of a pre-extracted SCFG.
-This implementation can also be used during a pre-processing stage to extract an SCFG for a specific test set.
+This article reports on an implementation of the basic techniques described in \newcite{Lopez2008} that was incoporated into the open source machine translation system Joshua \cite{Joshua-WMT}.
+The implementation described here enables users of Joshua to begin translating sentences using an aligned parallel corpus without having to extract an SCFG before decoding begins.
+The advantages of using this implementation are that any input sentence can be decoded (making it appropriate for live demos or real world use), and that the data structures require much less disk space than full phrase tables.
+This comes at the cost of slower running time for the decoder itself, since phrase translations have to be calculated on the fly.
+%This implementation can also be used during a preprocessing stage to extract an SCFG for a specific test set.
 
-The remainder of this paper is structured as follows.
+%The remainder of this paper is structured as follows.
 Section \ref{related-work} briefly overviews the grammar extraction mechanisms employed by other prominent open source machine translation systems.
@@
 
-While example-based machine translation has long used corpus information at run-time, statistical phrase-based approaches typically include a pre-processing stage where an aligned parallel corpus is split into phrases, and parameter values are calculated for each phrase using simple relative frequency estimates.
+While example-based machine translation has long used corpus information at run-time, statistical phrase-based approaches typically include a preprocessing stage where an aligned parallel corpus is split into phrases, and parameter values are calculated for each phrase using simple relative frequency estimates.
+%
+The phrase-based decoders Pharaoh \cite{Koehn2004} and Moses \cite{Moses} take this approach, providing users with scripts to estimate a translation model from a sentence-aligned parallel corpus.
+Similarly, Hiero \cite{Chiang2007} and the syntax-augmented machine translation (SAMT) system \cite{samt2006} both require a preprocessing stage to extract a SCFG translation model.
+Recent work in Moses (Huang and Koehn, p.c.) provides similar functionality for extracting an SCFG-based translation model during a preprocessing stage.
 
-The Pharaoh \cite{Koehn2004} and Moses \cite{Moses} phrase-based decoders take this approach, providing users with scripts to estimate a translation model from a sentence-aligned parallel corpus.
-Similarly, the Hiero \cite{Chiang2007} and syntax-augmented machine translation (SAMT) \cite{samt2006} approaches each require a pre-processing stage to extract an SCFG-based translation model.
-Recent work in Moses (Huang and Koehn, p.c.) provides similar functionality for extracting an SCFG-based translation model during a pre-processing stage.
-
-While \newcite{Callison-Burch2005b}, \newcite{Zhang2005}, and \newcite{Lopez2008} all describe implementations of traditional phrase-based models extended to take advantage of suffix array data structures to extract phrase translation options at run-time, generally available open-source implementations of these are not available.
-Preliminary work has investigated integrating these techniques into Moses, but this work is not complete\footnote{Much of this preliminary work was conducted by Chris Callison-Burch, Andreas Eisele, Juri Ganitkevitch, and Adam Lopez at the Second Machine Translation Marathon in 2008.}.
+\newcite{Callison-Burch2005b}, \newcite{Zhang2005}, and \newcite{Lopez2008} all describe implementations of traditional phrase-based models extended to take advantage of suffix array data structures to extract phrase translation options at run-time.
+However, functional open source implementations of these have yet to be made available.
+Preliminary work has investigated integrating these techniques into Moses, but this work is not complete.\footnote{Much of this preliminary work was conducted by Chris Callison-Burch, Andreas Eisele, Juri Ganitkevitch, and Adam Lopez at the Second Machine Translation Marathon in 2008.}
 
@@
 This implementation allows Hiero to use an aligned parallel corpus at run-time in lieu of a pre-extracted SCFG.
-However, this implementation is not available as generally available open source software.
-Cunei \cite{Phillips2009}\footnote{\url{http://sourceforge.net/projects/cunei}} is a statistical open-source EBMT system that uses suffix arrays to extract relevant phrase pairs from an aligned parallel corpus at run-time.
+However, this implementation is not available as open source software due to intellectual property restrictions imposed by the University of Maryland.
+Cunei \cite{Phillips2009}\footnote{\url{http://sourceforge.net/projects/cunei}} is a statistical open source EBMT system that uses suffix arrays to extract relevant phrase pairs from an aligned parallel corpus at run-time.
 
@@
 \label{data-structures}
+
+To extract hierarchical translation rules at run-time, the decoder must have access to the aligned parallel corpus.
+Internally, Joshua treats all source and target words as 32-bit integers.
+Each unique string that is encountered is assigned a unique integer.
+A hash map maintains the mapping from string to integer, while a corresponding list of strings maintains the mapping from integer back to string.
+Together these data structures comprise the {\bf symbol table}.
+
+A corpus can be considered a simple list whose size is equal to the number of words in the corpus.
+Using this approach with the symbol table, Joshua stores the source corpus as an integer array.
+An auxiliary array, with size equal to the number of sentences in the source corpus, is maintained.
+Elements in this auxiliary sentence array indicate the corpus index where each sentence begins.
+The target corpus is likewise represented by a {\bf corpus array} and corresponding sentence array.
 
@@
 \end{figure}
+
+Once the source and target corpus arrays are available, the corresponding {\bf suffix array}s can be constructed.
+Given a corpus array $c$ and a symbol table, a second array is created of equal size to the corpus array.
+This array $s$ is initialized such that $s[x] = x$.
+Where each integer in $c$ represents a word string, each integer in $s$ represents an index into $c$.
+The contents of $s$ are sorted, using the element comparison function defined in Figure \ref{sort}.
+After sorting, the indices of all instances in the corpus of any given phrase are located in a contiguous segment in the suffix array $s$.
+
+While a phrase-based decoder can simply look up any required phrase in a suffix array, hierarchical decoders must deal with discontinuous phrases that include gaps.
+To deal with such phrases, \namecite{Lopez2008} defines an incremental algorithm for constructing a specialized trie \cite{Fredkin1960} to represent the SCFG translation grammar.
+Given a source sentence, this algorithm constructs a {\bf prefix tree with suffix links} by first examining all possible contiguous source phrases, and uses the source suffix array to look up translations for contiguous phrases.
+Hierarchical phrases that consist of a contiguous phrase preceded or followed by a single nonterminal $X$ can be constructed directly from the corresponding contiguous phrase.
+In this manner, the tree is gradually constructed into a grammar containing contiguous phrases and simple hierarchical phrases.
 
@@
 
-To extract hierarchical translation rules at run-time, the decoder must have access to the aligned parallel corpus.
-Internally, Joshua treats all source and target words as 32-bit integers.
-Each unique string that is encountered is assigned a unique integer.
-A hash map maintains the mapping from string to integer, while a corresponding list of strings maintains the mapping from integer back to string; together, these data structures comprise the {\bf symbol table}.
-
-A corpus can be considered a simple list whose size is equal to the number of words in the corpus.
-Using this approach with the symbol table, Joshua stores the source corpus as an integer array.
-An auxiliary array, with size equal to the number of sentences in the source corpus, is maintained; elements in this auxiliary sentence array indicate the corpus index where each sentence begins.
-The target corpus is likewise represented by a {\bf corpus array} and corresponding sentence array.
-
-Once the source and target corpus arrays are available, the corresponding {\bf suffix array}s can be constructed.
-Given a corpus array $c$ and a symbol table, a second array is created of equal size to the corpus array.
-This array $s$ is initialized such that $s[x] = x$.
-Where each integer in $c$ represents a word string, each integer in $s$ represents an index into $c$.
-The contents of $s$ are sorted, using the element comparison function defined in figure \ref{sort}.
-After sorting, the indices of all instances in the corpus of any given phrase are located in a contiguous segment in the suffix array $s$.
-
-While a phrase-based decoder can simply look up any required phrase in a suffix array, hierarchical decoders must deal with discontinuous phrases that include gaps.
-To deal with such phrases, \namecite{Lopez2008} defines an incremental algorithm for constructing a specialized trie \cite{Fredkin1960} to represent the SCFG translation grammar.
-Given a source sentence, this algorithm constructs a {\bf prefix tree with suffix links} by first examining all possible contiguous source phrases, and uses the source suffix array to look up translations for contiguous phrases.
-Hierarchical phrases that consist of a contiguous phrase preceded or followed by a single nonterminal $X$ can be constructed directly from the corresponding contiguous phrase.
-In this manner, the tree is gradually constructed into a grammar containing contiguous phrases and simple hierarchical phrases.
-
-More complex hierarchical phrases are constructed according to the {\sc Query\_Intersect} function in figure \ref{query}.
-This function takes two smaller phrases $a\alpha$ and $\alpha{}b$, along with the list of indices where these phrases are located.
+More complex hierarchical phrases are constructed using the {\sc Query\_Intersect} function in Figure \ref{query}.
+This function takes two smaller phrases $a\alpha$ and $\alpha{}b$ ($a$ and $b$ represent single words and $\alpha$ represents a sequence), along with the list of indices where these phrases are located.
 These two lists can be efficiently processed to determine the locations where the two phrases intersect to form the more complex phrase $a\alpha{}b$.
@@
 
-Each node in the prefix tree corresponds to a unique source phrase; each node stores the complete list of all indices in the source corpus where that node's phrase occurs.
+Each node in the prefix tree corresponds to a unique source phrase.
+Each node stores the complete list of all indices in the source corpus where that node's phrase occurs.
 These locations are used in conjunction with the target corpus array and the word alignment data to construct SCFG translation rules.
@@
 Ideally, once translation rules have been extracted for a given source phrase, those rules would be stored and not calculated again.
-Unfortunately, memory constraints typically dictate that not all rules are stored.
+Memory constraints typically dictate that not all rules are stored.
 Rather than storing the translation rules for a given source phrase at the corresponding node in the prefix tree, a single least-recently-used (LRU) cache is maintained.
@@
 Memory-mapped version of the corpus array, suffix array, and alignment grids data structures are implemented and used by default.
-
-\begin{figure}[hptb]
-\begin{verbatimtab}[4]
-<project name="Joshua Grammar Extraction in Ant" default="extract">
-
-	<property name="classes.dir" value="bin"/>
-	<property name="experiment.dir" value="/path/to/my/data"/>
-
-	<taskdef name="extractRules" 
-		classname="joshua.prefix_tree.ExtractRules" 
-		classpath="${classes.dir}"/>
-
-	<taskdef name="compileJosh" 
-		classname="joshua.corpus.suffix_array.Compile" 
-		classpath="${classes.dir}"/>
-
-	<target name="compile" 
-		description="Compile aligned parallel corpus
-			 into memory-mapped files">
-
-		<compileJosh
-			sourceCorpus="${experiment.dir}/source.txt"
-			targetCorpus="${experiment.dir}/target.txt"
-			alignments="${experiment.dir}/align.txt"
-			outputDir="${experiment.dir}/model.josh"
-		/>
-	</target>
-
-	<target name="extract" 
-		description="Extract a translation grammar 
-			from a compiled parallel corpus">
-
-		<extractRules 
-			joshDir="${experiment.dir}/model.josh" 
-			testFile="${experiment.dir}/target-test.txt" 
-			outputFile="${experiment.dir}/test-set-specific.grammar"
-		/>    
-	</target>
-
-</project>
-\end{verbatimtab}
-\caption{Sample Apache Ant build file defining how to extract a translation grammar from an aligned parallel corpus.
-This example assumes that the files {\tt source.txt}, {\tt target.txt}, {\tt align.txt} and {\tt target-test.txt} already exist in the directory specified in the {\tt experiment.dir} variable.
-Running ant with this file will create the directory {\tt model.josh} and extract {\tt test-set-specific.grammar}.}
-\label{ant-file}
-\end{figure}
 
@@
 
-Given a word-aligned parallel corpus, the first step in extracting a grammar, either at run-time or during a pre-processing stage, is to compile the memory-mappable data structures to binary files on disk.
+Given a word-aligned parallel corpus, the first step in extracting a grammar, either at run-time or during a preprocessing stage, is to compile the memory-mappable data structures to binary files on disk.
 The {\tt joshua.corpus.suffix\_array.Compile} program takes four parameters: source corpus text file, target corpus text file, word alignments text file, and an output directory path.
-This program may be run directly, or may be accessed as an Apache Ant task (see figure \ref{ant-file}).
-The output directory, by convention, is named with the suffix .josh; this directory will contain binary representations of the symbol table, source and target corpus arrays, and source and target suffix arrays.
+%This program may be run directly, or may be accessed as an Apache Ant task (see Figure \ref{ant-file}). 
+The output directory, by convention, is named with the suffix {\tt {\tt .josh}}.
+This directory stores the binary representations of the symbol table, source and target corpus arrays, and the source and target suffix arrays.
+%
+These binary files are given canonical names inside the {\tt {\tt .josh}} directory, so that the decoder can use them simply by specifying the {\tt .josh} directory in the {\tt tm\_file=$\ldots$} line of the Joshua configuration file.
 
-Once these binary files have been constructed into a .josh directory, the decoder can be configured to extract translation rules directly at run-time.
-To run the decoder in this mode, specify the .josh directory in the {\tt tm\_file=$\ldots$} line of the Joshua configuration file.
-
-It is often useful (especially during MERT) to explicitly extract a test-set specific grammar in a pre-processing stage.
+It is often useful (especially during MERT) to extract a test set specific grammar once in a preprocessing step, since that test set will be translated many times and re-extracting the grammar each time would be wasteful.
 The {\tt joshua.prefix\_tree.ExtractRules} program performs this task.
-When run directly, this program accepts either three arguments (a compiled .josh directory, file name for grammar to extract, and test file) or five arguments (source corpus text file, target corpus text file, word alignments text file, file name for grammar to extract, and test file).
-This program can also be accessed as an Apache Ant task (see figure \ref{ant-file}).
+When run directly, this program accepts either three arguments (a compiled {\tt .josh} directory, file name for grammar to extract, and test file) or five arguments (source corpus text file, target corpus text file, word alignments text file, file name for grammar to extract, and test file).
+%This program can also be accessed as an Apache Ant task (see Figure \ref{ant-file}). 
 The following subsections document the mandatory and optional parameters that can be passed to this program through the {\tt extractRules} ant task.
@@
 
-	\item[maxPhraseSpan] Defines the maximum span (in the source corpus) of any extracted SCFG rule. Default value is ``10".
+	\item[maxPhraseSpan] Defines the maximum span (in the source corpus) of any extracted SCFG rule. Default value is {\tt 10}.
 
-	\item[maxPhraseLength] Defines the maximum number of tokens (terminals plus nonterminals) allowed in the source right-hand side of any extracted SCFG rule. Default value is ``10".
+	\item[maxPhraseLength] Defines the maximum number of tokens (terminals plus nonterminals) allowed in the source right-hand side of any extracted SCFG rule. Default value is {\tt 10}.
 
@@
 Note: the number and type of nonterminals is the same in the source and target right-hand sides of a SCFG rule.
-Default value is ``2".
+Default value is {\tt 2}.
 
@@
 
-	\item[encoding] Defines the file encoding scheme of the input test file and the output grammar file. Default value is ``UTF-8".
+	\item[encoding] Defines the file encoding scheme of the input test file and the output grammar file. Default value is {\tt UTF-8}.
 
 	\item[ruleSampleSize] When extracting SCFG rules for a given source language phrase, this option defines the number of instances of that source phrase will be sampled from the source training corpus for use in rule extraction.
-Default value is ``300".
+Default value is {\tt 300}.
 
@@
 
-The following parameters can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}} .
+The following parameters can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}}.
 
@@
 This option is provided for compatibility with Hiero's suffix array rule extractor \cite{Lopez2008}, in which this setting is set to true.
-Default value is ``true".
+Default value is {\tt true}.
 
@@
 For compatibility with \newcite{Lopez2008}, set this parameter to false.
-Default value is ``true".
+Default value is {\tt true}.
 
@@
 
-Additional options can be configured in the {\tt extractRules} target to affect behavior of the prefix tree.
+Additional options can be configured in the {\tt extractRules} target to change the behavior of the prefix tree.
 
@@
 If set to false, a new prefix tree will be created to process each sentence in the test file.
-Default value is ``true".
+Default value is {\tt true}.
 
@@
 If set to true, the tree will be printed after processing each sentence in the test file.
-Default value is ``false".
+Default value is {\tt false}.
 
@@
 \section*{Acknowledgements}
+
+This research was supported in part by the EuroMatrixPlus project funded by
+the European Commission under the Seventh Framework Programme, by the US National Science Foundation under grant IIS-0713448, and by the DARPA GALE program under contract number HR0011-06-2-0001.
+Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the sponsors.
 
@@
 \end{document}
+
+\begin{figure}[hptb]
+\begin{verbatimtab}[4]
+<project name="Joshua Grammar Extraction in Ant" default="extract">
+
+	<property name="classes.dir" value="bin"/>
+	<property name="experiment.dir" value="/path/to/my/data"/>
+
+	<taskdef name="extractRules" 
+		classname="joshua.prefix_tree.ExtractRules" 
+		classpath="${classes.dir}"/>
+
+	<taskdef name="compileJosh" 
+		classname="joshua.corpus.suffix_array.Compile" 
+		classpath="${classes.dir}"/>
+
+	<target name="compile" 
+		description="Compile aligned parallel corpus
+			 into memory-mapped files">
+
+		<compileJosh
+			sourceCorpus="${experiment.dir}/source.txt"
+			targetCorpus="${experiment.dir}/target.txt"
+			alignments="${experiment.dir}/align.txt"
+			outputDir="${experiment.dir}/model{\tt .josh}"
+		/>
+	</target>
+
+	<target name="extract" 
+		description="Extract a translation grammar 
+			from a compiled parallel corpus">
+
+		<extractRules 
+			joshDir="${experiment.dir}/model{\tt .josh}" 
+			testFile="${experiment.dir}/target-test.txt" 
+			outputFile="${experiment.dir}/test-set-specific.grammar"
+		/>    
+	</target>
+
+</project>
+\end{verbatimtab}
+\caption{Sample Apache Ant build file defining how to extract a translation grammar from an aligned parallel corpus.
+This example assumes that the files {\tt source.txt}, {\tt target.txt}, {\tt align.txt} and {\tt target-test.txt} already exist in the directory specified in the {\tt experiment.dir} variable.
+Running ant with this file will create the directory {\tt model{\tt .josh}} and extract {\tt test-set-specific.grammar}.}
+\label{ant-file}
+\end{figure}

### 2009-12-14 papers/mt_marathon10_SuffixArray/paper.tex
Added caption
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1242 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \end{algorithmic}
-\caption{}
+\caption{During suffix array creation, the contents of a corpus array are sorted using the element comparison function {\sc {Compare\_Elements}}}
 \label{sort}
@@
 It is often useful (especially during MERT) to extract a test set specific grammar once in a preprocessing step, since that test set will be translated many times and re-extracting the grammar each time would be wasteful.
-The {\tt joshua.prefix\_tree.ExtractRules} program performs this task.
+To perform this task, the program {\tt joshua.prefix\_tree.ExtractRules} can be used.
 When run directly, this program accepts either three arguments (a compiled {\tt .josh} directory, file name for grammar to extract, and test file) or five arguments (source corpus text file, target corpus text file, word alignments text file, file name for grammar to extract, and test file).

### 2009-12-27 papers/mt_marathon10_visualization/visualization.tex
initial import of color version. having figure-layout problems.
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1266 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

\documentclass[nologo]{pbml}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{wrapfig}
\usepackage{url}

\title{Visualizing Data Structures in Parsing-based Machine Translation}
%\subtitle{}

\author{firstname=Jonathan, surname=Weese,
       address={Department of Computer Science and Center for Language and Speech Processing \\
Johns Hopkins University \\
Baltimore, MD 21218, USA \\
{\tt jonny@cs.jhu.edu}}}
\author{firstname=Chris, surname=Callison-Burch, 
       address={Department of Computer Science and Center for Language and Speech Processing \\
Johns Hopkins University \\
Baltimore, MD 21218, USA \\
{\tt ccb@cs.jhu.edu}}}

\shorttitle{Visualizing Machine Translation}
\shortauthor{J. Weese, C. Callison-Burch}

\begin{document}
\maketitle
\begin{abstract}

As machine translation (MT) systems grow more complex and incorporate more
linguistic knowledge, it becomes more difficult to evaluate independent
pieces of the MT pipeline.
Being able to inspect many of the intermediate data structures used during
MT decoding allows a more fine-grained evaluation of MT performance, helping
to determine which parts of the current process are effective and which are
not.
In this article, we present an overview of the visualization tools that are
currently distributed with the Joshua \cite{Joshua-WMT} MT decoder.
We explain their use and present an example of how visually 
inspecting the decoder's data structures has led to useful improvements in
the MT model.

\end{abstract}

%{\sc
%Here are some citations to work into the text
%\begin{itemize}
%\item Syntax Augmented Machine Translation \cite{samt2006,zoll:08,Venugopal2009}
%\item Hiero \cite{Chiang2005,Chiang2007}
%\item Hypergraph citations \cite{billott-lang:1989:ACL,Klein2001}
%\end{itemize}
%}

\section{Introduction}

The Joshua machine translation decoder uses a 
formalism known as synchronous context-free grammars (SCFGs) \cite{Chiang2006}.
A probabilistic SCFG consists of a set of source-language terminal symbols, a set of target-language terminal symbols, a set of nonterminals
that is shared between both languages, and set of production rules of the
form
$$X \to \langle \gamma,\alpha,\sim,w \rangle$$
where $X$ is a nonterminal symbol, $\gamma$ is a (possibly mixed) sequence of
nonterminals and source terminals, $\alpha$ is a (again possibly mixed)
sequence of nonterminals and target-language terminals, $\sim$ is a one-to-one
correspondence between the nonterminals of $\gamma$ and $\alpha$, and $w$ is
a weight for the production rule.
%where $X \in N, \gamma \in [N \cup T_S]*$ is a (possibly mixed) sequence of
%nonterminals and source terminals, $\alpha \in [N \cup T_T]*$ is a 
%(again possibly mixed) sequence of nonterminals and target-language terminals,
% and $\sim$ is a one-to-one correspondence between the nonterminals of $\gamma$ and $\alpha$.
%$w$ is a weight for the production rule.

Using an SCFG to parse the input sentence automatically creates a corresponding target-language sentence.
We take the generated sentence as a candidate translation for the
input sentence. The sequence of rules used to generate $t_1^m$ is
called a {\it derivation}. The derivation encodes synchronous parse trees on
the source and target sides.

%As an example, consider the grammar with production rules
%\begin{align*}
%\textrm{PP} & \to \langle \textrm{ PREP}_1 \textrm{ NP}_2, \textrm{ PREP}_1 \textrm{ NP}_2 \rangle \\
%\textrm{NP} & \to \langle \textrm{ DET}_1 \textrm{ NN}_2, \textrm{ DET}_1 \textrm{ NN}_2 \rangle \\
%\textrm{PREP} & \to \langle \textit{ sur}, \textit{ on } \rangle \\
%\textrm{DET} & \to \langle \textit{ la}, \textit{ the } \rangle \\
%\textrm{NN} & \to \langle \textit{ lune}, \textit{ moon } \rangle
%\end{align*}
%The subscripts on nonterminals on the right hand side of a production rule
%represent the one-to-one correspondence $\sim$. For example, in the first rule,
%the PREP nonterminals are aligned, and the NP nonterminals are aligned.
%Figure \ref{toy-tree-figure} shows a parse tree for the French phrase
%``{\it sur la lune}'' and its automatically-generated candidate translation
%``on the moon.''

%In practice these grammars are automatically extracted and weighted using
%a bilingual parallel corpus. There are typically thousands of rules that may
%apply for a given input sentence and thousands of possible derivations and
%candidate translations.

Since we are parsing the input sentence with a probabilistic grammar, we can
generate many possible candidate parses for a particular sentence. These parses
may share a lot of structure --- for example, two different parse trees may
contain subtrees that are identical. 
As part of the search procdeure, the Joshua system compactly stores 
these structure-sharing
trees as a {\em hypergraph} or {\em packed forest} where identical subtrees
from many parses are represented by a single copy.
Structure-sharing hypergraphs are commonly used to represent the result of
parsing a sentence with a probabilistic CFG \cite{billott-lang:1989:ACL,Klein2001}.
So it is natural that Joshua's
SCFG-based model should use this data structure.

\begin{figure}[t]
\begin{center}
\includegraphics[height=3in]{images/hypergraph-example}
\end{center}
\caption{A hypergraph showing two candidate translations of {\it Je suis mon
ma\^{i}tre}.}
\label{hypergraph-example-figure}
\end{figure}

%\begin{wrapfigure}{r}{.66\linewidth}
%\begin{center}
%\includegraphics[width=\linewidth]{images/hypergraph-example}
%\end{center}
%\caption{A hypergraph showing two candidate translations of {\it Je suis mon ma\^{i}tre}.}
%\label{hypergraph-example-figure}
%\end{wrapfigure}    
%
Figure \ref{hypergraph-example-figure} shows an example of a hypergraph with
two candidate translations for the French sentence ``{\it Je suis mon
ma\^{i}tre}.'' Note that both translations --- ``I am my master'' and ``I
follow my master'' --- use the same derivation of ``my master'' as a phrasal
translation for the French {\it mon ma\^{i}tre}. Therefore, the hypergraph
maintains only one copy of this shared structure, saving space.

%The hypergraph is a representation of each of the choices
%made by the decoder during the decoding process:
%parsing with a SCFG is just choosing which production rules to use
%and in which order they should be applied.
%Each node of the hypergraph then represents a ``choice point'' that affects
%which candidate translation is evetually produced. In general, exponentially
%many derivations can be enumerated from the hypergraph by selecting different
%production rules at each choice point.

We focus on two data structures
used in the decoding process: first, each candidate translation has a {\em
derivation tree} describing how the decoder generated the candidate. Second,
there is a {\em hypergraph} that represents {\em all} the choices made by the
decoder when translating a single input sentence and shows us the relationships among all the various candidates.
%The visualization tools included with Joshua let the researcher inspect both
%of these data structures.

\section{Joshua's Visualization Tools}

The visualization tools are included in the Joshua Decoder release.\footnote{\url{http://cs.jhu.edu/~ccb/joshua/}} They have
only one outside dependency: the Java Universal Network/Graph Framework\footnote{\url{http://jung.sourceforge.net/}} (JUNG), a toolkit for drawing graphs in Java,
which is available under the BSD License.

The derivation-tree viewer does not depend on the Joshua decoder. It can
visualize any derivation tree as long as the tree representation is identical
to Joshua's output. The hypergraph visualizer, on the other hand, depends
on the decoding process itself, and therefore on the Joshua decoder.

\subsection{Synchronous Derivation Trees}

\begin{figure}[t]
\includegraphics[width=.95\linewidth]{images/pick-window-2} 
\includegraphics[width=\linewidth]{images/urdu-tree-window}
\caption{The Derivation Tree browser's sentence selection and tree-viewing windows.}
\label{derivation-figure}
\end{figure}

The Joshua decoder includes an option to output text-based representations
of the target-side parse tree rather than the plain candidate translation.
These textual representations can also be annotated with the source-side
span of each nonterminal in the tree. 
(These outputs can be chosen by setting {\tt use\_tree\_nbest} and {\tt 
include\_align\_index} to true in the Joshua configuration
file.)
The core of the derivation-tree visualizer
 takes a string that represents an annotated parse tree and uses JUNG to draw
a graph representing the tree.

Using the source-side annotations, the visualizer also draws the 
parse tree for the input sentence. Since the grammar is synchronous, there is
a one-to-one correspondence between nonterminals in the source- and target-side
trees. Any differences in the tree structure arise from possible reordering
of the nonterminals.
The bottom of Figure \ref{derivation-figure} shows part of the derivation tree 
generated by Joshua's output of 
% ccb - I manually inserted some newlines to make this wrap properly. 
{\tt (ROOT\{0-24\} ([GOAL]\{0-23\} ([GOAL]\{0-6\} ([NP]\{0-6\} ([NP]\{4-6\} this meeting) in ([NNP-GPE]\{0-1\} jerusalem))) ([S]\{6-23\} ([VP/NP]\{14-22\} is 
 \\
 ([VBG]\{19-21\} being) considered as ([NP+IN]\{14-18\} a ([NN]\{17-18\} show) of ([NN]\{15-16\} support) for)) ([NP]\{6-14\} palestinian ([JJ-GPE-ite\textbackslash NP]
 \\
\{7-14\} president ([NP-PERSON]\{8-10\} mahmoud abbas) ([PP]\{10-14\} for 
 \\
 ([NNP-GPE]\{12-13\} israel)))) .)))}

The tree visualizer takes the following arguments:  a file containing source-side sentences (one per line), a file containing the parallel target-side reference translations, and then one or more files containing n-best translations of the source sentences.
Multiple n-best files can be specified in order to contrast different runs
of the decoder on the same test set. For example, Figure \ref{example-derivations} shows one
sentence from the 2009 NIST Urdu--English evaluation, decoded once under a 
Hiero-style grammar with only one nonterminal symbol X, as introduced in
\cite{Chiang2005}, and once under a
syntactically-motivated grammar with a richer nonterminal set as  presented in \cite{samt2006}.

\begin{figure}[t]
\begin{center}
\includegraphics[height=2.75in]{images/hiero-tree}
\includegraphics[height=2.75in]{images/samt-tree}
\end{center}
\caption{An example visualization of two derivation trees for SCFGs that use a Hiero-style grammar and a syntactically-motivated grammar.}\label{example-derivations}
\end{figure}

When the visualizer starts up, it creates two types of windows. First there is
a window listing all the reference translations.
This window is shown on the top of Figure \ref{derivation-figure}.
(This is the reason for the
reference translation file --- an earlier version of the program had the 
sentences listed by the source side, but this was less useful for users who
had no knowledge of the source language.) The second type of window displays
a derivation tree.

The browser creates one tree-displaying window 
(as seen on the bottom of Figure \ref{derivation-figure})
for every n-best file that is
passed as an argument. This allows the user to easily compare derivation trees
created by different grammars (these are saved in different n-best files,
since they're the result of different runs of the decoder). In a tree window,
the user can click and drag to inspect different parts of the tree and can use
the scroll wheel to zoom in and out.

There are two ways to change trees: clicking the sentence in
the list window or using the left- and right-arrow buttons in the tree windows.
In either case, all the tree windows are synchronized; that is, whenever the
user changes to a different sentence, all tree windows are updated to show
the derivation for that sentence. The tree views are anchored; for example, if
a user were zoomed in on the first target-side nonterminals of one tree, when
he changed to a new sentence and the new tree was displayed, its view would
start zoomed in on that same area of the new sentence's tree.

The tree view windows also have a button to give the user more information about the current tree.  This button shows the source sentence, the reference translation, and the text of the translation candidate.

\subsection{Hypergraphs}

\begin{figure}[t]
%\includegraphics[width=.95\linewidth]{images/bw/hg-pick-window-2-bw}
\includegraphics[width=\linewidth]{images/hamid-ansari-derivations-cropped}
\caption{The visualization window for the hypergraph browser.}
\label{hypergraph-screenshot}
\end{figure}

The derivation tree viewer is used after running the decoder to inspect its
ouput. Since the hypergraph is used during the decoding process and is
discarded afterwards, the hypergraph visualizer depends on the decoder itself.
Internally, it works by setting a flag to build the graph visualization 
on-the-fly as the decoder processes the input.

The command line arguments for the hypergraph visualizer are slightly
different from the derivation tree viewer. A reference translation
is used in the same way as in the derivation tree viewer --- it allows the user
to choose a sentence to translate based on the reference translation. Now
because the decoder translates source sentences on demand, the other two
arguments are a source-sentence file and a Joshua configuration file that
should be used for the decoding.

On startup, the hypergraph viewer presents a list of sentences to translate.
%This window is shown on the top of Figure \ref{hypergraph-screenshot}.
The reference translations are displayed instead of the source to make it easier for the user. The user selects a sentence from the list and presses the ``Decode'' button at the bottom. Internally, the hypergraph viewer chooses the associated source sentence and calls Joshua to decode the sentence using the supplied
configuration file.

As the hypergraph structure is constructed, the
hypergraph viewer uses JUNG to build a corresponding graph. At first glance,
the graph that is displayed looks very similar to a derivation tree. In fact,
it is a hypergraph representation of the one-best candidate for the source
sentence.
Recall that the nodes of the hypergraph represent nonterminals or terminals in
the derivation of a given source sentence, and the hyperedges represent
production rules that are applied at each step.
Each node of the hypergraph is a rectangular box in this visualization, and
each hyperedge is a small circle. There may be many possible hyperedges leading
from a particular node, each one representing a different rule that could
be applied to that node. When only one hyperedge is visualized at each node,
the resulting graph represents one candidate translation.

Using the hypergraph viewer, we can inspect the choices that the decoder made
at each point. When the user clicks on a node of the hypergraph, a list of all
hyperedges leading from that node is displayed on the left. If the user selects
a hyperedge from the list, the corresponding subtree is displayed in the graph.
A user can even select multiple rules, and all of the resulting subtrees are
displayed side-by-side.
%The bottom of
Figure \ref{hypergraph-screenshot} shows this hypergraph view.
% On the list, two rules are selected, and the two corresponding subtrees,
% showing two distinct derivations of the phrase ``he said''  are generated
%in distinct colors below their parent node.
We can see four subtrees giving possible derivations of the 
candidate translation
``hamid ansari nominated for vice president.''

\section{Other Visualization Tools}

This section describes two other visualization tools developed by other
researchers, and contrasts the goals of the various visualization systems and
how the goals are reflected in design choices.

\subsection{The Chinese Room}

The Chinese Room \cite{albrecht-hwa-marai:2009:EACL} is a collaborative translation interface. It uses visualization
techniques to allow a user who has no knowledge of the source-side language to
collaborate with an MT system to create good translations. This is different
from earlier approaches to collaborative MT where the user was often assumed to
 be a professional translator.

%The most important difference between the Chinese Room and Joshua's
%visualization tools is that they are designed for different goals.
The Chinese
Room is designed to allow {\em translators} (even those with limited or no
knowledge of the source language) to produce good translations of the input
sentences. Joshua's visualization tools, on the other hand, are designed to
help {\em researchers} debug grammars and improve their translation models.
%Naturally, observing a user's interaction with the Chinese Room may lead to
%improvements in a translation model, but the user's goal is, in the end, to
%decode the input sentence given.
%So naturally there are differences in the two visualization schemes. 
The Chinese Room tries to give the user as much information as possible to
create a correct translation. This includes word alignments, glosses for
source words from a dictionary, source-side parse structure and so on. All
of this information would be useful for a translator.
%On the other hand, 
The current tools for Joshua are focused on improving
the grammars used in a translation model. We need comparatively less 
information to gain useful insight into this smaller domain. That is why we
have focused only on displaying the derivation trees and hypergraphs.
%We are interested in analyzing and correcting the mistakes that the decoder
%drew from the grammar, rather than helping a human to avoid those mistakes
%during the translation process.

\subsection{DerivTool}

DerivTool is a tool for interactively directing the decoding of a sentence
using a syntax-based MT model \cite{deneefe-knight-chan:2005:PosterDemo}. Users can choose which
rules to apply to a derivation at which time. In the end, the user ends up
building up a derivation tree for a particular candidate translation. DerivTool
is useful for analyzing grammars: a user can immediately tell when a needed
rule is missing (they can't continue their intended derivation) and can see
which rules are favored by the decoder at a particular point (since rules are
displayed in order of frequency).

%Since Joshua also visualizes derivation trees, we can consider DerivTool and
%Joshua as two tools that produce the same result (that is, they are visualizing
%the same data structure in the end). The main difference is whether the
% process is interactive or not. Both interactivity and batch-mode usage
% have their advantages.

Joshua's tools and DerivTool both produce derivation trees. The difference is that DerivTool is interactive and Joshua is not.
The main advantage of working in batch mode is that it is faster than building
up a derivation tree manually step-by-step. We run the decoder seperately,
letting it make the translation decisions. Afterwards,
the user can evaluate the quality of the trees produced. The Joshua
visualization tool can produce many derivation trees all at once (it reads
in a file containing the n-best derivations for each sentence of a test set).
This makes it easy to compare the different decisions that were possible at
decoding time without the user having to manually make the decisions himself.

\section{How Joshua's Visualization Tools Help}
\begin{figure}[t]
\begin{center}
\includegraphics[height=3in]{images/terminals-being-eaten-highlights}
\end{center}
\caption{An example of bad production rules that parse pieces of the source
sentence without producing any target-side output.}
\label{terminals-being-eaten}
\end{figure}

By visualizing derivation trees for different candidate translations, we
can find problems with the SCFGs that underly translations.
For example, we used the visualization tool to inspect derivation trees that were produced on the 2009 NIST
Urdu--English test set,
and noticed that many rules consumed part of the input sentence without producing any
output.  These rules were used often in the top candidates, but brought
down the translation quality. Having discovered that, we manually removed %rules of the form
%$$\textrm{X} \to \langle \gamma, \alpha, \sim, w\rangle$$
%where $\alpha$ is a string of nonterminals with no target-side terminal
%symbols. This helped improve the translation quality.
where the target side contained no terminal symbols.
This helped improve the translation quality.
In Figure \ref{terminals-being-eaten}, we can see rules of the form
$\textrm{[IN]} \to \langle \textrm{[IN]}\gamma,\textrm{[IN]}\rangle$ and
$\textrm{[.]} \to \langle \textrm{[.]}\gamma, \textrm{[.]}\rangle$ being
applied.

Pruning grammars of systematically bad production rules is a good way to
improve translation quality, and manual inspection of the output to see which
grammar rules have been used, and which ones correlate with low translation
quality, is an effective way to prune.
The derivation tree visualizer helps researchers too notice patterns in rule
application among different candidate translations.

Visualizing hypergraphs lets the researcher inspect the decisions that the
decoder made when choosing among different rules that could be applied at some
point in a derivation. This could be useful, for example, in determining if
a particular type of rule is being systematically overweighted or underweighted,

Visualizing the data structures involved in MT decoding allows the researcher
to determine empirical rules for improving the grammars involved.

%First, we can tell at a glance whether the parse tree seems ``reasonable.''
%That is, the intuition behind the SCFG formalism suggests that parse trees
%should have the right shape on both the source and target side. A single
%phrase should be the output of a single subtree; two phrases should be
%combined in a particular way, and so on. A grammar that produces excessively
%branching trees, for example, can be noticed immediately by looking at the
%tree itself.

%Secondly, we can do a more fine-grained analysis of the candidate. Maybe some
%parts of the translation look reasonable, but others are way off. By examining
%the parse tree, we can tell what rules of the grammar give rise to bad
%translations by seeing which subtrees correspond to the bad translation.
%
%By examining candidate translations with similar errors, we can see what kinds
%of grammars correspond to good translations and what kinds of grammars cause
%errors.

%Visualizing the hypergraph gives us another useful perspective on the decoding
%process. Before, we were comparing the parse trees of individual candidates.
%This let us get a sense of which rules of the SCFG correspond to good
%translations and which ones do not. But by visualizing the hypergraph, we
%can see which rules are competing with each other during the decoding process
%itself.
%

\section{Future Work}

There are still many improvements that should be made to these visualization
tools. %The first is that
We would like to be able to show the terminal alignments
that are induced by a particular derivation. But that requires more information
than is currently given by the Joshua decoder.
It only annotates nonterminals with
source-side spans.
As an example, consider the French phrase {\it l'objectif de $X$ est de}
 which corresponds to the English phrase {\it the goal of $X$ is to}. We know
 that those two phrases have corresponding spans, and that the two $X$ symbols
 also correspond. But despite this, we don't have word-level alignment
information: we don't know if {\it l'objectif de} corresponds to {\it the goal
of} or to {\it is to}. Such fine-grained information would be useful for
visualizing reordering in MT models.

Another improvement that we think would greatly increase the utility of these
visualization tools for research is to add support for exporting the displayed
trees as files. It would be nice to be able to  save an interesting tree in PDF format so that it could be easily embedded in a research paper.

There are other parts of the Joshua pipeline that might benefit
from visualization. During rule extraction, SCFG rules are automatically
generated given an aligned parallel corpus. Being able to visually inspect
both the aligner output and the results of the rule extraction on an individual
phrase could provide some insight into this process.
The decoder works by CKY
parsing, so it would be advantageous for researchers to be able to view the
parse chart that is produced --- which constituents are generates where, what
their associated weights are, which ones have been pruned, and so on. This
 would help to determine if translation errors are caused by search errors
 or something else.

%As noted above, the derivation-tree viewer is static. It reads in a
%representation of a parse tree and renders it. It would be easier for the user
%to interpret a particular tree if the derivation itself were presented as an
%animation. The user could immediately see the results of a single rule being
%applied, as well as the effect of any re-ordering on the parallel parse tree.

\section*{Acknowledgments}
This research was supported in part by the EuroMatrixPlus project funded by
the European Commission under the Seventh Framework Programme, and by the US National Science Foundation under grant IIS-0713448.  The views and findings are the authors' alone.

\bibliography{visualization}
\end{document}

### 2010-01-02 papers/mt_marathon10_xml_transliteration/mt_mar_translit.tex
Added the transliteration paper to the publications branch
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1279 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

\documentclass[logo]{pbml}

\usepackage{clrscode}

\begin{document}

\title{Integrating Output from Specialized Modules in Machine Translation}

\subtitle{Transliterations in Joshua}

\author{firstname=Ann, surname=Irvine,
       address={Center for Language and Speech Processing, Johns Hopkins University}}
\author{firstname=Mike, surname=Kayser, address={BBN Technologies}}
\author{firstname=Zhifei, surname=Li,
        address={Center for Language and Speech Processing, Johns Hopkins University}}
\author{firstname=Wren, surname=Thornton,
        address={Cognitive Science Program, Indiana University}}
\author{firstname=Chris, surname=Callison-Burch,
        address={Center for Language and Speech Processing, Johns Hopkins University}}

%\shorttitle{Integrating Output from Specialized Modules in Machine Translation}
\shorttitle{Specialized Modules in MT}
%\shortauthor{A. Irvine, M. Kayser, Z. Li, W. Thornton, C. Callison-Burch}
\shortauthor{Irvine, Kayser, Li, Thornton, Callison-Burch}

\maketitle

\begin{abstract}
In many cases in SMT we want to allow specialized modules to propose translation fragments to the decoder and allow them to compete with translations contained in the phrase table. Transliteration is one module that may produce such specialized output. In this paper, as an example, we build a specialized Urdu transliteration module and integrate its output into an Urdu--English MT system. The module marks-up the test text using an XML format, and the decoder allows alternate translations (transliterations) to compete.
\end{abstract}

\section{Introduction}
The phrase tables used in statistical machine translation (SMT) systems are often incomplete, and they may not take full advantage of the linguistic knowledge that we have about a language. However, many data-driven NLP tools exist for specific linguistic tasks. For this reason, it is often useful to create specialized modules that employ other methods of translation not so dependent on the training text. Such modules may include noun phrase taggers and translators \cite{Koehn2004a}, morphological analyzers, modality taggers and translators,\footnote{Verbal modality expresses the notions of possibility, necessity, permission, and obligation} and transliteration systems. These modules may then be integrated into the MT pipeline \cite{Dugast2007, Yang2006}.

Previous SMT systems have integrated the subtask output of specialized modules using an XML-markup on input text \cite{Koehn2004b, Senellart2003}. Here we present an XML format to integrate the output of our specialized transliteration module into the Joshua decoder \cite{JoshuaWMT}. It necessarily differs from the XML schemes used in phrase-based decoders because Joshua is a parsing-based decoder. We illustrate its use with an example transliteration module.

\section{Decoding Constraints}
Joshua \cite{JoshuaWMT} is an open source\footnote{http://cs.jhu.edu/$\sim$ccb/joshua/} SMT system that uses synchronous context free grammars (SCFGs) as its underlying formalism. SCFGs provide a convenient and theoretically grounded way of incorporating linguistic information into statistical models of translation. Joshua implements all the essential algorithms described in \cite{Chiang2007} and supports Hiero-style rules \cite{Chiang2005} as well as richer syntax augmented rules \cite{Zollman2008}. The version of Joshua that we have used in this work incorporates the grammar extraction software that comes as part of their open source SAMT toolkit.\footnote{http://www.cs.cmu.edu/zollmann/samt/} Joshua translates by applying the extracted SCFG rules to the source language text using a general chart parsing framework \cite{PBML09}.

A probabilistic SCFG consists of a set of source-language terminal symbols $T_S$, a set of target-language terminal symbols $T_T$, a set of nonterminals $N$ that is shared between both languages, and set of production rules of the form
$$X \to \langle \gamma,\alpha,\sim,w \rangle$$
where $X \in N, \gamma \in [N \cup T_S]*$ is a (possibly mixed) sequence of nonterminals and source terminals that form the lefthand side of the rule, $\alpha \in [N \cup T_T]*$ is a  (again possibly mixed) sequence of nonterminals and target-language terminals that form the righthand side of the rule, and $\sim$ is a one-to-one correspondence between the nonterminals of $\gamma$ and $\alpha$. $w$ is a weight for the production rule.

To support integrating specialized modules we introduced the ability to specify alternate translation rules in the document to be translated. In order to support both the use of alternate translation rules and regular decoding (without alternate rules), we introduced a new parameter in Joshua's configuration file for specifying which parser to use on the input file. Each input file parser reads in the file and emits a sequence of segments to be translated. In order to avoid storing the whole file in memory, the sequence is returned as a co-iterator.\footnote{We use some object instance which implements the \texttt{joshua.util.CoIterator} interface. Both iterators and co-iterators are examples of abstractions over enumerations. Whereas an iterator captures the notion of producing the elements, a co-iterator captures the notion of consuming those elements.} By using a co-iterator, the sequence is produced lazily and consumed on-line, invoking the chart parser as a co-routine. To avoid interrupting translation in the middle of a file, we want to detect malformed files before translation begins. So in order to detect malformed file errors eagerly, the file is read once with a co-iterator that consumes the segments but does nothing with them, and then re-read to produce segments for the chart parser.

Each translation segment consists of an ID, a source sentence, and a collection of spans covering the sentence where each span contains a collection of constraints. Spans containing only soft constraints are allowed to overlap, whereas hard constraint spans may not overlap. Constraints are drawn from three types: lefthand side (LHS) constraints, righthand side (RHS) constraints, and rule constraints. LHS constraints are hard constraints specifying that the span be treated as a specified nonterminal, thus filtering the regular grammar to generate translations only from that nonterminal. One use for LHS constraints is to integrate chunking or tagging information before decoding. RHS constraints are hard constraints filtering the regular grammar such that only rules generating the desired translation can be used. A use for RHS constraints would be integrating word sense disambiguation before decoding. Rule constraints specify a new grammar rule including a LHS nonterminal, source RHS (derived from the source sentence), target RHS, and feature values. Rule constraints can be either hard or soft; if they are hard, they override the regular grammar; if they are soft, they are considered an addition to the regular grammar and will compete with regular rules. Rule constraints for any given span must be marked collectively as all hard or all soft.

\subsection{XML Markup}
The XML format follows straightforwardly from the specification of segments. The XML file must be valid XML, and thus must have a root element. Underneath the root element are some number of \texttt{<seg>} tags with a required \texttt{id} attribute that specifies the input segment number. The source sentence is given as raw text under the \texttt{<seg>} tag. Each \texttt{<seg>} tag may contain one or more \texttt{<span>} tags with required \texttt{start} and \texttt{end} attributes and an optional \texttt{hard} attribute for rule constraints. Each \texttt{<span>} tag must contain one or more \texttt{<constraint>} tags each of which contains an optional \texttt{<lhs>} tag, containing nonterminal text, followed by an optional \texttt{<rhs>} tag with an optional \texttt{features} attribute and containing target text. Any other tags are ignored by the XML parser.

This specification of the XML format is overly liberal and could admit files which are non-sensical or which cannot be represented internally. In order to rule out such files, the generated objects are run through a type checker to ensure semantic validity. The type checker verifies the following invariants:

\begin{itemize}
\item Each constraint adheres to one of the three types, thus it has
	\begin{itemize}
	\item only a \texttt{<lhs>} tag, or
	\item only a \texttt{<rhs>} tag with no \texttt{features} attribute, or
	\item a \texttt{<lhs>} tag, a \texttt{<rhs>}, and a \texttt{features} attribute
	\end{itemize}
\item For each span,
	\begin{itemize}
	\item the start and end indices are within the width of the sentence
	\item the start index is smaller than the end index
	\end{itemize}
\item There are no overlapping hard spans
\end{itemize}

While the \texttt{features} attribute is considered optional in terms of the DTD for the XML grammar, that is only because DTDs are unable to capture the dependency relation between the three valid constraint types.\footnote{XML Schema and RelaxNG are also unable to capture all the necessary dependency relations. Even if they can capture the tag and attribute dependencies between the three constraint types, type checking would still be necessary due to the context-sensitive restrictions on valid start and end indices depending on the length in words of the source text.} If both \texttt{<lhs>} and \texttt{<rhs>} tags are provided but there is no \texttt{feature} attribute, this is considered a type error since the constraint does not belong to any of the three types: LHS constraints do not have \texttt{<rhs>} tags, RHS constraints do not have \texttt{<lhs>} tags, and rule constraints require a \texttt{features} attribute.

\section{Decoder Integration}

To enforce the three kinds of constraints (i.e., rule, LHS, and RHS) during decoding, we modified the regular chart-based decoding algorithm in Joshua.
Rule constraints can be hard or soft.
A rule constraint provides a new translation option for a source span,
in addition to those translation options (hereafter called grammar-based translations) provided by the regular grammars.
If the rule constraint is hard, all the grammar-based translations
will be disallowed in the final translation output. Otherwise, the new translation option will
compete with those grammar-based translations, in a probabilistic manner.
Different from a rule constraint, LHS and RHS constraints are always hard, meaning that
a grammar-based translation will be disallowed if its LHS or RHS does not match the LHS or RHS constraint.

Figure \ref{alg-constraint-decoding} presents the
modified algorithm, where lines \ref{line1}, \ref{line2}, and \ref{line3} are added to
the regular chart-parsing algorithm in order to support manual constraints.
As shown in line \ref{line1}, the algorithm first adds the rule constraints (regardless of being soft or hard) into the chart
so that the decoding algorithm will consider these rule translations as candidate translations.
To support the \textbf{hard} constraints, the algorithm will run through two filtering processes.
In line \ref{line2}, if a span is within the coverage of a \textbf{hard} rule constraint,
all the grammar-based rules applicable to this span will be disallowed.
Similarly, all the applicable grammar rules that do not match
any LHS or RHS constraints for the span will be filtered out, as shown in line \ref{line3}.

\begin{figure}[t]
\begin{center}
\begin{codebox}
\Procname{$\proc{\bf DecodingWithConstraints}(\id{grammars}, \id{ sentence}, \id{ constraints})$}
\li Add the \textrm{rule} constraints into the Chart \label{line1}
%\li regular chart seeding%
%\li $\For \id{w}  \kw{in }  [1,n]$
%\li \> $\For \id{i}  \kw{in }  [0,n-w]$
%\li \> $j = i + w$ \Comment each span [i,j]
\li For each \textrm{span} $[i,j]$ with increasing length (i.e., $j-i+1$)
\li \> Identify applicable grammar rules for the span
\li \> Filter the grammar rules based on \textbf{hard} \textrm{rule} constraints \label{line2}
\li \> Filter the grammar rules based on \textrm{LHS} and \textrm{RHS} constraints \label{line3}
\li \> Add the surviving grammar rules into the chart
\end{codebox}
\end{center}
\caption{Constraining Chart-based Decoding with Manual Constraints.}
\label{alg-constraint-decoding}
\vspace{-0.1in}
\end{figure}

\section{Transliteration Module}
Here we present a specialized transliteration module that uses Joshua's new XML markup. We developed an Urdu-English transliterator, which is useful because our LDC Urdu Language Pack bilingual parallel corpus \footnote{LDC catalog number LDC2009E12, "NIST Open MT08 Urdu Resources"} has only 88,108 sentence pairs with 1,586,065 English tokens and 1,664,409 Urdu tokens. In this data, we observed that 2\% of words in a development set were out-of-vocabulary (OOV) with respect to the training bitext. With the help of a human annotator, we found that approximately 33\% of these words were phonetically transliterable; for example, proper names or borrowed words. Introducing a module for generating transliterations and integrating that output into the output of an end-to-end MT system clearly has the potential to improve performance.

\subsection{Basic Framework} \label{transliteration-framework}
We treat transliteration as a \emph{monotone character translation} task, similar to the work of \cite{Knight1997}. We used the Joshua MT system to build an Urdu transliteration module using a semantically-informed framework, described below, and several sources of transliteration pairs for training.

At training time, given a list of Urdu--English name pairs, we first perform character-to-character alignment using the freely available Berkeley Word Aligner.\footnote{http://code.google.com/p/berkeleyaligner/} Next, we find character-sequence pairs which conform to the alignment graph for a word pair; these are analogous to phrase pairs in phrase-based statistical MT. We build a table of such character-sequence mappings, annotated with translation probabilities. Finally we extract a frequency-annotated list of 1.3 million names from the English Gigaword corpus using a named entity tagger \cite{Finkel2009}. We then use this to train a character language model prior. Having trained these components, we use them in conjunction with the off-the-shelf Joshua MT decoder.

During decoding, a novel Urdu word is segmented into sequences of characters, and each character sequence is translated to an English character sequence. Unlike in the closely analogous process of phrasal machine translation, in phonetic transliteration the translated character sequences are never reordered. Transliteration hypotheses are scored using a log-linear model which makes use of character sequence translation scores and a character-LM prior score. The result is a single English phonetic gloss of the Urdu word.

\subsection{Semantically-targeted transliteration} \label{semantically-targeted-transliteration}
We trained two transliteration systems, one for person names and one for all other semantic types (including non-names). These two systems shared all components except for the character LM and the dataset used for decoder weight tuning. For the person-name transliteration system, we trained our character language model from the large list of English person names automatically extracted from the Gigaword corpus. In the non-person-name transliteration system, we trained a model from a large list of English words, without regard to semantic type.

\subsection{Training the module}
In order to train the transliteration module, we gathered pairs of names that were likely to be transliterations of one another. We obtained unique name pairs from three sources: the Urdu--English parallel corpus (about 2,000 pairs, extracted from alignments and checked by a human annotator), Amazon's Mechanical Turk system (over 12,000 pairs), and linked people pages from the Urdu and English Wikipedias (about 1,000 pairs). Our final system used about 15,000 pairs of Urdu--English transliterated name pairs.\footnote{The complete name pair list is freely available at http://www.clsp.jhu.edu/$\sim$anni/} Transliteration performance improved with increasing amounts of training data, and our final module outperformed the baseline system available from the LDC Urdu Language Pack.

\subsection{Integrating the module}
When using the rule constraint mechanism to add externally constructed theories to Joshua's search space, it is necessary to specify a left-hand-side nonterminal for each rule. Since this nonterminal label is used by Joshua in decoding, it is helpful to choose a label which accurately represents the grammatical content of the covered phrase. Rather than compute a single guess for this label, we add multiple arcs to the search space, each with a different nonterminal. We use the automatically determined named-entity-category of the source word to construct the set of possible labels for a transliterated word.

In particular, before decoding we compute from training the N most frequent target nonterminal labels assigned to low-frequency words of each source name category (the name categories are: Person, Location, Organization, Geopolitical Entity, Facility). At decoding time, we use an automatic name tagger to compute the semantic category of each source word. For every transliteration candidate, we add N arcs to the search space, each with a different label from the semantically-targeted set. In our experiments, we use the value N=5.

An example of the XML markup for an input Urdu sentence is shown in Figure \ref{example-urdu-xml}. In this example, two word spans are tagged with constraints and hypothesis transliterations. Each span and transliteration hypothesis is tagged with the N most frequent target nonterminal labels (e.g. NNP, NN, etc. for a PERSON tag).

\begin{figure}[t]
\begin{center}
\includegraphics[height=3.5 in]{SampleSentence.pdf}
\end{center}
\caption{An example of XML markup on Urdu text.
Each span has 21 associated feature weights. The first 13 features represent the feature space of traditionally extracted Joshua translation rules. Since transliteration rules do not exist in the same feature space, these feature values are always set to 0 for transliteration rules.}
\label{example-urdu-xml}
\end{figure}

\section{Results}
We tested the impact of transliterator integration in a small number of blind submissions to the NIST MT09 Urdu--English evaluation. We integrated the transliterator into the current Joshua system. In one experiment, we transliterated all low-frequency words, while in a second experiment we transliterated only low-frequency person names. Our baseline for comparison was the identical Joshua system without transliterations.

We compared the baseline against the transliteration-aware systems both quantitatively and qualitatively. In a quantitative comparison, whose results are in Table \ref{transliteration-impact-quantitative}, transliteration yielded a small but notable BLEU improvement. As shown in the table, transliterating words of all semantic types yielded slightly better performance than transliterating only words marked as person names.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|}
\hline
\textbf{Joshua Translation System} & \textbf{NIST MT09 BLEU Score} \\ \hline
No Transliteration & .2958 \\ \hline
Transliterate names only & .2980 \\ \hline
Transliterate all types & .3010 \\ \hline
\end{tabular}
\end{center}
\caption{\label{transliteration-impact-quantitative} Impact of transliteration on BLEU in submissions to NIST MT09 evaluation. }
\end{table}

We also qualitatively compared the best transliteration-aware system with the baseline system via manual inspection of decoder output. As expected, some sentences showed clear improvement via the increased lexical coverage allowed by the transliteration model, while other sentences showed little benefit. In some sentences, the transliteration model hypothesized incorrect transliterations for OOV words. More effectively filtering such incorrect translation options, such as through a more developed measure of confidence,  is a potential avenue for future work. Tables \ref{transliteration-impact-example1} and \ref{transliteration-impact-example2} show examples of Joshua decoder output with and without the transliteration feature.

\begin{table}
\begin{center}
\begin{tabular}{|c|p{8cm}|}
\hline
\small \textbf{Without Transliteration} & \small \textbf{[UNKNOWN]} Members said that the economic plan, expensive, and will not be effective. \\ \hline
\small \textbf{With Transliteration} & \small \textbf{Republican} members said that the economic plan, expensive, and will not be effective.  \\ \hline
\small \textbf{Reference} & \small The republican members said that this economic plan is very ``expensive'' and will not be effective. \\ \hline
\small \textbf{Without Transliteration} & However, \small \textbf{[UNKNOWN]} said that he \small \textbf{[UNKNOWN]} president to respect their age and are also due to yell at them, but they were saying truth from miles away.  \\ \hline
\small \textbf{With Transliteration} & ``However, \small \textbf{Erdogan} said that he respects the \small \textbf{Israeli} President and his age as a result of which they yell at them , but they were saying the truth from miles away.  \\ \hline
\small \textbf{Reference} & However, later \small \textbf{Erdogan} said that he respects \small \textbf{Israeli} President and his age as well which is why he did not yell at him but whatever he was saying was miles away from truth.   \\ \hline
\end{tabular}
\end{center}
\normalsize
\caption{\label{transliteration-impact-example1} Examples of improvements from transliteration.}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|p{8cm}|}
\hline
\small \textbf{Without Transliteration} & In southern germany \small \textbf{[UNKNOWN]} a resident of the area of \small \textbf{[UNKNOWN]} \small \textbf{[UNKNOWN]} has left behind a big business group  \\ \hline
\small \textbf{With Transliteration} & A resident of the area of \small \textbf{Cuba} in south Germany \small \textbf{Adolf Merkel} has left behind a big business group  \\ \hline
\small \textbf{Reference} & \small \textbf{Adolf Merckle} of southern Germany's \small \textbf{Swabia} area has left a large business group behind  \\ \hline
\end{tabular}
\end{center}
\normalsize
\caption{\label{transliteration-impact-example2} Impact of transliteration. Note that the location name ``Swabia'' was incorrectly transliterated to ``Cuba.'' This example indicates the future room for improvement.}
\end{table}

\section{Conclusion}
\vspace{-.15cm}
In this work, we created an XML format to markup the output of specialized subtask modules and integrate alternate translations into the SMT decoder. We created a transliteration module using a character-based statistical MT system and several thousand pairs of transliterated words.  The results are promising. In particular, a qualitative analysis suggests that the transliterations were able to appropriately compete with the phrase-based translation output. This work has also opened the door to integrating additional specialized translation modules. Such modules have a potential to increase translation performance, particularly in low-resource conditions.\footnote{This research was supported in part by the European Commission through the EuroMatrixPlus project, by the US National Science Foundation under grant IIS-0713448, and by the Human Language Technology Center of Excellence.  Any opinions, findings, and conclusions or recommendations expressed in this material are the authors' and do not necessarily reflect the views of the sponsor. }

\bibliography{mt_mar_translit}

\end{document}

### 2010-04-06 papers/wmt10.pipeline/paper.tex
Working on short paper
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1559 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage[%
  colorlinks,%
  linkcolor=black,%
  anchorcolor=black,%
  citecolor=black,%
  filecolor=black,%
  menucolor=black,%
  pagecolor=black,%
  urlcolor=black,%
  %hyperindex=false,%
  bookmarksopen=true,%
  pdfusetitle,%
  pdfpagelabels,
  naturalnames
]{hyperref}
%\usepackage[colorlinks]{hyperref}

\usepackage{acl2010}
\usepackage[colorinlistoftodos]{todonotes}

\title{Empirical Questions in Machine Translation\
 --- Reproducible Results for the JHU WMT10 System}

\author{Lane Schwartz\
\thanks{Research conducted as a visiting researcher at Johns Hopkins University}\
\\University of Minnesota\\Minneapolis, MN}

\date{}

\begin{document}
\maketitle

\begin{abstract}
\end{abstract}

%\listoftodos

\section{Introduction}

\todo[inline,caption={Motivation}]{
Papers are not reproducible.
\begin{itemize}
\item Don't list code
\item Don't provide data
\item Don't publish scripts to reproduce
\end{itemize}

MT is complex, much more so than other CL research (parsing, tagging, ASR, etc)
}

\todo[inline,caption={Call to action}]{
\cite{pedersen2010}

CL Research can and should have an extremely low barrier to entry. Allow for lone researchers and small groups the opportunity to participate.
}

\section{Related Work}

\todo[inline,caption={Open code examples}]{
Moses \cite{Moses}, Joshua \cite{Joshua-WMT}, SAMT \cite{samt2006}, Cunei \cite{Phillips2009}
}

\todo[inline,caption={Open data examples}]{
Europarl \cite{Koehn-europarl}, JRC-Acquis \cite{Steinberger-2006}, Fr-En $10^9$ \cite{WMT09-Findings}, News Commentary \cite{Callison-Burch2007a,Callison-Burch2008a}
}

\todo[inline,caption={Open workflow examples}]{
\cite{LoonyBin} and \cite{experiment.perl}
}

\section{Framework}

\missingfigure{Graph of dependencies and steps}

GNU Make \cite{gnumake}

\todo[inline,color=yellow,caption={parallel make}]{
Many tasks, such as preprocessing numerous training files, are not dependent on one another. In such cases {\tt make} can be configured to execute multiple processes simultaneously on a single multi-processor machine. In cases where scheduled distributed computing environments such as the Sun Grid Engine are configured, well-designed make files can be run by make variants ({\tt distmake}, SGE {\tt qmake}, Sun Studio {\tt dmake}) which distribute tasks to multiple machines using the distributed scheduler.
}

\section{Reproducible Results}

\bibliographystyle{acl}
\bibliography{bibliography}

\end{document}

### 2010-04-20 papers/wmt10/joshua-progress.tex
add a latex for Joshua progress paper for WMT10, the author list will change by removing names who do not contribute to the new functions
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1590 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

%
% File acl2010.tex
%
% Contact  jshin@csie.ncnu.edu.tw or pkoehn@inf.ed.ac.uk
%%
%% Based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2010}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage[colorlinks]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
%\setlength\titlebox{6.5cm}    % You can expand the title box if you
% really have to

\title{Discriminative Training in Joshua with Expectation Semirings\footnote{ 
Please cite Li et al. (2009a) if you use Joshua in your research, and not this demonstration description paper. }}

\author{
Zhifei Li,\,\,\,
 Lane Schwartz,$^\star$\,\,\,
Chris Callison-Burch,\,\,\,
Chris Dyer,$^\dagger$\,\,\,
Juri Ganitkevitch,\,\,\, \\
{\bf Sanjeev Khudanpur,\,\,\, 
Wren N.\,G.\,Thornton,\,\,\,
Ziyuan Wang\,\,
Jonathan Weese\,\,
{\textnormal{and}}
\,\,\,Omar F. Zaidan}\\
Center for Language and Speech Processing, Johns Hopkins University, Baltimore, MD\\
$\dagger$ Computational Linguistics and Information Processing Lab, University of Maryland, College Park, MD\\
$\star$ Natural Language Processing Lab, University of Minnesota, Minneapolis, MN }

\date{}

\begin{document}
\maketitle

\begin{abstract}
 This paper describes Joshua, an open source hierarchical phrase-based statistical machine translation system. Results for the WMT 2010 shared task are reported for six language pairs.
\end{abstract}

\listoftodos

\section{Supporting SAMT}
\todo{juri or ccb}

\section{Semiring Parsing}
\todo{zhifei}

\section{Discriminative Training}
\todo{zhifei}

\section{Visualization}
\todo{jonny}

\section{Manual Constraints}
\todo{anni}

%\todo{Add info about grammar extraction; MT Marathon paper and speedups}
%
%\section{Results}
%
%\todo{Describe results}
%
%\begin{table}[h]
%\begin{center}
%\begin{tabular}{|l|l||c|c|c|}
%\hline
%\bf Source & \bf Target & \bf BLEU & \bf BLEU- & \bf TER \\
%& & & \bf cased & \\
%\hline
%German & English & 21.3 & 19.5 & 0.660 \\ \hline
%English & German & 15.2  & 14.6  & 0.738 \\ \hline
%French & English & 27.7 & 26.4 & 0.614 \\ \hline
%English & French & 23.8 & 22.8 & 0.681 \\ \hline
%Spanish & English & 29.0 & 27.6 & 0.595 \\ \hline
%English & Spanish & 28.1 & 26.5 & 0.596  \\ \hline
%\end{tabular}
%\end{center}
%\caption{\label{scores} Automatic metric scores for the test set newstest2010 }
%\end{table}
%
%
%\begin{table}[h]
%\begin{center}
%\begin{tabular}{|l|l||c|}
%\hline
%\bf Source & \bf Target & \bf BLEU  \\
%\hline
%German & English & 18.19  \\ \hline
%English & German & 13.57 \\ \hline
%French & English & 26.41 \\ \hline
%English & French & 25.28 \\ \hline
%Spanish & English & 25.28 \\ \hline
%English & Spanish & 24.02  \\ \hline
%\end{tabular}
%\end{center}
%\caption{\label{devtest-scores} Automatic metric scores for the development test set newstest2009}
%\end{table}

\end{document}

### 2010-04-21 papers/wmt10.pipeline/paper.tex
Work on system description paper
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1593 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \documentclass[11pt]{article}
+%\usepackage[%
+%  colorlinks,%
+%  linkcolor=black,%
+%  anchorcolor=black,%
+%  citecolor=black,%
+%  filecolor=black,%
+%  menucolor=black,%
+%  pagecolor=black,%
+%  urlcolor=black,%
+%  %hyperindex=false,%
+%  bookmarksopen=true,%
+%  pdfusetitle,%
+%  pdfpagelabels,
+%  naturalnames
+%]{hyperref}
 \usepackage{times}
@@
 \usepackage{latexsym}
-\usepackage[%
-  colorlinks,%
-  linkcolor=black,%
-  anchorcolor=black,%
-  citecolor=black,%
-  filecolor=black,%
-  menucolor=black,%
-  pagecolor=black,%
-  urlcolor=black,%
-  %hyperindex=false,%
-  bookmarksopen=true,%
-  pdfusetitle,%
-  pdfpagelabels,
-  naturalnames
-]{hyperref}
+
 %\usepackage[colorlinks]{hyperref}
+
+\usepackage{tikz}
+\usetikzlibrary{fit,shapes.misc,shapes}
 
@@
 
-\title{Empirical Questions in Machine Translation\
- --- Reproducible Results for the JHU WMT10 System}
+%\title{Empirical Questions in Machine Translation\
+ %--- Reproducible Results for the JHU WMT10 System}
+\title{Reproducible Results in Parsing-Based Machine Translation:\\
+The JHU Shared Task Submission}
 
@@
 \thanks{Research conducted as a visiting researcher at Johns Hopkins University}\
-\\University of Minnesota\\Minneapolis, MN}
+\\University of Minnesota\\Minneapolis, MN\\
+{\tt lane@cs.umn.edu} \And
+Chris Callison-Burch \\
+Johns Hopkins University \\
+Baltimore, MD\\
+{\tt ccb@cs.jhu.edu}}
 
@@
 
+\todo[inline,caption={Parsing-based SMT}] {
+Parsing-based SMT: \cite{SCALE-report} Hiero grammar extraction \cite{Lopez2008}
+}
+
 \section{Framework}
 
-\missingfigure{Graph of dependencies and steps}
+%\missingfigure{Graph of dependencies and steps}
+\begin{figure*}
+%\include{mt-workflow-preprocess}
+\include{figure.pgf}
+%\begin{tikzpicture}[shape=rounded rectangle]
+%\node (PreprocessLabel) {\underline{\bf Preprocess Data}};
+%\node (CompressedData) [style=draw,below of= PreprocessLabel]  {Compressed Data};
+%\node (DecompressedData) [style=draw,below of=CompressedData] {Decompressed Data};a
+%\node (RemoveXML) [style=draw,below of=DecompressedData,text width=28mm]  {Test Set Data Stripped of XML};
+%\node (Tokenize) [style=draw,below of= RemoveXML]  {Tokenized Data};
+%\node (Normalize) [style=draw,below of=Tokenize] {Normalized Data};
+%\node [style=draw,shape=rectangle,fit=(PreprocessLabel) (CompressedData) (DecompressedData) (RemoveXML) (Tokenize) (Normalize)] {};
+%
+%\draw [->] (CompressedData) -- (DecompressedData);
+%\draw [->] (DecompressedData) -- (RemoveXML);
+%\draw [->] (RemoveXML) -- (Tokenize);
+%\draw [->] (Tokenize) -- (Normalize);
+%\draw [->] (DecompressedData) -- (Tokenize);
+%\end{tikzpicture}
+
+\caption{Machine translation workflow}
+\end{figure*}
 
@@
 In such cases {\tt make} can be configured to execute multiple processes simultaneously on a single multi-processor machine.
-In cases where scheduled distributed computing environments such as the Sun Grid Engine are configured, well-designed make files can be run by make variants ({\tt distmake}, SGE {\tt qmake}, Sun Studio {\tt dmake}) which distribute tasks to multiple machines using the distributed scheduler.
+In cases where scheduled distributed computing environments such as the Sun Grid Engine are configured, make files can be processed by scheduler-aware {\tt make} variants ({\tt distmake}, SGE {\tt qmake}, Sun Studio {\tt dmake}) which distribute outstanding tasks to available distributed machines using the relevant distributed scheduler.
 }
@@
 \section{Reproducible Results}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|l||c|c|c|}
+\hline
+\bf Source & \bf Target & \bf BLEU & \bf BLEU- & \bf TER \\
+& & & \bf cased & \\
+\hline
+German & English & 21.3 & 19.5 & 0.660 \\ \hline
+English & German & 15.2  & 14.6  & 0.738 \\ \hline
+French & English & 27.7 & 26.4 & 0.614 \\ \hline
+English & French & 23.8 & 22.8 & 0.681 \\ \hline
+Spanish & English & 29.0 & 27.6 & 0.595 \\ \hline
+English & Spanish & 28.1 & 26.5 & 0.596  \\ \hline
+\end{tabular}
+\end{center}
+\caption{\label{scores} Automatic metric scores for the test set newstest2010 }
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|l||c|}
+\hline
+\bf Source & \bf Target & \bf BLEU  \\
+\hline
+German & English & 18.19  \\ \hline
+English & German & 13.57 \\ \hline
+French & English & 26.41 \\ \hline
+English & French & 25.28 \\ \hline
+Spanish & English & 25.28 \\ \hline
+English & Spanish & 24.02  \\ \hline
+\end{tabular}
+\end{center}
+\caption{\label{devtest-scores} Automatic metric scores for the development test set newstest2009}
+\end{table}
+
+\todo[inline,caption={URL}]{
+List how to download: \url{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz}
+}
+
+\section*{Acknowledgements}
+This work was supported by the DARPA GALE program (Contract No HR0011-06-2-0001).
 

### 2010-04-22 papers/wmt10.pipeline/paper.tex
Work on shared task paper
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1595 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \author{Lane Schwartz\
-\thanks{Research conducted as a visiting researcher at Johns Hopkins University}\
+\thanks{\ Research conducted as a visiting researcher at Johns Hopkins University}\
 \\University of Minnesota\\Minneapolis, MN\\
@@
 
-\todo[inline,caption={Motivation}]{
-Papers are not reproducible.
-\begin{itemize}
-\item Don't list code
-\item Don't provide data
-\item Don't publish scripts to reproduce
-\end{itemize}
+%\todo[inline,color=yellow,caption={Call to action}]{
 
-MT is complex, much more so than other CL research (parsing, tagging, ASR, etc)
-}
+Research investigating natural language process and computational linguistics can and should have an extremely low barrier to entry.
+The data with which we work is customarily available in common electronic formats.
+The computational techniques which we apply can typically be performed on commodity computing resources which are widely available.
+In short, there should be no reason why small research groups and even lone researchers should not be able to join and make substantive contributions furthering our field.
+%
+The reality is less encouraging.
 
-\todo[inline,caption={Call to action}]{
-\cite{pedersen2010}
+Many published articles describe novel techniques and provide interesting results, yet fail to describe technical details in sufficient detail to allow their results to be reproduced by other researchers.
+While there are notable and laudable exceptions, many publications fail to provide the source code and scripts necessary to reproduce results.
+The use of private data, not freely available for download by any interested researcher only compounds these problems.
+\newcite{pedersen2010} rightly argues that the implementation details so often ignored in publications are in fact essential for our research to be reproducible science.
 
-CL Research can and should have an extremely low barrier to entry. Allow for lone researchers and small groups the opportunity to participate.
-}
+Reproducibility in machine translation is made more challenging by the complexity of experimental workflows.
+Results in machine translation tasks are dependent on a cascade of processing steps and configurations.
+While interesting subsets of these usually appear in experimental descriptions, many steps (preprocessing techniques, alignment parameters, translation rule extraction parameters, language model parameters, list of features used) are invariably omitted, even though these configurations are often critical to reproducing results.
+
+This paper describes the Johns Hopkins University submission to the 2010 Workshop on Statistical Machine Translation shared translation task.
+Links to the software, scripts, and configurations used to run the experiments described herein are provided.
+The remainder of this paper is structured as follows.
+Section \ref{Related Work} lists the major examples of publicly available open source machine translation systems, parallel corpora, and machine translation workflow management systems.
+Section \ref{Framework} describes the experimental design used to run the shared task translations.
+Section \ref{Results} presents the shared task results.
 
 \section{Related Work}
+\label{Related Work}
 
-\todo[inline,caption={Open code examples}]{
-Moses \cite{Moses}, Joshua \cite{Joshua-WMT}, SAMT \cite{samt2006}, Cunei \cite{Phillips2009}
-}
+The last four years have witnessed the implementation and release of numerous open source machine translation systems.
+The widely used Moses system implements the standard phrase-based translation model.
+Parsing-based translation models are implemented by SAMT \cite{samt2006} and Joshua \cite{Joshua-WMT}.
+Cunei \cite{Phillips2009} implements statistical example-based translation.
+The SRILM \cite{Stolcke2002} and IRSTLM \cite{IRSTLM} toolkits enable efficient training and querying of n-gram language models.
+\newcite{Phramer} and \newcite{schwartz08mclc} provide other open-source implementations of phrase-based and hierarchical decoders, respectively.
 
-\todo[inline,caption={Open data examples}]{
-Europarl \cite{Koehn-europarl}, JRC-Acquis \cite{Steinberger-2006}, Fr-En $10^9$ \cite{WMT09-Findings}, News Commentary \cite{Callison-Burch2007a,Callison-Burch2008a}
-}
+Freely available parallel corpora for numerous European languages have also been released in recent years.
+These include the Europarl \cite{Koehn-europarl} and JRC-Acquis \cite{Steinberger-2006} legislative corpora, each of which includes data for most EU language pairs.
+The smaller News Commentary \cite{Callison-Burch2007a,Callison-Burch2008a} provide smaller amounts of parallel data in the news genre.
+The recent Fr-En $10^9$ \cite{WMT09-Findings} corpus aggregates huge numbers of parallel French-English sentences from the web.
 
-\todo[inline,caption={Open workflow examples}]{
-\cite{LoonyBin} and \cite{experiment.perl}
-}
+Systems to address the complex workflows required to run non-trivial machine translation experiments have also been developed.
+These include {\tt experiment.perl} \cite{experiment.perl}, developed as a workflow management system at the University of Edinburgh, and LoonyBin \cite{LoonyBin}, a general hyperworkflow management utility from Carnegie Melon University.
 
 \todo[inline,caption={Parsing-based SMT}] {
-Parsing-based SMT: \cite{SCALE-report} Hiero grammar extraction \cite{Lopez2008}
+Hiero grammar extraction \cite{Lopez2008}
 }
@@
 \section{Framework}
+\label{Framework}
 
@@
 
-\caption{Machine translation workflow}
+\caption{Machine translation workflow.
+The scripts and configuration files used to implement and run this workflow are available for download at \url{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz}}
 \end{figure*}
@@
 
+\todo[inline,caption={experimental design}] {
+
+\begin{itemize}
+\item Data used: Europarl \cite{Koehn-europarl}, Fr-En $10^9$ \cite{WMT09-Findings}, News Commentary \cite{Callison-Burch2007a,Callison-Burch2008a}. No restricted data from the LDC. 
+\item Preprocessing: Removed XML markup; Tokenized and lowercased training data using scripts provided for the WMT10 shared task
+\item Training corpus subsampling, as described in \cite{Joshua-WMT}
+\item Corpus aligned using Berkeley aligner \cite{liang06alignment}
+\item Grammar extraction using Joshua suffix array grammar extraction \cite{Schwartz-PBML}; for Fr-En corpus used \cite{Lopez2008} for speed
+\item Language models built using SRILM \cite{Stolcke2002}.
+\item Decoder used was Joshua \cite{Joshua-WMT}
+\item Minimum error rate training \cite{Zaidan2009}
+\end{itemize}
+}
+
+\todo[inline,color=yellow] {
+Our experiments used hierarchical phrase-based grammars containing exactly two nonterminals - the wildcard nonterminal X, and S, used to glue together neighboring constituents.
+Recent work has shown that parsing-based machine translation using SAMT \cite{samt2006} grammars with rich nonterminal sets can demonstrate substantial gains over hierarchical grammars for certain language pairs \cite{SCALE-report}.
+Joshua supports such grammars; the experimental workflow presented here could easily be extended in future research to incorporate the use of SAMT grammars with additional language pairs.
+}
+
 \section{Reproducible Results}
+\label{Results}
 

### 2010-04-22 papers/wmt10.pipeline/paper.tex
More work
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1598 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \documentclass[11pt]{article}
-%\usepackage[%
-%  colorlinks,%
-%  linkcolor=black,%
-%  anchorcolor=black,%
-%  citecolor=black,%
-%  filecolor=black,%
-%  menucolor=black,%
-%  pagecolor=black,%
-%  urlcolor=black,%
-%  %hyperindex=false,%
-%  bookmarksopen=true,%
-%  pdfusetitle,%
-%  pdfpagelabels,
-%  naturalnames
-%]{hyperref}
 \usepackage{times}
@@
 \usepackage{latexsym}
-
-%\usepackage[colorlinks]{hyperref}
-
 \usepackage{tikz}
 \usetikzlibrary{fit,shapes.misc,shapes}
-
 \usepackage{acl2010}
@@
 
-%\title{Empirical Questions in Machine Translation\
- %--- Reproducible Results for the JHU WMT10 System}
 \title{Reproducible Results in Parsing-Based Machine Translation:\\
@@
 
-%\listoftodos
-
 \section{Introduction}
-
-%\todo[inline,color=yellow,caption={Call to action}]{
 
@@
 The widely used Moses system implements the standard phrase-based translation model.
-Parsing-based translation models are implemented by SAMT \cite{samt2006} and Joshua \cite{Joshua-WMT}.
+Parsing-based translation models are implemented by Joshua \cite{Joshua-WMT} and SAMT \cite{samt2006}.
 Cunei \cite{Phillips2009} implements statistical example-based translation.
-The SRILM \cite{Stolcke2002} and IRSTLM \cite{IRSTLM} toolkits enable efficient training and querying of n-gram language models.
-\newcite{Phramer} and \newcite{schwartz08mclc} provide other open-source implementations of phrase-based and hierarchical decoders, respectively.
+\newcite{Phramer} and \newcite{schwartz08mclc} respectively provide additional open-source implementations of phrase-based and hierarchical decoders.
+
+The SRILM \cite{Stolcke2002}, IRSTLM \cite{IRSTLM}, and RandLM \cite{Talbot2007a} toolkits enable efficient training and querying of n-gram language models.
 
@@
 
-Systems to address the complex workflows required to run non-trivial machine translation experiments have also been developed.
+Open source systems to address the complex workflows required to run non-trivial machine translation experiments have also been developed.
 These include {\tt experiment.perl} \cite{experiment.perl}, developed as a workflow management system at the University of Edinburgh, and LoonyBin \cite{LoonyBin}, a general hyperworkflow management utility from Carnegie Melon University.
-
-\todo[inline,caption={Parsing-based SMT}] {
-Hiero grammar extraction \cite{Lopez2008}
-}
 
@@
 
-%\missingfigure{Graph of dependencies and steps}
 \begin{figure*}
-%\include{mt-workflow-preprocess}
 \include{figure.pgf}
-%\begin{tikzpicture}[shape=rounded rectangle]
-%\node (PreprocessLabel) {\underline{\bf Preprocess Data}};
-%\node (CompressedData) [style=draw,below of= PreprocessLabel]  {Compressed Data};
-%\node (DecompressedData) [style=draw,below of=CompressedData] {Decompressed Data};a
-%\node (RemoveXML) [style=draw,below of=DecompressedData,text width=28mm]  {Test Set Data Stripped of XML};
-%\node (Tokenize) [style=draw,below of= RemoveXML]  {Tokenized Data};
-%\node (Normalize) [style=draw,below of=Tokenize] {Normalized Data};
-%\node [style=draw,shape=rectangle,fit=(PreprocessLabel) (CompressedData) (DecompressedData) (RemoveXML) (Tokenize) (Normalize)] {};
-%
-%\draw [->] (CompressedData) -- (DecompressedData);
-%\draw [->] (DecompressedData) -- (RemoveXML);
-%\draw [->] (RemoveXML) -- (Tokenize);
-%\draw [->] (Tokenize) -- (Normalize);
-%\draw [->] (DecompressedData) -- (Tokenize);
-%\end{tikzpicture}
 
 \caption{Machine translation workflow.
+Nodes in grey indicate software and scripts.
 The scripts and configuration files used to implement and run this workflow are available for download at \url{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz}}

### 2010-04-22 papers/wmt10/joshua-progress.tex
Fixed title footnote by changing it to use \thanks
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1591 0ae5e6b2-d358-4f09-a895-f82f13dd62a4
add the full list of names
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1599 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\title{Discriminative Training in Joshua with Expectation Semirings\footnote{ 
-Please cite Li et al. (2009a) if you use Joshua in your research, and not this demonstration description paper. }}
+\title{Joshua: An Open Source Toolkit for Parsing-based Machine Translation\thanks{ 
+Please cite Li et al. (2009a) if you use Joshua in your research, and not this paper. }}
 
@@
 Zhifei Li,\,\,\,
- Lane Schwartz,$^\star$\,\,\,
 Chris Callison-Burch,\,\,\,
 Chris Dyer,$^\dagger$\,\,\,
-Juri Ganitkevitch,\,\,\, \\
-{\bf Sanjeev Khudanpur,\,\,\, 
+Juri Ganitkevitch,\,\,\,
+Ann Irvine,\,\,\, 
+Sanjeev Khudanpur,\,\,\,\\ {\bf
+Lane Schwartz,$^\star$\,\,\, 
 Wren N.\,G.\,Thornton,\,\,\,
-Ziyuan Wang\,\,
+Ziyuan Wang,\,\,
 Jonathan Weese\,\,
@@
 \begin{abstract}
- This paper describes Joshua, an open source hierarchical phrase-based statistical machine translation system. Results for the WMT 2010 shared task are reported for six language pairs.
+ This paper describes Joshua 2.0, an open source hierarchical phrase-based statistical machine translation system.
 \end{abstract}

### 2010-04-22 papers/wmt10.pipeline/paper.tex
Updating paper
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1600 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \usepackage{acl2010}
-\usepackage[colorinlistoftodos]{todonotes}
+%\usepackage[colorinlistoftodos]{todonotes}
 
@@
 While there are notable and laudable exceptions, many publications fail to provide the source code and scripts necessary to reproduce results.
-The use of private data, not freely available for download by any interested researcher only compounds these problems.
+The use of restricted data, not freely available for download by any interested researcher only compounds these problems.
 \newcite{pedersen2010} rightly argues that the implementation details so often ignored in publications are in fact essential for our research to be reproducible science.
@@
 Section \ref{Related Work} lists the major examples of publicly available open source machine translation systems, parallel corpora, and machine translation workflow management systems.
-Section \ref{Framework} describes the experimental design used to run the shared task translations.
+Section \ref{Framework} describes the experimental workflow used to run the shared task translations, with the corresponding experimental design in section \ref{Design}.
 Section \ref{Results} presents the shared task results.
+
+\begin{figure*}[htbp]
+\include{figure.pgf}
+\caption{Machine translation workflow.
+Nodes in grey indicate software and scripts.
+The scripts and configuration files used to implement and run this workflow are available for download at \url{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz}}
+\label{dependencyGraph}
+\end{figure*}
 
@@
 The last four years have witnessed the implementation and release of numerous open source machine translation systems.
-The widely used Moses system implements the standard phrase-based translation model.
-Parsing-based translation models are implemented by Joshua \cite{Joshua-WMT} and SAMT \cite{samt2006}.
+The widely used Moses system \cite{moses} implements the standard phrase-based translation model.
+Parsing-based translation models are implemented by Joshua \cite{Joshua-WMT}, SAMT \cite{samt2006}, and cdec \cite{cdec}.
 Cunei \cite{Phillips2009} implements statistical example-based translation.
@@
 
-\section{Framework}
+\section{Experimental Workflow Management}
 \label{Framework}
 
-\begin{figure*}
-\include{figure.pgf}
+Running a statistical machine translation system to achieve state-of-the-art performance involves the configuration and execution of numerous interdependent intermediate tools.
+To manage task dependencies and tool configuration, our shared task workflow consists of a set of dependency scripts written for GNU Make \cite{gnumake}.
 
-\caption{Machine translation workflow.
-Nodes in grey indicate software and scripts.
-The scripts and configuration files used to implement and run this workflow are available for download at \url{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz}}
-\end{figure*}
+Figure \ref{dependencyGraph} shows a graph depicting the steps in our experimental workflow, and the dependencies between steps.
+Each node in the graph represents a step in the workflow; each step is with a Make script that defines how to run the tools required in that step.
+In each experiment, an additional configuration script is provided for each experimental step, defining the parameters to be used when running that step in the current experiment.
+Optional front-end wrapper scripts can also be provided, allowing for a complete experiment to be run - from downloading data and software through truecasing translated results - by executing a single make file.
 
-GNU Make \cite{gnumake}
-
-\todo[inline,color=yellow,caption={parallel make}]{
+This framework is also conducive to parallelization.
 Many tasks, such as preprocessing numerous training files, are not dependent on one another.
@@
 In cases where scheduled distributed computing environments such as the Sun Grid Engine are configured, make files can be processed by scheduler-aware {\tt make} variants ({\tt distmake}, SGE {\tt qmake}, Sun Studio {\tt dmake}) which distribute outstanding tasks to available distributed machines using the relevant distributed scheduler.
-}
 
-\todo[inline,caption={experimental design}] {
+\section{Experimental Configuration}
+\label{Design}
 
-\begin{itemize}
-\item Data used: Europarl \cite{Koehn-europarl}, Fr-En $10^9$ \cite{WMT09-Findings}, News Commentary \cite{Callison-Burch2007a,Callison-Burch2008a}. No restricted data from the LDC. 
-\item Preprocessing: Removed XML markup; Tokenized and lowercased training data using scripts provided for the WMT10 shared task
-\item Training corpus subsampling, as described in \cite{Joshua-WMT}
-\item Corpus aligned using Berkeley aligner \cite{liang06alignment}
-\item Grammar extraction using Joshua suffix array grammar extraction \cite{Schwartz-PBML}; for Fr-En corpus used \cite{Lopez2008} for speed
-\item Language models built using SRILM \cite{Stolcke2002}.
-\item Decoder used was Joshua \cite{Joshua-WMT}
-\item Minimum error rate training \cite{Zaidan2009}
-\end{itemize}
-}
+\begin{table*}[htb]
+\begin{center}
+\begin{tabular}{|l|l|l|}
+\hline
+\bf Source & \bf Target & \bf Parallel Corpora \\ \hline
+German & English & news-commentary10.de-en europarl-v5.de-en \\ \hline
+English & German & news-commentary10.de-en europarl-v5.de-en \\ \hline
+French & English & news-commentary10.fr-en europarl-v5.fr-en giga-fren.release2 undoc.2000.en-fr \\ \hline
+English & French & news-commentary10.fr-en europarl-v5.fr-en giga-fren.release2 undoc.2000.en-fr \\ \hline
+Spanish & English & news-commentary10.es-en europarl-v5.es-en undoc.2000.en-es \\ \hline
+English & Spanish & news-commentary10.es-en europarl-v5.es-en undoc.2000.en-es  \\ \hline
+\end{tabular}
+\end{center}
+\caption{Parallel training data used for training translation model, per language pair}
+\label{parallelCorpora}
+\end{table*}
 
-\todo[inline,color=yellow] {
+\begin{table*}[htb]
+\begin{center}
+\begin{tabular}{|l|l|}
+\hline
+\bf Target & \bf Monolingual Corpora \\ \hline
+English & europarl-v5.en news-commentary10.en news.en.shuffled undoc.2000.en-fr.en giga-fren.release2.en \\ \hline
+French & europarl-v5.fr news-commentary10.fr news.fr.shuffled undoc.2000.en-fr.fr giga-fren.release2.fr \\ \hline
+German & europarl-v5.de news-commentary10.de news.de.shuffled \\ \hline
+Spanish & europarl-v5.es news-commentary10.es news.es.shuffled undoc.2000.en-es.es\\ \hline
+\end{tabular}
+\end{center}
+\caption{Monolingual training data used for training language model, per target language}
+\label{monolingualCorpora}
+\end{table*}
+
+Experimental workflows were configured\footnote{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz} and run for six language pairs in the translation shared task: English-French, English-German, English-Spanish, French-English, German-English, and Spanish-English.
+
+In all experiments, only data freely available for download was used.
+No restricted data from the LDC or other sources was used.
+Table \ref{parallelCorpora} lists the parallel corpora used in training the translation model for each experiment.
+The monolingual corpora used in training each target language model are listed in table \ref{monolingualCorpora}.
+In all experiments, news-test2008 was used as a development tuning corpus during minimum error rate training; newstest2009 was used as a development test set.
+The shared task data set newstest2010 was used as a final blind test set.
+
+All data was automatically downloaded, unzipped, and preprocessed prior to use.
+Files provided in XML format were converted to plain text by selecting lines with {\tt <seg>} tags, then removing the beginning and end tags for each segment; this processing was applied using GNU grep and sed.
+The {\tt tokenize.perl} and {\tt lowercase.perl} scripts provided for the shared task\footnote{http://www.statmt.org/wmt08/scripts.tgz with md5sum: \\ {\tt tokenize.perl} 45cd1832827131013245eca76481441a \\ {\tt lowercase.perl} a1958ab429b1e29d379063c3b9cd7062} were applied to all data.
+
+Interpolated n-gram language models for the four target languages were built using the SRI Language Model Toolkit\footnote{ http://www-speech.sri.com/projects/srilm \\ SRILM version 1.5.7.
+Our experimental workflow requires that SRILM be compiled separately, with the \$SRILM environment variable set to the install location.}, with n-gram order set to 5.
+The \newcite{ChenGoodman98} technique for modified Kneser-Ney discounting \cite{Kneser1995} was applied during language model training.
+
+Following \newcite{Joshua-WMT}, a subset of the available training sentences is selected via subsampling; training sentences are selected based on the estimated likelihood of that sentence being useful later for translating a particular test corpus.
+
+Given a subsampled parallel training corpus, word alignment is performed using the Berkeley aligner\footnote{http://berkeleyaligner.googlecode.com/files/berkeleyaligner\\\_unsupervised-2.1.tar.gz --- Berkeley aligner version 2.1} \cite{liang06alignment}.
+
+For each language pair, a synchronous context free translation grammar is extracted for a particular test set, following the methods of \newcite{Lopez2008} as implemented in \cite{Schwartz-PBML}.
+For the largest training sets (French-English and English-French) the original \cite{Lopez2008} implementation included with Hiero was used to save time during training\footnote{It is expected that using the Joshua implementation should result in nearly identical results, albeit with somewhat more time required to extract the grammar.}.
+
 Our experiments used hierarchical phrase-based grammars containing exactly two nonterminals - the wildcard nonterminal X, and S, used to glue together neighboring constituents.
@@
 Joshua supports such grammars; the experimental workflow presented here could easily be extended in future research to incorporate the use of SAMT grammars with additional language pairs.
-}
+The Z-MERT implementation \cite{Zaidan2009} of minimum error rate training was used for parameter tuning.
+Tuned grammars were used by Joshua to translate all test sets.
+
+The code used for subsampling, grammar extraction, decoding, and minimum error rate training is provided with Joshua\footnote{http://sourceforge.net/projects/joshua/files/joshua/1.3/joshua-1.3.tgz/download --- Joshua version 1.3}, with the exception of the original \cite{Lopez2008} grammar extraction code.
 
@@
 
-\todo[inline,caption={URL}]{
-List how to download: \url{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz}
-}
-
 \section*{Acknowledgements}

### 2010-04-23 papers/wmt10.pipeline/paper.tex
Draft complete
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1601 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \begin{abstract}
+We present the JHU submission to the 2010 WMT shared translation task.
+We describe processing steps using open data and open source software used in our submission, and provide the scripts and configurations required to train, tune, and test the machine translation system described herein.
 \end{abstract}
@@
 
+Because of the use of subsampling, the extracted translation grammars are targeted for use with a specific test set.
+Our experiments were begun prior to the release of the blind newstest2010 shared task test set.
+Subsampling was performed using the development tuning set, news-test2008, and the development test set, newstest2009.
+Once the newstest2010 test set was released, the process of subsampling, alignment, and grammar extraction was repeated to obtain translation grammars targeted for use with the shared task test set.
+
 Our experiments used hierarchical phrase-based grammars containing exactly two nonterminals - the wildcard nonterminal X, and S, used to glue together neighboring constituents.
@@
 Joshua supports such grammars; the experimental workflow presented here could easily be extended in future research to incorporate the use of SAMT grammars with additional language pairs.
-The Z-MERT implementation \cite{Zaidan2009} of minimum error rate training was used for parameter tuning.
+
+The Z-MERT implementation \cite{Zaidan2009} of minimum error rate training \cite{Och2003c} was used for parameter tuning.
 Tuned grammars were used by Joshua to translate all test sets.
+The Joshua decoder produces n-best lists of translations.
 
-The code used for subsampling, grammar extraction, decoding, and minimum error rate training is provided with Joshua\footnote{http://sourceforge.net/projects/joshua/files/joshua/1.3/joshua-1.3.tgz/download --- Joshua version 1.3}, with the exception of the original \cite{Lopez2008} grammar extraction code.
+Rather than simply selecting the top candidate from each list, we take the preferred candidate after perform minimum Bayes risk rescoring \cite{Kumar2004b}.
 
-\section{Reproducible Results}
+Once a single translation has been extracted for each sentence in the test set, we repeat the procedures described above to train language and translation models for use in translating lowercased results into a more human-readable truecased form.
+A truecase language model is trained as above, but on the tokenized (but not normalized) monolingual target language corpus.
+Monotone word alignments are deterministically created, mapping normalized lowercase training text to the original truecase text.
+As in bilingual translation, subsampling is performed for the training set, and a translation grammar for lowercase-to-truecase is extracted.
+No tuning is performed.
+The Joshua decoder is used to translate the lowercased target language test results into truecase format.
+The {\tt detokenize.perl} and {\tt wrap-xml.perl} scripts provided for the shared task were manually applied to truecased translation results prior to final submission of results.
+
+The code used for subsampling, grammar extraction, decoding, minimum error rate training, and minimum Bayes risk rescoring is provided with Joshua\footnote{http://sourceforge.net/projects/joshua/files/joshua/1.3/joshua-1.3.tgz/download --- Joshua version 1.3}, with the exception of the original \cite{Lopez2008} grammar extraction implementation.
+
+\section{Experimental Results}
 \label{Results}
+
+The experiments described in sections \ref{Framework} and \ref{Design} above provided truecased translations for six language pairs in the translation shared task: English-French, English-German, English-Spanish, French-English, German-English, and Spanish-English.
+Table \ref{scores} lists the automatic metric scores for the newstest2010 test set, according to the BLEU \cite{Papineni2002} and TER \cite{TER} metrics.
 
@@
 \end{table}
+
+The submitted system ranked highest among shared task participants for the German-English task, according to TER.
+
+In order to provide points of comparison with the 2009 Workshop on Statistical Machine Translation shared translation task participants,  table \ref{devtest-scores} lists automatic metric scores for our systems' translations of the newstest2009 test set, which we used as a development test set.
 

### 2010-04-23 papers/wmt10.pipeline/paper.tex
Tidying up
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1602 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \\University of Minnesota\\Minneapolis, MN\\
-{\tt lane@cs.umn.edu} \And
-Chris Callison-Burch \\
-Johns Hopkins University \\
-Baltimore, MD\\
-{\tt ccb@cs.jhu.edu}}
+{\tt lane@cs.umn.edu}% \And
+%Chris Callison-Burch \\
+%Johns Hopkins University \\
+%Baltimore, MD\\
+%{\tt ccb@cs.jhu.edu}
+}
 
@@
 \begin{abstract}
-We present the JHU submission to the 2010 WMT shared translation task.
-We describe processing steps using open data and open source software used in our submission, and provide the scripts and configurations required to train, tune, and test the machine translation system described herein.
+We present the Johns Hopkins University submission to the 2010 WMT shared translation task.
+We describe processing steps using open data and open source software used in our submission, and provide the scripts and configurations required to train, tune, and test our machine translation system.
 \end{abstract}
@@
 
-Research investigating natural language process and computational linguistics can and should have an extremely low barrier to entry.
+Research investigating natural language processing and computational linguistics can and should have an extremely low barrier to entry.
 The data with which we work is customarily available in common electronic formats.
@@
 These include the Europarl \cite{Koehn-europarl} and JRC-Acquis \cite{Steinberger-2006} legislative corpora, each of which includes data for most EU language pairs.
-The smaller News Commentary \cite{Callison-Burch2007a,Callison-Burch2008a} provide smaller amounts of parallel data in the news genre.
+The smaller News Commentary corpora \cite{Callison-Burch2007a,Callison-Burch2008a} provide smaller amounts of parallel data in the news genre.
 The recent Fr-En $10^9$ \cite{WMT09-Findings} corpus aggregates huge numbers of parallel French-English sentences from the web.
@@
 Figure \ref{dependencyGraph} shows a graph depicting the steps in our experimental workflow, and the dependencies between steps.
-Each node in the graph represents a step in the workflow; each step is with a Make script that defines how to run the tools required in that step.
+Each node in the graph represents a step in the workflow; each step is implemented as a Make script that defines how to run the tools required in that step.
 In each experiment, an additional configuration script is provided for each experimental step, defining the parameters to be used when running that step in the current experiment.
@@
 
-Following \newcite{Joshua-WMT}, a subset of the available training sentences is selected via subsampling; training sentences are selected based on the estimated likelihood of that sentence being useful later for translating a particular test corpus.
+Following \newcite{Joshua-WMT}, a subset of the available training sentences was selected via subsampling; training sentences are selected based on the estimated likelihood of each sentence being useful later for translating a particular test corpus.
 
@@
 Our experiments were begun prior to the release of the blind newstest2010 shared task test set.
-Subsampling was performed using the development tuning set, news-test2008, and the development test set, newstest2009.
+Subsampling was performed for the development tuning set, news-test2008, and the development test set, newstest2009.
 Once the newstest2010 test set was released, the process of subsampling, alignment, and grammar extraction was repeated to obtain translation grammars targeted for use with the shared task test set.

### 2010-04-23 papers/wmt10/joshua-progress.tex
more stuff
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1604 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\title{Joshua: An Open Source Toolkit for Parsing-based Machine Translation\thanks{ 
-Please cite Li et al. (2009a) if you use Joshua in your research, and not this paper. }}
+\newcommand{\joshua}{\textbf{Joshua}}
+
+\title{2010 Progress Report to Joshua Toolkit : An Open Source Toolkit for Parsing-based Machine Translation\thanks{ 
+Please cite Li et al. (2009a) if you use Joshua in your research, and not this paper.}}
 
@@
 Chris Callison-Burch,\,\,\,
-Chris Dyer,$^\dagger$\,\,\,
+%Chris Dyer,$^\dagger$\,\,\,
 Juri Ganitkevitch,\,\,\,
 Ann Irvine,\,\,\, 
-Sanjeev Khudanpur,\,\,\,\\ {\bf
+%Sanjeev Khudanpur,\,\,\,
 Lane Schwartz,$^\star$\,\,\, 
+\\ {\bf
 Wren N.\,G.\,Thornton,\,\,\,
@@
 Center for Language and Speech Processing, Johns Hopkins University, Baltimore, MD\\
-$\dagger$ Computational Linguistics and Information Processing Lab, University of Maryland, College Park, MD\\
+%$\dagger$ Computational Linguistics and Information Processing Lab, University of Maryland, College Park, MD\\
 $\star$ Natural Language Processing Lab, University of Minnesota, Minneapolis, MN }
@@
 \begin{document}
+
 \maketitle
-
 \begin{abstract}
- This paper describes Joshua 2.0, an open source hierarchical phrase-based statistical machine translation system.
+We describe the progress we have made in the past year to \textbf{Joshua} \cite{joshua-wmt09}, 
+an open-source toolkit for parsing-based machine translation.
+The main new functions we have added into the toolkit include: supporting for the syntax-augmented machine translation (SAMT) grammar,
+supporting mannual constraints in the input to constrain the decoding,  
+semiring parsing for unified dynamic programming, hypergraph-based 
+discriminative training for better feature engineering, visualization of the derivation
+trees and hypergraphs, and a unified pipeline for MT experiments.
 \end{abstract}
@@
 
+\section{Introduction}
+
+\textbf{Joshua} \cite{joshua-wmt09} is an open-source toolkit (written in Java) for parsing-based machine translation.
+In the first version, it  has inimplemented all the essential algorithms described 
+in \newcite{hiero-2007}: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.  
+It has also implemented suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007} 
+and minimum error rate training \cite{och-mert}. 
+Additionally, parallel and distributed computing techniques have been exploited 
+to make it scalable \cite{Joshua-old}. In the design of \joshua, we have applied general
+principles of software engineering to achieve three
+major goals: extensibility, end-to-end coherence,
+and scalability.
+
+The toolkit has then been used by the authors and several
+other groups in their daily research, and thus much progress has
+been made to the first version. 
+In this report, we will describe the main new functions we have added into the 
+toolkit, which include: supporting for the syntax-augmented machine translation (SAMT) grammar \cite{???},
+supporting mannual constraints in the input to constrain the decoding \cite{???},  
+semiring parsing for unified dynamic programming \cite{li-eisner:2009:EMNLP}, 
+hypergraph-based discriminative training for better feature engineering \cite{oracle-extraction-naacl09}, 
+visualization of the derivation trees and hypergraphs \cite{???}, 
+and a unified pipeline for MT experiments \cite{???}.
+
+Below we give a short description for each of the new functions added.
+
 \section{Supporting SAMT}
 \todo{juri or ccb}
+
+\section{Supporting Manual Constraints}
+\todo{anni}
 
@@
 
+\cite{li-eisner:2009:EMNLP}
+
 \section{Discriminative Training}
 \todo{zhifei}
+
+\cite{variational-decoding-acl09}
+\cite{oracle-extraction-naacl09}
 
@@
 
-\section{Manual Constraints}
-\todo{anni}
+\section{MT Pipeline}
+\todo{lane}
 
-%\todo{Add info about grammar extraction; MT Marathon paper and speedups}
-%
-%\section{Results}
-%
-%\todo{Describe results}
-%
-%\begin{table}[h]
-%\begin{center}
-%\begin{tabular}{|l|l||c|c|c|}
-%\hline
-%\bf Source & \bf Target & \bf BLEU & \bf BLEU- & \bf TER \\
-%& & & \bf cased & \\
-%\hline
-%German & English & 21.3 & 19.5 & 0.660 \\ \hline
-%English & German & 15.2  & 14.6  & 0.738 \\ \hline
-%French & English & 27.7 & 26.4 & 0.614 \\ \hline
-%English & French & 23.8 & 22.8 & 0.681 \\ \hline
-%Spanish & English & 29.0 & 27.6 & 0.595 \\ \hline
-%English & Spanish & 28.1 & 26.5 & 0.596  \\ \hline
-%\end{tabular}
-%\end{center}
-%\caption{\label{scores} Automatic metric scores for the test set newstest2010 }
-%\end{table}
-%
-%
-%\begin{table}[h]
-%\begin{center}
-%\begin{tabular}{|l|l||c|}
-%\hline
-%\bf Source & \bf Target & \bf BLEU  \\
-%\hline
-%German & English & 18.19  \\ \hline
-%English & German & 13.57 \\ \hline
-%French & English & 26.41 \\ \hline
-%English & French & 25.28 \\ \hline
-%Spanish & English & 25.28 \\ \hline
-%English & Spanish & 24.02  \\ \hline
-%\end{tabular}
-%\end{center}
-%\caption{\label{devtest-scores} Automatic metric scores for the development test set newstest2009}
-%\end{table}
+\bibliographystyle{acl}
+\bibliography{machinetranslation}
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Changed title
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1606 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\title{2010 Progress Report to Joshua Toolkit : An Open Source Toolkit for Parsing-based Machine Translation\thanks{ 
-Please cite Li et al. (2009a) if you use Joshua in your research, and not this paper.}}
+\title{Joshua 2.0: An Open Source Toolkit for Machine Translation \\with Syntax, Semirings and Other Goodies }
 
@@
 \textbf{Joshua} \cite{joshua-wmt09} is an open-source toolkit (written in Java) for parsing-based machine translation.
-In the first version, it  has inimplemented all the essential algorithms described 
+The first version implemented all the essential algorithms described 
 in \newcite{hiero-2007}: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.  
-It has also implemented suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007} 
+The initial release also implemented suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007} 
 and minimum error rate training \cite{och-mert}. 
-Additionally, parallel and distributed computing techniques have been exploited 
+Additionally, parallel and distributed computing techniques were exploited 
 to make it scalable \cite{Joshua-old}. In the design of \joshua, we have applied general
@@
 
-\section{Supporting SAMT}
+\section{Support for Syntax Augmented MT}
 \todo{juri or ccb}
 
-\section{Supporting Manual Constraints}
+\section{Manual Constraints on the Input}
 \todo{anni}

### 2010-04-23 papers/wmt10/joshua-progress.tex
added joshua-viz to bibliographies; added a few paragraphs for wmt10 about visualization
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1607 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \section{Visualization}
-\todo{jonny}
+
+We have introduced tools for visualizing two of the main data structures used
+in Joshua \cite{joshua-viz}. Each visualization tool is implemented as a
+stand-alone program that hooks into the pipeline, but they are distributed with
+Joshua. The first program is a hypergraph visualizer. The user can choose from
+a set of input sentences, then call the decoder to build the hypergraph.
+
+The second is a derivation-tree visualizer. Joshua can produce output formatted
+as a parse tree, where each nonterminal has been annotated with its source-side
+span. The visualizer can read in multiple n-best lists in this format, then
+display the resulting derivation trees side-by-side. We have found that
+visually inspecting these derivation trees is advantageous in debugging many
+grammars.
+
+We would like to add visualization tools for more parts of the pipeline. For
+example, a chart visualizer would make it easier for researchers to tell where
+search errors were happening during decoding, and why. An alignment visualizer
+for aligned parallel corpora might help to determine how grammar extraction 
+could be improved.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
add semiring parsing
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1608 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 Chris Callison-Burch,\,\,\,
-%Chris Dyer,$^\dagger$\,\,\,
+Chris Dyer,$^\dagger$\,\,\,
 Juri Ganitkevitch,\,\,\,
+\\ {\bf
 Ann Irvine,\,\,\,  
-%Sanjeev Khudanpur,\,\,\,
+Sanjeev Khudanpur,\,\,\,
 Lane Schwartz,$^\star$\,\,\, 
+Wren N.\,G.\,Thornton,\,\,\,}
 \\ {\bf
-Wren N.\,G.\,Thornton,\,\,\,
 Ziyuan Wang,\,\,
@@
 {\textnormal{and}}
-\,\,\,Omar F. Zaidan}\\
+\,\,\,Omar F. Zaidan
+}\\
 Center for Language and Speech Processing, Johns Hopkins University, Baltimore, MD\\
-%$\dagger$ Computational Linguistics and Information Processing Lab, University of Maryland, College Park, MD\\
+$\dagger$ Computational Linguistics and Information Processing Lab, University of Maryland, College Park, MD\\
 $\star$ Natural Language Processing Lab, University of Minnesota, Minneapolis, MN }
@@
 \section{Semiring Parsing}
-\todo{zhifei}
 
-\cite{li-eisner:2009:EMNLP}
+In \joshua, we use a {\em hypergraph} (or {\em packed forest}) to compactly 
+represent the exponentially many derivation trees generated by the decoder for a 
+particular foreign input sentence.
+Given a hypergraph, we may perform many atomic inference operations
+such as finding one-best, $k$-best, or expectations over the hypergraph.
+To perform each operation, we may implement a dedicated dynamic programming algorithm.
+However, a more general framework to specify these algorithms is semiring-weighted parsing \cite{semiringparsing}.
+Within this framework, we implemented all the algorithms (i.e.,
+inside, outside, and inside-outside speedup) described by \newcite{li-eisner:2009:EMNLP},
+and also the first-order expectation semiring \cite{eisner-expectation-semiring} and its
+second-order version \cite{li-eisner:2009:EMNLP}.
 
-\section{Discriminative Training}
+% The original first-order expectation semiring allows us to efficiently
+% compute a vector of first-order statistics (expectations; first
+% derivatives) on the set of paths in a lattice or the set of trees in a
+% hypergraph.  The second-order expectation semiring {\em additionally} computes
+% a matrix of second-order statistics (expectations of {\em products}; second
+% derivatives (Hessian); derivatives of expectations).
+
+%% use of expectation and variance semiring
+The first- and second-order expectation semirings can be used compute many interesting quantities over the hypergraph.
+These quantities include expected hypothesis length, feature expectation, entropy, cross-entropy, Kullback-Leibler divergence,
+Bayes risk, variance of hypothesis length, gradient of entropy and Bayes risk, covariance and Hessian matrix, and so on.
+% The second-order expectation semiring is essential for many interesting training paradigms 
+% such as deterministic annealing \cite{Rose98deterministicannealing}, minimum risk \cite{smith-eisner:2006:POS},
+% active and semi-supervised learning \cite{entropy-mini,semiCRF-jiao}.
+% In these settings, we must compute the gradient of entropy or risk. 
+% The semiring can also be used for second-order gradient optimization algorithms.
+
+\section{Hypergraph-based Discriminative Training}
 \todo{zhifei}

### 2010-04-23 papers/wmt10/joshua-progress.tex
add hook for lattice parsing
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1609 0ae5e6b2-d358-4f09-a895-f82f13dd62a4
Added a clean bibliography
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1610 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\title{Joshua 2.0: An Open Source Toolkit for Machine Translation \\with Syntax, Semirings and Other Goodies }
+\title{Joshua 2.0: A Toolkit for Parsing-Based Machine Translation \\with Syntax, Semirings and Other Goodies }
 
@@
 In this report, we will describe the main new functions we have added into the 
-toolkit, which include: supporting for the syntax-augmented machine translation (SAMT) grammar \cite{???},
-supporting mannual constraints in the input to constrain the decoding \cite{???},  
+toolkit, which include: supporting for the syntax-augmented machine translation (SAMT) grammar \cite{samt2006},
+supporting manual constraints in the input to constrain the decoding \cite{PBML-2010-Josua-transliteration},  
 semiring parsing for unified dynamic programming \cite{li-eisner:2009:EMNLP}, 
 hypergraph-based discriminative training for better feature engineering \cite{oracle-extraction-naacl09}, 
-visualization of the derivation trees and hypergraphs \cite{???}, 
-and a unified pipeline for MT experiments \cite{???}.
+visualization of the derivation trees and hypergraphs \cite{PBML-2010-Josua-visualization}, 
+and a unified pipeline for MT experiments \cite{Schwartz-wmt10-pipline}.
 
@@
 \todo{anni}
+
+\section{Lattice Parsing}
+\todo{Chris Dyer}
 
@@
 We have introduced tools for visualizing two of the main data structures used
-in Joshua \cite{joshua-viz}. Each visualization tool is implemented as a
+in Joshua \cite{PBML-2010-Josua-visualization}. Each visualization tool is implemented as a
 stand-alone program that hooks into the pipeline, but they are distributed with
@@
 \bibliographystyle{acl}
-\bibliography{machinetranslation}
+\bibliography{machinetranslation-clean}
 

### 2010-04-23 papers/wmt10/scratch.tex
Added a clean bibliography
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1610 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

--- 2010-04-22 papers/wmt10.pipeline/paper.tex
@@
+%
+% File acl2010.tex
+%
+% Contact  jshin@csie.ncnu.edu.tw or pkoehn@inf.ed.ac.uk
+%%
+%% Based on the style files for ACL-IJCNLP-2009, which were, in turn,
+%% based on the style files for EACL-2009 and IJCNLP-2008...
+
+%% Based on the style files for EACL 2006 by 
+%%e.agirre@ehu.es or Sergi.Balari@uab.es
+%% and that of ACL 08 by Joakim Nivre and Noah Smith
+
 \documentclass[11pt]{article}
+\usepackage{acl2010}
 \usepackage{times}
@@
 \usepackage{latexsym}
-\usepackage{tikz}
-\usetikzlibrary{fit,shapes.misc,shapes}
-\usepackage{acl2010}
-%\usepackage[colorinlistoftodos]{todonotes}
+\usepackage[colorlinks]{hyperref}
+\usepackage[colorinlistoftodos]{todonotes}
+%\setlength\titlebox{6.5cm}    % You can expand the title box if you
+% really have to
 
-\title{Reproducible Results in Parsing-Based Machine Translation:\\
-The JHU Shared Task Submission}
+\title{Joshua: An Open Source Toolkit for Parsing-based Machine Translation}
 
-\author{Lane Schwartz\
-\thanks{\ Research conducted as a visiting researcher at Johns Hopkins University}\
-\\University of Minnesota\\Minneapolis, MN\\
-{\tt lane@cs.umn.edu} \And
-Chris Callison-Burch \\
-Johns Hopkins University \\
-Baltimore, MD\\
-{\tt ccb@cs.jhu.edu}}
+\author{
+Zhifei Li,\,\,\,
+ Lane Schwartz,$^\star$\,\,\,
+Chris Callison-Burch,\,\,\, %add a footnote to mention that CCB is the project leader
+Chris Dyer,$^\dagger$\,\,\,
+Juri Ganitkevitch,\,\,\, \\
+{\bf Sanjeev Khudanpur,\,\,\, 
+Wren N.\,G.\,Thornton,\,\,\,
+Jonathan Weese\,\,
+{\textnormal{and}}\,\,\,Omar F. Zaidan}\\
+Center for Language and Speech Processing, Johns Hopkins University, Baltimore, MD\\
+$\dagger$ Computational Linguistics and Information Processing Lab, University of Maryland, College Park, MD\\
+$\star$ Natural Language Processing Lab, University of Minnesota, Minneapolis, MN }
 
@@
 \begin{abstract}
+ This paper describes Joshua, an open source hierarchical phrase-based statistical machine translation system. Results for the WMT 2010 shared task are reported for six language pairs.
 \end{abstract}
 
-\section{Introduction}
+\listoftodos
 
-Research investigating natural language process and computational linguistics can and should have an extremely low barrier to entry.
-The data with which we work is customarily available in common electronic formats.
-The computational techniques which we apply can typically be performed on commodity computing resources which are widely available.
-In short, there should be no reason why small research groups and even lone researchers should not be able to join and make substantive contributions furthering our field.
-%
-The reality is less encouraging.
+\section{Grammar Extraction}
 
-Many published articles describe novel techniques and provide interesting results, yet fail to describe technical details in sufficient detail to allow their results to be reproduced by other researchers.
-While there are notable and laudable exceptions, many publications fail to provide the source code and scripts necessary to reproduce results.
-The use of restricted data, not freely available for download by any interested researcher only compounds these problems.
-\newcite{pedersen2010} rightly argues that the implementation details so often ignored in publications are in fact essential for our research to be reproducible science.
+\todo{Add info about grammar extraction; MT Marathon paper and speedups}
 
-Reproducibility in machine translation is made more challenging by the complexity of experimental workflows.
-Results in machine translation tasks are dependent on a cascade of processing steps and configurations.
-While interesting subsets of these usually appear in experimental descriptions, many steps (preprocessing techniques, alignment parameters, translation rule extraction parameters, language model parameters, list of features used) are invariably omitted, even though these configurations are often critical to reproducing results.
+\section{Results}
 
-This paper describes the Johns Hopkins University submission to the 2010 Workshop on Statistical Machine Translation shared translation task.
-Links to the software, scripts, and configurations used to run the experiments described herein are provided.
-The remainder of this paper is structured as follows.
-Section \ref{Related Work} lists the major examples of publicly available open source machine translation systems, parallel corpora, and machine translation workflow management systems.
-Section \ref{Framework} describes the experimental workflow used to run the shared task translations, with the corresponding experimental design in section \ref{Design}.
-Section \ref{Results} presents the shared task results.
-
-\begin{figure*}[htbp]
-\include{figure.pgf}
-\caption{Machine translation workflow.
-Nodes in grey indicate software and scripts.
-The scripts and configuration files used to implement and run this workflow are available for download at \url{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz}}
-\label{dependencyGraph}
-\end{figure*}
-
-\section{Related Work}
-\label{Related Work}
-
-The last four years have witnessed the implementation and release of numerous open source machine translation systems.
-The widely used Moses system \cite{moses} implements the standard phrase-based translation model.
-Parsing-based translation models are implemented by Joshua \cite{Joshua-WMT}, SAMT \cite{samt2006}, and cdec \cite{cdec}.
-Cunei \cite{Phillips2009} implements statistical example-based translation.
-\newcite{Phramer} and \newcite{schwartz08mclc} respectively provide additional open-source implementations of phrase-based and hierarchical decoders.
-
-The SRILM \cite{Stolcke2002}, IRSTLM \cite{IRSTLM}, and RandLM \cite{Talbot2007a} toolkits enable efficient training and querying of n-gram language models.
-
-Freely available parallel corpora for numerous European languages have also been released in recent years.
-These include the Europarl \cite{Koehn-europarl} and JRC-Acquis \cite{Steinberger-2006} legislative corpora, each of which includes data for most EU language pairs.
-The smaller News Commentary \cite{Callison-Burch2007a,Callison-Burch2008a} provide smaller amounts of parallel data in the news genre.
-The recent Fr-En $10^9$ \cite{WMT09-Findings} corpus aggregates huge numbers of parallel French-English sentences from the web.
-
-Open source systems to address the complex workflows required to run non-trivial machine translation experiments have also been developed.
-These include {\tt experiment.perl} \cite{experiment.perl}, developed as a workflow management system at the University of Edinburgh, and LoonyBin \cite{LoonyBin}, a general hyperworkflow management utility from Carnegie Melon University.
-
-\section{Experimental Workflow Management}
-\label{Framework}
-
-Running a statistical machine translation system to achieve state-of-the-art performance involves the configuration and execution of numerous interdependent intermediate tools.
-To manage task dependencies and tool configuration, our shared task workflow consists of a set of dependency scripts written for GNU Make \cite{gnumake}.
-
-Figure \ref{dependencyGraph} shows a graph depicting the steps in our experimental workflow, and the dependencies between steps.
-Each node in the graph represents a step in the workflow; each step is with a Make script that defines how to run the tools required in that step.
-In each experiment, an additional configuration script is provided for each experimental step, defining the parameters to be used when running that step in the current experiment.
-Optional front-end wrapper scripts can also be provided, allowing for a complete experiment to be run - from downloading data and software through truecasing translated results - by executing a single make file.
-
-This framework is also conducive to parallelization.
-Many tasks, such as preprocessing numerous training files, are not dependent on one another.
-In such cases {\tt make} can be configured to execute multiple processes simultaneously on a single multi-processor machine.
-In cases where scheduled distributed computing environments such as the Sun Grid Engine are configured, make files can be processed by scheduler-aware {\tt make} variants ({\tt distmake}, SGE {\tt qmake}, Sun Studio {\tt dmake}) which distribute outstanding tasks to available distributed machines using the relevant distributed scheduler.
-
-\section{Experimental Configuration}
-\label{Design}
-
-\begin{table*}[htb]
-\begin{center}
-\begin{tabular}{|l|l|l|}
-\hline
-\bf Source & \bf Target & \bf Parallel Corpora \\ \hline
-German & English & news-commentary10.de-en europarl-v5.de-en \\ \hline
-English & German & news-commentary10.de-en europarl-v5.de-en \\ \hline
-French & English & news-commentary10.fr-en europarl-v5.fr-en giga-fren.release2 undoc.2000.en-fr \\ \hline
-English & French & news-commentary10.fr-en europarl-v5.fr-en giga-fren.release2 undoc.2000.en-fr \\ \hline
-Spanish & English & news-commentary10.es-en europarl-v5.es-en undoc.2000.en-es \\ \hline
-English & Spanish & news-commentary10.es-en europarl-v5.es-en undoc.2000.en-es  \\ \hline
-\end{tabular}
-\end{center}
-\caption{Parallel training data used for training translation model, per language pair}
-\label{parallelCorpora}
-\end{table*}
-
-\begin{table*}[htb]
-\begin{center}
-\begin{tabular}{|l|l|}
-\hline
-\bf Target & \bf Monolingual Corpora \\ \hline
-English & europarl-v5.en news-commentary10.en news.en.shuffled undoc.2000.en-fr.en giga-fren.release2.en \\ \hline
-French & europarl-v5.fr news-commentary10.fr news.fr.shuffled undoc.2000.en-fr.fr giga-fren.release2.fr \\ \hline
-German & europarl-v5.de news-commentary10.de news.de.shuffled \\ \hline
-Spanish & europarl-v5.es news-commentary10.es news.es.shuffled undoc.2000.en-es.es\\ \hline
-\end{tabular}
-\end{center}
-\caption{Monolingual training data used for training language model, per target language}
-\label{monolingualCorpora}
-\end{table*}
-
-Experimental workflows were configured\footnote{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz} and run for six language pairs in the translation shared task: English-French, English-German, English-Spanish, French-English, German-English, and Spanish-English.
-
-In all experiments, only data freely available for download was used.
-No restricted data from the LDC or other sources was used.
-Table \ref{parallelCorpora} lists the parallel corpora used in training the translation model for each experiment.
-The monolingual corpora used in training each target language model are listed in table \ref{monolingualCorpora}.
-In all experiments, news-test2008 was used as a development tuning corpus during minimum error rate training; newstest2009 was used as a development test set.
-The shared task data set newstest2010 was used as a final blind test set.
-
-All data was automatically downloaded, unzipped, and preprocessed prior to use.
-Files provided in XML format were converted to plain text by selecting lines with {\tt <seg>} tags, then removing the beginning and end tags for each segment; this processing was applied using GNU grep and sed.
-The {\tt tokenize.perl} and {\tt lowercase.perl} scripts provided for the shared task\footnote{http://www.statmt.org/wmt08/scripts.tgz with md5sum: \\ {\tt tokenize.perl} 45cd1832827131013245eca76481441a \\ {\tt lowercase.perl} a1958ab429b1e29d379063c3b9cd7062} were applied to all data.
-
-Interpolated n-gram language models for the four target languages were built using the SRI Language Model Toolkit\footnote{ http://www-speech.sri.com/projects/srilm \\ SRILM version 1.5.7.
-Our experimental workflow requires that SRILM be compiled separately, with the \$SRILM environment variable set to the install location.}, with n-gram order set to 5.
-The \newcite{ChenGoodman98} technique for modified Kneser-Ney discounting \cite{Kneser1995} was applied during language model training.
-
-Following \newcite{Joshua-WMT}, a subset of the available training sentences is selected via subsampling; training sentences are selected based on the estimated likelihood of that sentence being useful later for translating a particular test corpus.
-
-Given a subsampled parallel training corpus, word alignment is performed using the Berkeley aligner\footnote{http://berkeleyaligner.googlecode.com/files/berkeleyaligner\\\_unsupervised-2.1.tar.gz --- Berkeley aligner version 2.1} \cite{liang06alignment}.
-
-For each language pair, a synchronous context free translation grammar is extracted for a particular test set, following the methods of \newcite{Lopez2008} as implemented in \cite{Schwartz-PBML}.
-For the largest training sets (French-English and English-French) the original \cite{Lopez2008} implementation included with Hiero was used to save time during training\footnote{It is expected that using the Joshua implementation should result in nearly identical results, albeit with somewhat more time required to extract the grammar.}.
-
-Our experiments used hierarchical phrase-based grammars containing exactly two nonterminals - the wildcard nonterminal X, and S, used to glue together neighboring constituents.
-Recent work has shown that parsing-based machine translation using SAMT \cite{samt2006} grammars with rich nonterminal sets can demonstrate substantial gains over hierarchical grammars for certain language pairs \cite{SCALE-report}.
-Joshua supports such grammars; the experimental workflow presented here could easily be extended in future research to incorporate the use of SAMT grammars with additional language pairs.
-The Z-MERT implementation \cite{Zaidan2009} of minimum error rate training was used for parameter tuning.
-Tuned grammars were used by Joshua to translate all test sets.
-
-The code used for subsampling, grammar extraction, decoding, and minimum error rate training is provided with Joshua\footnote{http://sourceforge.net/projects/joshua/files/joshua/1.3/joshua-1.3.tgz/download --- Joshua version 1.3}, with the exception of the original \cite{Lopez2008} grammar extraction code.
-
-\section{Reproducible Results}
-\label{Results}
+\todo{Describe results}
 
@@
 
-\section*{Acknowledgements}
-This work was supported by the DARPA GALE program (Contract No HR0011-06-2-0001).
-
-\bibliographystyle{acl}
-\bibliography{bibliography}
-
 \end{document}

### 2010-04-23 papers/wmt10/joshua-progress.tex
Pointed the tex to the clean bibliography
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1613 0ae5e6b2-d358-4f09-a895-f82f13dd62a4
add variational hook
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1614 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\title{Joshua 2.0: A Toolkit for Parsing-Based Machine Translation \\with Syntax, Semirings and Other Goodies }
+\title{Joshua 2.0: A Toolkit for Parsing-Based Machine Translation \\with Syntax, Semirings, Discriminative Training and Other Goodies }
 
@@
 Bayes risk, variance of hypothesis length, gradient of entropy and Bayes risk, covariance and Hessian matrix, and so on.
+
 % The second-order expectation semiring is essential for many interesting training paradigms 
@@
 
+\section{Variational Decoding}
+Statistical models in machine translation exhibit spurious ambiguity.
+That is, the probability of an output string is split
+among many distinct derivations (e.g., trees or
+segmentations). In principle, the goodness of a
+string is measured by the total probability of its
+many derivations. However, finding the best string
+(e.g., during decoding) is then computationally intractable.
+Therefore, the first version implemented the Viterbi ?????Joshua most systems use a simple
+Viterbi approximation that measures the goodness 
+of a string using only its most probable derivation.
+
+Instead, we develop a variational approximation \cite{variational-decoding-acl09},
+which considers all the derivations but still
+allows tractable decoding.
+
 \section{Hypergraph-based Discriminative Training}
-\todo{zhifei}
+Discriminative training with a large number of features has 
+great potential to improve the MT performance.
+We have implemented the hypergraph-based
 
-\cite{variational-decoding-acl09}
 \cite{oracle-extraction-naacl09}
@@
 \bibliographystyle{acl}
-\bibliography{machinetranslation-clean}
+\bibliography{machinetranslation}
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
done with variational decoding
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1615 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \usepackage[colorinlistoftodos]{todonotes}
+\usepackage{xspace}
+
 %\setlength\titlebox{6.5cm}    % You can expand the title box if you
@@
 
-\newcommand{\joshua}{\textbf{Joshua}}
+\newcommand{\joshua}{\textbf{Joshua}\xspace}
+\newcommand{\Q}{\mathcal{Q}}
 
@@
 \section{Variational Decoding}
+
 Statistical models in machine translation exhibit spurious ambiguity.
@@
 many derivations. However, finding the best string
-(e.g., during decoding) is then computationally intractable.
-Therefore, the first version implemented the Viterbi ?????Joshua most systems use a simple
+(e.g., during decoding) is then NP-hard.
+The first version of \joshua implemented the
 Viterbi approximation that measures the goodness 
@@
 
-Instead, we develop a variational approximation \cite{variational-decoding-acl09},
-which considers all the derivations but still
+The Viterbi approximation is efficient, but ignores most of the derivations in the hypergraph.
+We implemented the variational decoding \cite{variational-decoding-acl09}, which works as follows.
+Given a foreign string, the MT system first produces a hypergraph, 
+which encodes a probability distribution $p$ over 
+possible output strings and their derivations.
+The variational decoding constructs a second distribution $q \in \Q$ 
+that approximates $p$ as well as possible, 
+and then finds the best string according to $q$ (instead of $p$).
+The $q$ distribution is parameterized by an $n$-gram model, under which 
+the last two steps can be performed efficiently via dynamic programming.
+In this way, variational decoding considers all the derivations but still
 allows tractable decoding.

### 2010-04-23 papers/wmt10/joshua-progress.tex
finished with training
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1616 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \section{Hypergraph-based Discriminative Training}
+
 Discriminative training with a large number of features has 
-great potential to improve the MT performance.
-We have implemented the hypergraph-based
+potential to improve the MT performance.
+We have implemented the hypergraph-based minimum risk training \cite{li-eisner:2009:EMNLP},
+which minimizes the {\em expected loss} of the reference translations.
+The minimum-risk objective can be optimized by a gradient-based method, where
+the risk and its gradient can be computed using a second-order expectation semiring.
+For optimization, we use both L-BFGS\footnote{http://en.wikipedia.org/wiki/L-BFGS} 
+and Rprop\footnote{http://en.wikipedia.org/wiki/Rprop}.
 
-\cite{oracle-extraction-naacl09}
+We have also implemented the average Peceptron algorithm and the forest-reranking \cite{zhifei-forest-reranking-galebook}.
+Since the reference translation may not be in the hypergraph due to pruning or inherent
+defficiency of the translation grammar, we need to use an {\em oracle translation} (i.e., the translation in
+the hypergraph that is most simmilar to the reference translation) as a surrogate for training.
+We implemented the {\em oracle extraction} algorithm described by \newcite{oracle-extraction-naacl09}
+for this purpose.
+
+Given the current infrastructure, other training methods 
+(e.g., maximum conditional likelihood or MIRA as used by \newcite{chiang-knight-wang:2009:NAACLHLT09})
+can also be easily supported with minimum coding.
+We plan to implement a large number of feature functions in \joshua so that exhaustive 
+feature engineering is possible for MT.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
zhifei is done
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1617 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 an open-source toolkit for parsing-based machine translation.
-The main new functions we have added into the toolkit include: supporting for the syntax-augmented machine translation (SAMT) grammar,
+The main new functions we have added into the toolkit include: 
+supporting for the syntax-augmented machine translation (SAMT) grammar,
 supporting mannual constraints in the input to constrain the decoding,  
-semiring parsing for unified dynamic programming, hypergraph-based 
-discriminative training for better feature engineering, visualization of the derivation
-trees and hypergraphs, and a unified pipeline for MT experiments.
+lattice parsing for dealing with input uncertainty,
+semiring parsing for unified dynamic programming, 
+variational decoding for approximating the intractable MAP decoding,
+hypergraph-based discriminative training for better feature engineering, 
+visualization of the derivation trees and hypergraphs, 
+and a unified pipeline for MT experiments.
 \end{abstract}
@@
 in \newcite{hiero-2007}: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.  
-The initial release also implemented suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007} 
+It also implemented suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007} 
 and minimum error rate training \cite{och-mert}. 
@@
 been made to the first version. 
-In this report, we will describe the main new functions we have added into the 
-toolkit, which include: supporting for the syntax-augmented machine translation (SAMT) grammar \cite{samt2006},
-supporting manual constraints in the input to constrain the decoding \cite{PBML-2010-Josua-transliteration},  
-semiring parsing for unified dynamic programming \cite{li-eisner:2009:EMNLP}, 
-hypergraph-based discriminative training for better feature engineering \cite{oracle-extraction-naacl09}, 
+In this paper, we will describe the main new functions we have added into the 
+toolkit, which include: 
+\begin{itemize}
+\item
+supporting for the syntax-augmented machine translation (SAMT) grammar \cite{samt2006},
+\item
+supporting mannual constraints in the input to constrain the decoding \cite{PBML-2010-Josua-transliteration},  
+\item
+lattice parsing for dealing with input uncertainty \cite{dyer-muresan-resnik:2008:ACLMain},
+\item
+semiring parsing for unified dynamic programming \cite{li-eisner:2009:EMNLP} , 
+\item
+variational decoding for approximating the intractable MAP decoding \cite{variational-decoding-acl09},
+\item
+hypergraph-based discriminative training for better feature engineering \cite{zhifei-forest-reranking-galebook}, 
+\item
 visualization of the derivation trees and hypergraphs \cite{PBML-2010-Josua-visualization}, 
+\item
 and a unified pipeline for MT experiments \cite{Schwartz-wmt10-pipline}.
-
+\end{itemize}
 Below we give a short description for each of the new functions added.
@@
 
-We have also implemented the average Peceptron algorithm and the forest-reranking \cite{zhifei-forest-reranking-galebook}.
+We have also implemented the average Perceptron algorithm and the forest-reranking \cite{zhifei-forest-reranking-galebook}.
 Since the reference translation may not be in the hypergraph due to pruning or inherent

### 2010-04-23 papers/wmt10.pipeline/paper.tex
working on figure
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1603 0ae5e6b2-d358-4f09-a895-f82f13dd62a4
Possibly final draft
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1618 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
-\documentclass[11pt]{article}
+\documentclass[11pt, a4paper]{article}
 \usepackage{times}
@@
 \usepackage{acl2010}
+%\usepackage[pdftex]{graphicx}
 %\usepackage[colorinlistoftodos]{todonotes}
@@
 \caption{Machine translation workflow.
-Nodes in grey indicate software and scripts.
+Square nodes in grey indicate software and scripts.
 The scripts and configuration files used to implement and run this workflow are available for download at \url{http://sourceforge.net/projects/joshua/files/joshua/1.3/wmt2010-experiment.tgz}}
@@
 
-The Z-MERT implementation \cite{Zaidan2009} of minimum error rate training \cite{Och2003c} was used for parameter tuning.
+The Z-MERT implementation \cite{Zaidan2009} of minimum error rate training \cite{Och2003c} was used for parameter tuning\footnote{Z-MERT seed value was set to 12341234}.
 Tuned grammars were used by Joshua to translate all test sets.

### 2010-04-23 papers/wmt10/joshua-progress.tex
Minor fixes, typos, etc
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1619 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \usepackage{latexsym}
-\usepackage[colorlinks]{hyperref}
+%\usepackage[colorlinks]{hyperref}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 
-The toolkit has then been used by the authors and several
+The toolkit has been used by the authors and several
 other groups in their daily research, and thus much progress has
 been made to the first version. 
-In this paper, we will describe the main new functions we have added into the 
+In this paper, we describe the main new functions we have added into the 
 toolkit, which include: 
@@
 \item
-supporting for the syntax-augmented machine translation (SAMT) grammar \cite{samt2006},
+supporting for syntax-augmented machine translation (SAMT) grammar \cite{samt2006},
 \item
-supporting mannual constraints in the input to constrain the decoding \cite{PBML-2010-Josua-transliteration},  
+supporting manual constraints in the input to constrain the decoding \cite{PBML-2010-Josua-transliteration},  
 \item
@@
 \item
-and a unified pipeline for MT experiments \cite{Schwartz-wmt10-pipline}.
+and a unified framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}.
 \end{itemize}

### 2010-04-23 papers/wmt10/joshua-progress.tex
Rewording manual constraints
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1620 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 supporting for the syntax-augmented machine translation (SAMT) grammar,
-supporting mannual constraints in the input to constrain the decoding,  
+support for external modules to posit translations for spans in the input sentence that constrain decoding,  
 lattice parsing for dealing with input uncertainty,
@@
 \item
-supporting for syntax-augmented machine translation (SAMT) grammar \cite{samt2006},
+support for the syntax-augmented machine translation (SAMT) grammar \cite{samt2006},
 \item
-supporting manual constraints in the input to constrain the decoding \cite{PBML-2010-Josua-transliteration},  
+support for external modules to posit translations for spans in the input sentence that constrain decoding,    \cite{PBML-2010-Josua-transliteration},  
 \item
@@
 
-\section{Manual Constraints on the Input}
+\section{Specifying Constraints on the Rules}
 \todo{anni}

### 2010-04-23 papers/wmt10/joshua-progress.tex
Lane's section added
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1621 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \section{MT Pipeline}
-\todo{lane}
+
+Many research publications describe novel techniques and provide interesting results, yet fail to describe technical details in sufficient detail to allow their results to be reproduced by other researchers.
+Reproducibility in machine translation is made more challenging by the complexity of experimental workflows.
+Results in machine translation tasks are dependent on a cascade of processing steps and configurations.
+
+\newcite{Schwartz-wmt10-pipline} describes a workflow framework for designing and running reproducible machine translation experiments using Joshua.
+Each step in the machine translation workflow (data preprocessing, grammar training, MERT, decoding, etc) is modeled by a GNU Make script which defines how to run the tools used in that step, and an auxiliary configuration file that defines the exact parameters to be used in that step during a particular run of an experiment.
+Workflows configured using this framework allow a complete experiment to be run - from downloading data and software through truecasing translated results - by executing a single make file.
+
+This framework encourages researchers to supplement research publications with links to the complete set of scripts and configurations that were actually used to run the experiment.
+The scripts and configurations used to run the Johns Hopkins University submission to the 2010 WMT shared translation task are provided in \newcite{Schwartz-wmt10-pipline}.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited abstract
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1622 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \begin{abstract}
-We describe the progress we have made in the past year to \textbf{Joshua} \cite{joshua-wmt09}, 
-an open-source toolkit for parsing-based machine translation.
-The main new functions we have added into the toolkit include: 
-supporting for the syntax-augmented machine translation (SAMT) grammar,
-support for external modules to posit translations for spans in the input sentence that constrain decoding,  
+We describe the progress we have made in the past year on \textbf{Joshua} \cite{joshua-wmt09}, 
+an open source toolkit for parsing based machine translation.
+The new functionality includes: 
+support for translation grammars with a rich set of syntactic nonterminals,
+the ability for external modules to posit constraints on spans in the input sentence should be translated,  
 lattice parsing for dealing with input uncertainty,
-semiring parsing for unified dynamic programming, 
+a semiring framework that provides a unified way of doing various dynamic programming calculations, 
 variational decoding for approximating the intractable MAP decoding,
 hypergraph-based discriminative training for better feature engineering, 
-visualization of the derivation trees and hypergraphs, 
-and a unified pipeline for MT experiments.
+visualization of the derivation trees, 
+and a cleaner pipeline for MT experiments.
 \end{abstract}
-
-\listoftodos
 
@@
 
-\textbf{Joshua} \cite{joshua-wmt09} is an open-source toolkit (written in Java) for parsing-based machine translation.
+\textbf{Joshua} \cite{joshua-wmt09} is an open-source toolkit for parsing-based machine translation (written in Java). 
 The first version implemented all the essential algorithms described 
@@
 
-\section{Specifying Constraints on the Rules}
+\section{Specifying Constraints on Rules}
 \todo{anni}

### 2010-04-23 papers/wmt10/joshua-progress.tex
dyer's edits
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1623 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\section{Lattice Parsing}
-\todo{Chris Dyer}
-
 \section{Semiring Parsing}
@@
 
+\section{Lattice Parsing}
+
+The bottom-up parsing algorithm used to generate the translation hypergraph has been generalized to support translation from not only an input sentence but from a \emph{word lattice} with runtime and memory overhead that is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.
+This enables the exploration of a distribution over input sentences during decoding, and the best translation is selected from among any of them.
+This functionality is useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or word segmenters.
+
 \section{Variational Decoding}
@@
 many derivations. However, finding the best string
-(e.g., during decoding) is then NP-hard.
+during decoding is then NP-hard.
 The first version of \joshua implemented the
@@
 
-The Viterbi approximation is efficient, but ignores most of the derivations in the hypergraph.
-We implemented the variational decoding \cite{variational-decoding-acl09}, which works as follows.
-Given a foreign string, the MT system first produces a hypergraph, 
+The Viterbi approximation is efficient, but it ignores most of the derivations in the hypergraph.
+We implemented variational decoding \cite{variational-decoding-acl09}, which works as follows.
+First, given a foreign string (or lattice), the MT system produces a hypergraph, 
 which encodes a probability distribution $p$ over 
 possible output strings and their derivations.
-The variational decoding constructs a second distribution $q \in \Q$ 
-that approximates $p$ as well as possible, 
-and then finds the best string according to $q$ (instead of $p$).
-The $q$ distribution is parameterized by an $n$-gram model, under which 
-the last two steps can be performed efficiently via dynamic programming.
-In this way, variational decoding considers all the derivations but still
+Second, a distribution $q$ is selected that approximates $p$ as well as possible but comes from a family of distributions $\mathcal{Q}$ where inference is tractable.
+Third, the best string according to $q$ (instead of $p$) is found.
+In our implementation, the $q$ distribution is parameterized by an $n$-gram model, under which 
+the second and third steps can be performed efficiently and exactly via dynamic programming.
+In this way, variational decoding considers all derivations in the hypergraph but still
 allows tractable decoding.
@@
 
-We have also implemented the average Perceptron algorithm and the forest-reranking \cite{zhifei-forest-reranking-galebook}.
+We have also implemented the average Perceptron algorithm and forest-reranking \cite{zhifei-forest-reranking-galebook}.
 Since the reference translation may not be in the hypergraph due to pruning or inherent

### 2010-04-23 papers/wmt10/joshua-progress.tex
dyer's edits
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1624 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 possible output strings and their derivations.
-Second, a distribution $q$ is selected that approximates $p$ as well as possible but comes from a family of distributions $\mathcal{Q}$ where inference is tractable.
+Second, a distribution $q$ is selected that approximates $p$ as well as possible but comes from a family of distributions $\mathcal{Q}$ in which inference is tractable.
 Third, the best string according to $q$ (instead of $p$) is found.

### 2010-04-23 papers/wmt10/joshua-progress.tex
Added two refs to our PBML pubs
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1625 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\textbf{Joshua} \cite{joshua-wmt09} is an open-source toolkit for parsing-based machine translation (written in Java). 
-The first version implemented all the essential algorithms described 
+\textbf{Joshua} is an open-source toolkit for parsing-based machine translation (written in Java). 
+The initial release of Joshua  \cite{joshua-wmt09} implemented all the essential algorithms described 
 in \newcite{hiero-2007}: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.  
-It also implemented suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007} 
-and minimum error rate training \cite{och-mert}. 
+It also implemented suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction} 
+and minimum error rate training \cite{och-mert,Zaidan2009}. 
 Additionally, parallel and distributed computing techniques were exploited 
@@
 \item
-support for external modules to posit translations for spans in the input sentence that constrain decoding,    \cite{PBML-2010-Josua-transliteration},  
+support for external modules to posit translations for spans in the input sentence that constrain decoding,    \cite{PBML-2010-Joshua-transliteration},  
 \item
@@
 \item
-visualization of the derivation trees and hypergraphs \cite{PBML-2010-Josua-visualization}, 
+visualization of the derivation trees and hypergraphs \cite{PBML-2010-Joshua-visualization}, 
 \item
@@
 We have introduced tools for visualizing two of the main data structures used
-in Joshua \cite{PBML-2010-Josua-visualization}. Each visualization tool is implemented as a
+in Joshua \cite{PBML-2010-Joshua-visualization}. Each visualization tool is implemented as a
 stand-alone program that hooks into the pipeline, but they are distributed with

### 2010-04-23 papers/wmt10/joshua-progress.tex
Revising intro
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1626 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \textbf{Joshua} is an open-source toolkit for parsing-based machine translation (written in Java).
-The initial release of Joshua  \cite{joshua-wmt09} implemented all the essential algorithms described 
-in \newcite{hiero-2007}: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.  
-It also implemented suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction} 
-and minimum error rate training \cite{och-mert,Zaidan2009}. 
-Additionally, parallel and distributed computing techniques were exploited 
-to make it scalable \cite{Joshua-old}. In the design of \joshua, we have applied general
-principles of software engineering to achieve three
-major goals: extensibility, end-to-end coherence,
-and scalability.
+The initial release of Joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.
+The Joshua release also included re-implementations of suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.
+Additionally, it included parallel and distributed computing techniques for salability \cite{Joshua-old}.
 
-The toolkit has been used by the authors and several
-other groups in their daily research, and thus much progress has
-been made to the first version. 
-In this paper, we describe the main new functions we have added into the 
-toolkit, which include: 
+This paper describes the additions to the toolkit over the past year, which together form the 2.0 release.
+The software has been heavily used by the authors and several other groups in their daily research, and has been substantially refined since the first release.
+The most important new functions in the toolkit are:
 \begin{itemize}

### 2010-04-23 papers/wmt10/joshua-progress.tex
Revising intro
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1627 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\textbf{Joshua} is an open-source toolkit for parsing-based machine translation (written in Java).
+\textbf{Joshua} is an open-source toolkit for parsing-based machine translation that is written in Java.
 The initial release of Joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.
-The Joshua release also included re-implementations of suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.
+The Joshua 1.0 release also included re-implementations of suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.
 Additionally, it included parallel and distributed computing techniques for salability \cite{Joshua-old}.
@@
 \item
-support for the syntax-augmented machine translation (SAMT) grammar \cite{samt2006},
+Support for any style of synchronous context free grammar (SCFG)
+% instead of just Hiero grammars.  This 
+including syntax augment machine translation (SAMT) grammars \cite{samt2006},
 \item
-support for external modules to posit translations for spans in the input sentence that constrain decoding,    \cite{PBML-2010-Joshua-transliteration},  
+Support for external modules to posit translations for spans in the input sentence that constrain decoding,    \cite{PBML-2010-Joshua-transliteration},  
 \item
-lattice parsing for dealing with input uncertainty \cite{dyer-muresan-resnik:2008:ACLMain},
+Lattice parsing for dealing with input uncertainty, including ambiguous output from speech recognizers or Chinese word segmenters  \cite{dyer-muresan-resnik:2008:ACLMain},
 \item
-semiring parsing for unified dynamic programming \cite{li-eisner:2009:EMNLP} , 
+Semiring parsing for unified dynamic programming \cite{li-eisner:2009:EMNLP} , 
 \item
-variational decoding for approximating the intractable MAP decoding \cite{variational-decoding-acl09},
+Improvements to decoding through variational methods and approximations to intractable MAP decoding \cite{variational-decoding-acl09},
 \item
-hypergraph-based discriminative training for better feature engineering \cite{zhifei-forest-reranking-galebook}, 
+Hypergraph-based discriminative training for better feature engineering \cite{zhifei-forest-reranking-galebook}, 
 \item
-visualization of the derivation trees and hypergraphs \cite{PBML-2010-Joshua-visualization}, 
+Visualization of the derivation trees and hypergraphs \cite{PBML-2010-Joshua-visualization}, 
 \item
-and a unified framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}.
+A unified framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}.
 \end{itemize}

### 2010-04-23 papers/wmt10/joshua-progress.tex
Revising Lane's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1628 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \item
-A unified framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}.
+A convenient framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}.
 \end{itemize}
@@
 
-Many research publications describe novel techniques and provide interesting results, yet fail to describe technical details in sufficient detail to allow their results to be reproduced by other researchers.
-Reproducibility in machine translation is made more challenging by the complexity of experimental workflows.
-Results in machine translation tasks are dependent on a cascade of processing steps and configurations.
-
-\newcite{Schwartz-wmt10-pipline} describes a workflow framework for designing and running reproducible machine translation experiments using Joshua.
-Each step in the machine translation workflow (data preprocessing, grammar training, MERT, decoding, etc) is modeled by a GNU Make script which defines how to run the tools used in that step, and an auxiliary configuration file that defines the exact parameters to be used in that step during a particular run of an experiment.
-Workflows configured using this framework allow a complete experiment to be run - from downloading data and software through truecasing translated results - by executing a single make file.
+Reproducing other researchers' machine translation experiments is difficult because the pipeline is too complex to fully detail in short conference papers.
+We have put together a workflow framework for designing and running reproducible machine translation experiments using Joshua \newcite{Schwartz-wmt10-pipline}.
+Each step in the machine translation workflow (data preprocessing, grammar training, MERT, decoding, etc) is modeled by a Make script that defines how to run the tools used in that step, and an auxiliary configuration file that defines the exact parameters to be used in that step for a particular experimental setup.
+Workflows configured using this framework allow a complete experiment to be run - from downloading data and software through scoring the final translated results - by executing a single Makefile.
 
 This framework encourages researchers to supplement research publications with links to the complete set of scripts and configurations that were actually used to run the experiment.
-The scripts and configurations used to run the Johns Hopkins University submission to the 2010 WMT shared translation task are provided in \newcite{Schwartz-wmt10-pipline}.
-
+The Johns Hopkins University submission to the WMT10 shared translation task was implemented in this framework, and can be easily and exactly reproduced by other researchers by typing a single command.
 \bibliographystyle{acl}

### 2010-04-23 papers/wmt10/joshua-progress.tex
Revising intro
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1629 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 % instead of just Hiero grammars.  This 
-including syntax augment machine translation (SAMT) grammars \cite{samt2006},
+including syntax augment machine translation (SAMT) grammars \cite{samt2006}
 \item
-Support for external modules to posit translations for spans in the input sentence that constrain decoding,    \cite{PBML-2010-Joshua-transliteration},  
+Support for external modules to posit translations for spans in the input sentence that constrain decoding   \cite{PBML-2010-Joshua-transliteration}
 \item
-Lattice parsing for dealing with input uncertainty, including ambiguous output from speech recognizers or Chinese word segmenters  \cite{dyer-muresan-resnik:2008:ACLMain},
+Lattice parsing for dealing with input uncertainty, including ambiguous output from speech recognizers or Chinese word segmenters  \cite{dyer-muresan-resnik:2008:ACLMain}
 \item
-Semiring parsing for unified dynamic programming \cite{li-eisner:2009:EMNLP} , 
+A semiring architecture over hypergraphs that allows many inference operations to be implemented easily and elegantly \cite{li-eisner:2009:EMNLP}
 \item
-Improvements to decoding through variational methods and approximations to intractable MAP decoding \cite{variational-decoding-acl09},
+Improvements to decoding through variational decoding and other approximate methods that overcome intractable MAP decoding \cite{variational-decoding-acl09}
 \item
-Hypergraph-based discriminative training for better feature engineering \cite{zhifei-forest-reranking-galebook}, 
+Hypergraph-based discriminative training for better feature engineering \cite{zhifei-forest-reranking-galebook}
 \item
-Visualization of the derivation trees and hypergraphs \cite{PBML-2010-Joshua-visualization}, 
+Visualization of the derivation trees and hypergraphs \cite{PBML-2010-Joshua-visualization}
 \item
-A convenient framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}.
+A convenient framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}
 \end{itemize}
-Below we give a short description for each of the new functions added.
+The sections below give a short description of each of these new functions.
 
@@
 
-\section{MT Pipeline}
+\section{Pipeline for Running MT Experiments}
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
anni section added
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1630 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \section{Specifying Constraints on Rules}
-\todo{anni}
+
+Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly in low-resource settings.
+We have implemented this functionality in Joshua by supporting an XML format for proposing alternate translation rules (constraints) for a particular word span to the decoder \cite{pbml-2010-Joshua-modules}.
+That is, processing that is separate from the MT engine can suggest translations for some set of source side words and phrases.
+The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights.
+We modified the chart-based decoder to support these constraints.
 
@@
 We have introduced tools for visualizing two of the main data structures used
-in Joshua \cite{PBML-2010-Joshua-visualization}. Each visualization tool is implemented as a
+in Joshua \cite{PBML-2010-Joshua-transliteration}. Each visualization tool is implemented as a
 stand-alone program that hooks into the pipeline, but they are distributed with

### 2010-04-23 papers/wmt10/joshua-progress.tex
Revising Zhifei's text
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1631 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \textbf{Joshua} is an open-source toolkit for parsing-based machine translation that is written in Java.
-The initial release of Joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.
-The Joshua 1.0 release also included re-implementations of suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.
+The initial release of Joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart parsing, $n$-gram language model integration, beam and cube pruning, and $k$-best extraction.
+The Joshua 1.0 release also included re-implementations of suffix array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.
 Additionally, it included parallel and distributed computing techniques for salability \cite{Joshua-old}.
@@
 \item
-Support for external modules to posit translations for spans in the input sentence that constrain decoding   \cite{PBML-2010-Joshua-transliteration}
+Support for external modules to posit translations for spans in the input sentence that constrain decoding   \cite{PBML-2010-Joshua-modules}
 \item
@@
 \end{itemize}
-The sections below give a short description of each of these new functions.
+The sections below give short descriptions for each of these new functions.
 
@@
 Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly in low-resource settings.
-We have implemented this functionality in Joshua by supporting an XML format for proposing alternate translation rules (constraints) for a particular word span to the decoder \cite{pbml-2010-Joshua-modules}.
+We have implemented this functionality in Joshua by supporting an XML format for proposing alternate translation rules (constraints) for a particular word span to the decoder \cite{PBML-2010-Joshua-modules}.
 That is, processing that is separate from the MT engine can suggest translations for some set of source side words and phrases.
@@
 particular foreign input sentence.
-Given a hypergraph, we may perform many atomic inference operations
-such as finding one-best, $k$-best, or expectations over the hypergraph.
-To perform each operation, we may implement a dedicated dynamic programming algorithm.
+Given a hypergraph, we may perform many atomic inference operations,
+such as finding one-best or $k$-best translations, or computing expectations over the hypergraph.
+For each such operation, we could implement a dedicated dynamic programming algorithm.
 However, a more general framework to specify these algorithms is semiring-weighted parsing \cite{semiringparsing}.
-Within this framework, we implemented all the algorithms (i.e.,
-inside, outside, and inside-outside speedup) described by \newcite{li-eisner:2009:EMNLP},
-and also the first-order expectation semiring \cite{eisner-expectation-semiring} and its
-second-order version \cite{li-eisner:2009:EMNLP}.
+We have implemented  the inside algorithm, the outside algorithm, and the inside-outside speedup described by \newcite{li-eisner:2009:EMNLP},
+plut the first-order expectation semiring \cite{eisner-expectation-semiring} and its
+second-order version \cite{li-eisner:2009:EMNLP}.  All of these use our newly implemented semiring framework.
 
@@
 %% use of expectation and variance semiring
-The first- and second-order expectation semirings can be used compute many interesting quantities over the hypergraph.
-These quantities include expected hypothesis length, feature expectation, entropy, cross-entropy, Kullback-Leibler divergence,
+The first- and second-order expectation semi-rings can also be used to compute many interesting quantities over hypergraphs.
+These quantities include expected translation length, feature expectation, entropy, cross-entropy, Kullback-Leibler divergence,
 Bayes risk, variance of hypothesis length, gradient of entropy and Bayes risk, covariance and Hessian matrix, and so on.
@@
 We have introduced tools for visualizing two of the main data structures used
-in Joshua \cite{PBML-2010-Joshua-transliteration}. Each visualization tool is implemented as a
+in Joshua \cite{PBML-2010-Joshua-visualization}. Each visualization tool is implemented as a
 stand-alone program that hooks into the pipeline, but they are distributed with

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Anni's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1632 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly in low-resource settings.
-We have implemented this functionality in Joshua by supporting an XML format for proposing alternate translation rules (constraints) for a particular word span to the decoder \cite{PBML-2010-Joshua-modules}.
-That is, processing that is separate from the MT engine can suggest translations for some set of source side words and phrases.
+Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly for low-resource languages.
+We have implemented an XML interface that allows external modules to propose alternate translation rules (constraints) for a particular word span to the decoder \cite{PBML-2010-Joshua-modules}.
+Processing that is separate from the MT engine can suggest translations for some set of source side words and phrases.
 The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights.
+Additionally the XML format allows constraints on the righthand side of SCFG rules, which allows constraints like forcing a particular span to be a NP.
 We modified the chart-based decoder to support these constraints.

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Chris Dyer's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1633 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights.
-Additionally the XML format allows constraints on the righthand side of SCFG rules, which allows constraints like forcing a particular span to be a NP.
-We modified the chart-based decoder to support these constraints.
+Additionally the XML format allows constraints on the lefthand side of SCFG rules, which allows constraints like forcing a particular span to be a NP.
+We modified Joshua's chart-based decoder to support these constraints.
 
@@
 
-In \joshua, we use a {\em hypergraph} (or {\em packed forest}) to compactly 
-represent the exponentially many derivation trees generated by the decoder for a 
-particular foreign input sentence.
+In \joshua, we use a hypergraph (or packed forest) to compactly 
+represent the exponentially many derivation trees generated by the decoder for an input sentence.
 Given a hypergraph, we may perform many atomic inference operations,
@@
 
-\section{Lattice Parsing}
+\section{Word Lattice Input}
 
-The bottom-up parsing algorithm used to generate the translation hypergraph has been generalized to support translation from not only an input sentence but from a \emph{word lattice} with runtime and memory overhead that is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.
-This enables the exploration of a distribution over input sentences during decoding, and the best translation is selected from among any of them.
-This functionality is useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or word segmenters.
+We generalized the bottom-up parsing algorithm that generates the translation hypergraph so that it supports translation from word lattices instead of just sentences.
+Our implementation has runtime and memory overhead that is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.
+This allows distribution over input sentences to be explored during decoding, and allows the best translation to be selected from among all of them.
+This is especially useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or word segmenters, which can encode their alternative analyses as confusion networks or lattices.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Chris Dyer's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1634 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-We generalized the bottom-up parsing algorithm that generates the translation hypergraph so that it supports translation from word lattices instead of just sentences.
-Our implementation has runtime and memory overhead that is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.
-This allows distribution over input sentences to be explored during decoding, and allows the best translation to be selected from among all of them.
-This is especially useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or word segmenters, which can encode their alternative analyses as confusion networks or lattices.
+We generalized the bottom-up parsing algorithm that generates the translation hypergraph so that it supports translation of word lattices instead of just sentences.
+Our implementation's runtime and memory overhead is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.
+Accepting lattice-based input allows the decoder to explore a distribution over input sentences,  allowing it to select the best translation from among all of them.
+This is especially useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or Chinese word segmenters, which can encode their alternative analyses as confusion networks or lattices.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Anni's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1635 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\section{Specifying Constraints on Rules}
+\section{Specifying Constraints on Translation}
 
@@
 The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights.
-Additionally the XML format allows constraints on the lefthand side of SCFG rules, which allows constraints like forcing a particular span to be a NP.
+In addition to specifying translations, the XML format allows constraints on the lefthand side of SCFG rules, which allows constraints like forcing a particular span to be translated as an NP.
 We modified Joshua's chart-based decoder to support these constraints.

### 2010-04-23 papers/wmt10/joshua-progress.tex
Revising Zhifei's text
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1636 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 among many distinct derivations (e.g., trees or
-segmentations). In principle, the goodness of a
+segmentations) that have the same yield. In principle, the goodness of a
 string is measured by the total probability of its
@@
 The first version of \joshua implemented the
-Viterbi approximation that measures the goodness 
-of a string using only its most probable derivation.
+Viterbi approximation, which measures the goodness 
+of a translation using only its most probable derivation.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Jonny's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1637 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \textbf{Joshua} is an open-source toolkit for parsing-based machine translation that is written in Java.
-The initial release of Joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart parsing, $n$-gram language model integration, beam and cube pruning, and $k$-best extraction.
-The Joshua 1.0 release also included re-implementations of suffix array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.
+The initial release of \joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart parsing, $n$-gram language model integration, beam and cube pruning, and $k$-best extraction.
+The \joshua 1.0 release also included re-implementations of suffix array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.
 Additionally, it included parallel and distributed computing techniques for salability \cite{Joshua-old}.
@@
 In addition to specifying translations, the XML format allows constraints on the lefthand side of SCFG rules, which allows constraints like forcing a particular span to be translated as an NP.
-We modified Joshua's chart-based decoder to support these constraints.
+We modified \joshua's chart-based decoder to support these constraints.
 
@@
 
-We have introduced tools for visualizing two of the main data structures used
-in Joshua \cite{PBML-2010-Joshua-visualization}. Each visualization tool is implemented as a
-stand-alone program that hooks into the pipeline, but they are distributed with
-Joshua. The first program is a hypergraph visualizer. The user can choose from
+We created tools for visualizing two of the main data structures used
+in \joshua \cite{PBML-2010-Joshua-visualization}. The first visualizer displays hypergraphs. The user can choose from
 a set of input sentences, then call the decoder to build the hypergraph.
-
-The second is a derivation-tree visualizer. Joshua can produce output formatted
-as a parse tree, where each nonterminal has been annotated with its source-side
+The second  visualizer displays derivation trees.
+Setting a flag in the configuration file causes the decoder to output parse trees instead of strings, where each nonterminal is annotated with its source-side
 span. The visualizer can read in multiple n-best lists in this format, then
 display the resulting derivation trees side-by-side. We have found that
-visually inspecting these derivation trees is advantageous in debugging many
+visually inspecting these derivation trees is useful for debugging
 grammars.
@@
 Reproducing other researchers' machine translation experiments is difficult because the pipeline is too complex to fully detail in short conference papers.
-We have put together a workflow framework for designing and running reproducible machine translation experiments using Joshua \newcite{Schwartz-wmt10-pipline}.
+We have put together a workflow framework for designing and running reproducible machine translation experiments using \joshua \newcite{Schwartz-wmt10-pipline}.
 Each step in the machine translation workflow (data preprocessing, grammar training, MERT, decoding, etc) is modeled by a Make script that defines how to run the tools used in that step, and an auxiliary configuration file that defines the exact parameters to be used in that step for a particular experimental setup.

### 2010-04-23 papers/wmt10/joshua-progress.tex
Added Juri's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1638 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\section{Support for Syntax Augmented MT}
-\todo{juri or ccb}
+\section{Support for Syntax-Augmented MT}
+The syntax-augmented approach to hierachical phrase-based machine
+translation introduced by \newcite{samt2006} aims at improving
+translation quality by introducing syntactic information into the
+decoding process. To this end, one side of the bitext is parsed and a
+syntax-augmented grammar containing additional nonterminal symbols is
+extracted. These new symbols are obtained by annotating the
+nonterminals in Hiero-style rules
+\begin{center}
+\begin{tabular}{c}
+  {\small \tt [X]} $\Rightarrow$ {\small \tt [X,1]} sans {\small \tt [X,2]} 
+  $\vert$ {\tt \small [X,1]} without {\small \tt [X,2]} \\
+\end{tabular}
+\end{center}
+extracted from the bitext, the nonterminals are labeled according
+to what constituents cover the nonterminal span on the parsed side of
+the bitext. 
+\begin{center}
+\begin{tabular}{c}
+  {\small \tt [NP] $\Rightarrow$  ne [V,1] pas | did not [V,1] } \\
+  {\small \tt [VP] $\Rightarrow$  a b [V,1] | c d [X,1] } 
+\end{tabular}
+\end{center}
+To ensure good coverage, SAMT concatenates syntactic constituents or
+uses CCG-style labels when there is no single constituent covering a
+span. The resulting grammar allows for derivations that take into
+account the syntactic
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Lane's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1639 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 The sections below give short descriptions for each of these new functions.
+
+\section{Support for Syntax-based Translation}
 
@@
 Reproducing other researchers' machine translation experiments is difficult because the pipeline is too complex to fully detail in short conference papers.
-We have put together a workflow framework for designing and running reproducible machine translation experiments using \joshua \newcite{Schwartz-wmt10-pipline}.
+We have put together a workflow framework for designing and running reproducible machine translation experiments using \joshua \cite{Schwartz-wmt10-pipline}.
 Each step in the machine translation workflow (data preprocessing, grammar training, MERT, decoding, etc) is modeled by a Make script that defines how to run the tools used in that step, and an auxiliary configuration file that defines the exact parameters to be used in that step for a particular experimental setup.
-Workflows configured using this framework allow a complete experiment to be run - from downloading data and software through scoring the final translated results - by executing a single Makefile.
-
+Workflows configured using this framework allow a complete experiment to be run -- from downloading data and software through scoring the final translated results -- by executing a single Makefile.
 This framework encourages researchers to supplement research publications with links to the complete set of scripts and configurations that were actually used to run the experiment.
-The Johns Hopkins University submission to the WMT10 shared translation task was implemented in this framework, and can be easily and exactly reproduced by other researchers by typing a single command.
+The Johns Hopkins University submission for the WMT10 shared translation task was implemented in this framework, so it can be easily and exactly reproduced.
 \bibliographystyle{acl}

### 2010-04-23 papers/wmt10/joshua-progress.tex
SAMT edits
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1640 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
   {\small \tt [X]} $\Rightarrow$ {\small \tt [X,1]} sans {\small \tt [X,2]} 
-  $\vert$ {\tt \small [X,1]} without {\small \tt [X,2]} \\
+  $\vert$ {\tt \small [X,1]} without {\small \tt [X,2]}
 \end{tabular}
 \end{center}
-extracted from the bitext, the nonterminals are labeled according
-to what constituents cover the nonterminal span on the parsed side of
+extracted from the bitext, the nonterminals are labeled according to
+which constituents cover the nonterminal span on the parsed side of
 the bitext.
@@
 \begin{tabular}{c}
-  {\small \tt [NP] $\Rightarrow$  ne [V,1] pas | did not [V,1] } \\
-  {\small \tt [VP] $\Rightarrow$  a b [V,1] | c d [X,1] } 
+  {\small \tt [VP]} $\Rightarrow$ {\small \tt [VBN,1]} sans {\small
+    \tt [NP,2]} $\vert$ $\mathit{VBN}_1$ without {\small \tt [NP,2]} \\
+  {\small \tt [NP]} $\Rightarrow$ {\small \tt [NP,1]} sans {\small
+    \tt [NP,2]} 
+  $\vert$ {\small \tt [NP,1]} without {\small \tt [NP,2]} 
 \end{tabular}
 \end{center}
-To ensure good coverage, SAMT concatenates syntactic constituents or
-uses CCG-style labels when there is no single constituent covering a
-span. The resulting grammar allows for derivations that take into
-account the syntactic
+To ensure good coverage, the SAMT approach concatenates syntactic
+constituents or uses CCG-style labels when there is no single
+constituent spanning a nonterminal. The resulting grammar
+
+\joshua's decoder supports SAMT-style grammars as well as non-syntactic
+nonterminal annotations. An SAMT grammar
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
More syntax edits.
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1641 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 constituents or uses CCG-style labels when there is no single
-constituent spanning a nonterminal. The resulting grammar
+constituent spanning a nonterminal. The derivations licensed by the
+resulting grammar are constrained to match the syntactic structure,
+typically leading to more consistent reorderings.
 
 \joshua's decoder supports SAMT-style grammars as well as non-syntactic
-nonterminal annotations. An SAMT grammar
+nonterminal annotations obtained from other sources, such as named
+entity taggers.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Juri's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1642 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

--- 2010-04-23 papers/wmt10/joshua-progress.tex
@@
 
-\section{Support for Syntax-Augmented MT}
-The syntax-augmented approach to hierachical phrase-based machine
-translation introduced by \newcite{samt2006} aims at improving
-translation quality by introducing syntactic information into the
-decoding process. To this end, one side of the bitext is parsed and a
-syntax-augmented grammar containing additional nonterminal symbols is
-extracted. These new symbols are obtained by annotating the
-nonterminals in Hiero-style rules
-\begin{center}
+The initial release of \joshua supported only Hiero-style SCFGs, which use a single nonterminal symbol X.
+This release includes support for arbitrary SCFGs, including ones that use a rich set of linguistic nonterminal symbols.
+In particular we have added support for \newcite{samt2006}'s syntax-augmented machine
+translation. SAMT grammar extraction is identical to Hiero grammar extraction, except that one side of the parallel is parsed, and syntactic labels replace the X
+nonterminals in Hiero-style rules.  Instead of extracting this Hiero rule from the bitext
+
+\noindent
 \begin{tabular}{c}
@@
 \end{tabular}
-\end{center}
-extracted from the bitext, the nonterminals are labeled according to
+
+\noindent
+the nonterminals can be labeled according to
 which constituents cover the nonterminal span on the parsed side of
-the bitext.
-\begin{center}
+the bitext.  This constrains what types of phrases the decoder can use when producing a translation.
+
+\noindent
 \begin{tabular}{c}
-  {\small \tt [VP]} $\Rightarrow$ {\small \tt [VBN,1]} sans {\small
-    \tt [NP,2]} $\vert$ $\mathit{VBN}_1$ without {\small \tt [NP,2]} \\
-  {\small \tt [NP]} $\Rightarrow$ {\small \tt [NP,1]} sans {\small
-    \tt [NP,2]} 
-  $\vert$ {\small \tt [NP,1]} without {\small \tt [NP,2]} 
+  {\small \tt [VP]} $\Rightarrow$ {\small \tt [VBN]} sans {\small
+    \tt [NP]} $\vert$ {\small \tt [VBN]} without {\small \tt [NP]} \\
+  {\small \tt [NP]} $\Rightarrow$ {\small \tt [NP]} sans {\small
+    \tt [NP]} 
+  $\vert$ {\small \tt [NP]} without {\small \tt [NP]} 
 \end{tabular}
-\end{center}
-To ensure good coverage, the SAMT approach concatenates syntactic
-constituents or uses CCG-style labels when there is no single
-constituent spanning a nonterminal. The resulting grammar
 
-\joshua's decoder supports SAMT-style grammars as well as non-syntactic
-nonterminal annotations. An SAMT grammar
+\noindent
+Unlike GHKM, SAMT has the same coverage as Hiero, because it allows non-constituent phrases to get syntactic
+labels using CCG-style slash notation. 
+The derivations created using syntactically motivated grammars exhibit more coherent syntactic structure than Hiero and typically result in better reordering, especially for languages with word orders that diverge from English, like Urdu.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Juri's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1643 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 \noindent
-Unlike GHKM, SAMT has the same coverage as Hiero, because it allows non-constituent phrases to get syntactic
+Unlike GHKM \cite{Galley2004}, SAMT has the same coverage as Hiero, because it allows non-constituent phrases to get syntactic
 labels using CCG-style slash notation.
-The derivations created using syntactically motivated grammars exhibit more coherent syntactic structure than Hiero and typically result in better reordering, especially for languages with word orders that diverge from English, like Urdu.
+Experimentally, we have found that the derivations created using syntactically motivated grammars exhibit more coherent syntactic structure than Hiero and typically result in better reordering, especially for languages with word orders that diverge from English, like Urdu \cite{SCALE-report}.
 

### 2010-04-23 papers/wmt10/joshua-progress.tex
Added funding acknowledgments
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1644 0ae5e6b2-d358-4f09-a895-f82f13dd62a4
Submitted version
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1645 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 support for translation grammars with a rich set of syntactic nonterminals,
-the ability for external modules to posit constraints on spans in the input sentence should be translated,  
+the ability for external modules to posit constraints on how spans in the input sentence should be translated,  
 lattice parsing for dealing with input uncertainty,
@@
 The Johns Hopkins University submission for the WMT10 shared translation task was implemented in this framework, so it can be easily and exactly reproduced.
+
+\section*{Acknowledgements}
+
+Research funding was provided by the NSF under grant IIS-0713448, by the European Commission through the EuroMatrixPlus project, and by the DARPA GALE program under Contract No.
+HR0011-06-2-0001.
+The views and findings are the authors' alone.
+
 \bibliographystyle{acl}

### 2010-04-23 papers/wmt10/joshua-progress.tex
Added MERT section.
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1646 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 hypergraph-based discriminative training for better feature engineering, 
+a parallelized MERT module,
+document-level and tail-based MERT,
 visualization of the derivation trees, 
@@
 Hypergraph-based discriminative training for better feature engineering \cite{zhifei-forest-reranking-galebook}
+\item
+A parallelization of MERT's computations, and supporting document-level and tail-based optimization \cite{atypical-mert}
 \item
@@
 
+\section{Minimum Error Rate Training}
+
+\joshua's MERT module optimizes parameter weights so as to maximize performance
+on a development set as measuered by an automatic evaluation metric, such as
+Bleu \cite{och-mert}.
+
+We have parallelized our MERT module in two ways: parallelizing the computation
+of metric scores, and parallelizing the search over parameters. The computation
+of metric scores is a computational concern when tuning to a metric that is
+slow to compute, such as translation edit rate \cite{snover-etal:2006:ter}.
+Since scoring a candidate is independent from scoring any other candidate, we
+parallelize this computation using a multi-threaded solution\footnote{Based on
+sample code by Kenneth Heafield.}. Similarly, we parallelize the optimization
+of the intermediate initial weight vectors, also using a multi-threaded
+solution.
+
+Another feature is the module's awareness of document information, and the
+capability to perform optimization of {\em document-based} variants of the
+automatic metric \cite{atypical-mert}. For example, in document-based Bleu,
+a Bleu score is calculated {\em for each document}, and the tuned score is the
+average of those document scores. The MERT module can furthermore be instructed
+to target a specific subset of those documents, namely the {\em tail} subset,
+where only the subset of documents with the lowest document Bleu scores are
+considered.\footnote{This feature is of interest to GALE teams, for instance,
+since GALE's evaluation criteria place a lot of focus on translation quality of
+tail documents.}
+
+More details on the MERT method and the
+implementation can be found in \newcite{Zaidan2009}.\footnote{The module
+is also available as a standalone application, {\em Z-MERT}, that can be used
+with other MT systems. (Software and documentation at: \url{http://cs.jhu.edu/~ozaidan/zmert}.)}
+
 \section{Visualization}

### 2010-04-23 papers/wmt10/joshua-progress.tex
Justified the longer paragraphs.
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1647 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\textbf{Joshua} is an open-source toolkit for parsing-based machine translation that is written in Java.
-The initial release of \joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart parsing, $n$-gram language model integration, beam and cube pruning, and $k$-best extraction.
-The \joshua 1.0 release also included re-implementations of suffix array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.
-Additionally, it included parallel and distributed computing techniques for salability \cite{Joshua-old}.
+\joshua is an open-source toolkit for parsing-based machine translation that is
+written in Java. The initial release of \joshua  \cite{joshua-wmt09} was
+a re-implementation of the Hiero system \cite{hiero-2007} and all its
+associated algorithms, including: chart parsing, $n$-gram language model
+integration, beam and cube pruning, and $k$-best extraction.  The \joshua 1.0
+release also included re-implementations of suffix array grammar extraction
+\cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction} and
+minimum error rate training \cite{och-mert,Zaidan2009}.  Additionally, it
+included parallel and distributed computing techniques for salability \cite{Joshua-old}.
 
-This paper describes the additions to the toolkit over the past year, which together form the 2.0 release.
-The software has been heavily used by the authors and several other groups in their daily research, and has been substantially refined since the first release.
-The most important new functions in the toolkit are:
+This paper describes the additions to the toolkit over the past year, which
+together form the 2.0 release. The software has been heavily used by the
+authors and several other groups in their daily research, and has been
+substantially refined since the first release.  The most important new
+functions in the toolkit are:
 \begin{itemize}
@@
 
-Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly for low-resource languages.
-We have implemented an XML interface that allows external modules to propose alternate translation rules (constraints) for a particular word span to the decoder \cite{PBML-2010-Joshua-modules}.
-Processing that is separate from the MT engine can suggest translations for some set of source side words and phrases.
-The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights.
-In addition to specifying translations, the XML format allows constraints on the lefthand side of SCFG rules, which allows constraints like forcing a particular span to be translated as an NP.
-We modified \joshua's chart-based decoder to support these constraints.
+Integrating output from specialized modules (like transliterators,
+morphological analyzers, and modality translators) into the MT pipeline can
+improve translation performance, particularly for low-resource languages. We
+have implemented an XML interface that allows external modules to propose
+alternate translation rules (constraints) for a particular word span to the
+decoder \cite{PBML-2010-Joshua-modules}. Processing that is separate from the
+MT engine can suggest translations for some set of source side words and
+phrases. The XML format allows for both hard constraints, which must be used,
+and soft constraints, which compete with standard extracted translation rules,
+as well as specifying associated feature weights. In addition to specifying
+translations, the XML format allows constraints on the lefthand side of SCFG
+rules, which allows constraints like forcing a particular span to be translated
+as an NP.  We modified \joshua's chart-based decoder to support these
+constraints.
 
@@
 
-We generalized the bottom-up parsing algorithm that generates the translation hypergraph so that it supports translation of word lattices instead of just sentences.
-Our implementation's runtime and memory overhead is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.
-Accepting lattice-based input allows the decoder to explore a distribution over input sentences,  allowing it to select the best translation from among all of them.
-This is especially useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or Chinese word segmenters, which can encode their alternative analyses as confusion networks or lattices.
+We generalized the bottom-up parsing algorithm that generates the translation
+hypergraph so that it supports translation of word lattices instead of just
+sentences.  Our implementation's runtime and memory overhead is proportional to
+the size of the lattice, rather than the number of paths in the lattice
+\cite{dyer-muresan-resnik:2008:ACLMain}.  Accepting lattice-based input allows
+the decoder to explore a distribution over input sentences,  allowing it to
+select the best translation from among all of them.  This is especially useful
+when \joshua is used to translate the output of statistical preprocessing
+components, such as speech recognizers or Chinese word segmenters, which can
+encode their alternative analyses as confusion networks or lattices.
 
@@
 
-Reproducing other researchers' machine translation experiments is difficult because the pipeline is too complex to fully detail in short conference papers.
-We have put together a workflow framework for designing and running reproducible machine translation experiments using \joshua \cite{Schwartz-wmt10-pipline}.
-Each step in the machine translation workflow (data preprocessing, grammar training, MERT, decoding, etc) is modeled by a Make script that defines how to run the tools used in that step, and an auxiliary configuration file that defines the exact parameters to be used in that step for a particular experimental setup.
-Workflows configured using this framework allow a complete experiment to be run -- from downloading data and software through scoring the final translated results -- by executing a single Makefile.
-This framework encourages researchers to supplement research publications with links to the complete set of scripts and configurations that were actually used to run the experiment.
-The Johns Hopkins University submission for the WMT10 shared translation task was implemented in this framework, so it can be easily and exactly reproduced.
+Reproducing other researchers' machine translation experiments is difficult
+because the pipeline is too complex to fully detail in short conference papers.
+We have put together a workflow framework for designing and running
+reproducible machine translation experiments using \joshua \cite{Schwartz-wmt10-pipline}.
+Each step in the machine translation workflow (data preprocessing, grammar
+training, MERT, decoding, etc) is modeled by a Make script that defines how to
+run the tools used in that step, and an auxiliary configuration file that
+defines the exact parameters to be used in that step for a particular
+experimental setup. Workflows configured using this framework allow a complete
+experiment to be run -- from downloading data and software through scoring the
+final translated results -- by executing a single Makefile.
+
+This framework encourages researchers to supplement research publications with
+links to the complete set of scripts and configurations that were actually used
+to run the experiment. The Johns Hopkins University submission for the WMT10
+shared translation task was implemented in this framework, so it can be easily
+and exactly reproduced.
 

### 2010-06-03 papers/wmt10/joshua-progress.tex
Fixed typos mentioned by reviewer 2, added citations for R-PROP and L-BFGS.
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1699 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

@@
 
-\joshua is an open-source toolkit for parsing-based machine translation that is
-written in Java. The initial release of \joshua  \cite{joshua-wmt09} was
-a re-implementation of the Hiero system \cite{hiero-2007} and all its
-associated algorithms, including: chart parsing, $n$-gram language model
-integration, beam and cube pruning, and $k$-best extraction.  The \joshua 1.0
-release also included re-implementations of suffix array grammar extraction
-\cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction} and
-minimum error rate training \cite{och-mert,Zaidan2009}.  Additionally, it
-included parallel and distributed computing techniques for salability \cite{Joshua-old}.
+\joshua is an open-source toolkit for parsing-based machine
+translation that is written in Java. The initial release of \joshua
+\cite{joshua-wmt09} was a re-implementation of the Hiero system
+\cite{hiero-2007} and all its associated algorithms, including: chart
+parsing, $n$-gram language model integration, beam and cube pruning,
+and $k$-best extraction.  The \joshua 1.0 release also included
+re-implementations of suffix array grammar extraction
+\cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}
+and minimum error rate training \cite{och-mert,Zaidan2009}.
+Additionally, it included parallel and distributed computing
+techniques for scalability \cite{Joshua-old}.
 
@@
 
-The initial release of \joshua supported only Hiero-style SCFGs, which use a single nonterminal symbol X.
-This release includes support for arbitrary SCFGs, including ones that use a rich set of linguistic nonterminal symbols.
-In particular we have added support for \newcite{samt2006}'s syntax-augmented machine
-translation. SAMT grammar extraction is identical to Hiero grammar extraction, except that one side of the parallel is parsed, and syntactic labels replace the X
-nonterminals in Hiero-style rules.  Instead of extracting this Hiero rule from the bitext
+The initial release of \joshua supported only Hiero-style SCFGs, which
+use a single nonterminal symbol X.  This release includes support for
+arbitrary SCFGs, including ones that use a rich set of linguistic
+nonterminal symbols.  In particular we have added support for
+\newcite{samt2006}'s syntax-augmented machine translation. SAMT
+grammar extraction is identical to Hiero grammar extraction, except
+that one side of the parallel corpus is parsed, and syntactic labels
+replace the X nonterminals in Hiero-style rules.  Instead of
+extracting this Hiero rule from the bitext
 
@@
 \noindent
-the nonterminals can be labeled according to
-which constituents cover the nonterminal span on the parsed side of
-the bitext.  This constrains what types of phrases the decoder can use when producing a translation.
+the nonterminals can be labeled according to which constituents cover
+the nonterminal span on the parsed side of the bitext.  This
+constrains what types of phrases the decoder can use when producing a
+translation.
 
@@
 \noindent
-Unlike GHKM \cite{Galley2004}, SAMT has the same coverage as Hiero, because it allows non-constituent phrases to get syntactic
-labels using CCG-style slash notation.
-Experimentally, we have found that the derivations created using syntactically motivated grammars exhibit more coherent syntactic structure than Hiero and typically result in better reordering, especially for languages with word orders that diverge from English, like Urdu \cite{SCALE-report}.
+Unlike GHKM \cite{Galley2004}, SAMT has the same coverage as Hiero,
+because it allows non-constituent phrases to get syntactic labels
+using CCG-style slash notation. Experimentally, we have found that the
+derivations created using syntactically motivated grammars exhibit
+more coherent syntactic structure than Hiero and typically result in
+better reordering, especially for languages with word orders that
+diverge from English, like Urdu \cite{SCALE-report}.
 
@@
 
-Discriminative training with a large number of features has 
-potential to improve the MT performance.
-We have implemented the hypergraph-based minimum risk training \cite{li-eisner:2009:EMNLP},
+Discriminative training with a large number of features has potential
+to improve the MT performance.  We have implemented the
+hypergraph-based minimum risk training \cite{li-eisner:2009:EMNLP},
 which minimizes the {\em expected loss} of the reference translations.
-The minimum-risk objective can be optimized by a gradient-based method, where
-the risk and its gradient can be computed using a second-order expectation semiring.
-For optimization, we use both L-BFGS\footnote{http://en.wikipedia.org/wiki/L-BFGS} 
-and Rprop\footnote{http://en.wikipedia.org/wiki/Rprop}.
+The minimum-risk objective can be optimized by a gradient-based
+method, where the risk and its gradient can be computed using a
+second-order expectation semiring.  For optimization, we use both
+L-BFGS \cite{lbfgs} and Rprop \cite{rprop}.
 
